### 2024-10-06 To_do_list.qmd
Added intitial files

---
format: pdf
---
### **Person A: Data Gathering, Exploration & Preprocessing**
- **Task 1: Data Collection & Preprocessing**
  - Download the "Presidential general election polls (current cycle)" data.
  - Explore the dataset to understand its structure, variables, and context.
  - Perform data cleaning (e.g., handle missing values, outliers).
  - Document any high-level cleaning steps and ensure a reproducible data preparation workflow.

- **Task 2: Exploratory Data Analysis (EDA)**
  - Generate summary statistics for all relevant variables.
  - Create visualizations (e.g., histograms, boxplots, scatter plots) to understand variable distributions and relationships.
  - Identify key predictors to include in the model, and justify their selection.

- **Task 3: Pollster Deep-Dive for Appendix A**
  - Pick a pollster from the data sample.
  - Research the methodology of this pollster, focusing on survey design (population, frame, sample recruitment, etc.).
  - Write a detailed overview of the survey’s strengths, weaknesses, and key features.

- **Task 4: Data Section Write-Up**
  - Write a section on the dataset, including background, context, and detailed examination of variables.
  - Include all visualizations and summary statistics in a clear format.

---

### **Person B: Model Building, Analysis, & Validation**
- **Task 1: Model Development**
  - Decide on the appropriate model to use (linear or generalized linear model).
  - Justify the model choice based on the dataset and research goals.
  - Implement the model in R, Python, or a combination, and ensure reproducibility.

- **Task 2: Model Explanation & Assumptions**
  - Clearly define all model components using mathematical notation and plain English explanations.
  - Discuss and document underlying assumptions, limitations, and potential issues.
  - Validate model performance using techniques like test/train splits, RMSE calculations, or sensitivity analyses.

- **Task 3: Results Section Write-Up**
  - Create tables, graphs, and visualizations to present the model's findings.
  - Write a clear and concise interpretation of the model results.

- **Task 4: Discussion Section Write-Up**
  - Discuss key takeaways from the model and results.
  - Highlight what the paper contributes to understanding the election forecast.
  - Address any weaknesses and propose future steps or improvements.

---

### **Person C: Appendices, Documentation & Reproducibility**
- **Task 1: Idealized Survey Design for Appendix B**
  - Design an ideal methodology and survey for forecasting the U.S. presidential election on a $100K budget.
  - Detail all aspects of the methodology (sampling approach, recruitment, data validation).
  - Create the survey on a platform like Google Forms, include a link in the appendix, and export the survey content as a PDF.

- **Task 2: Reproducible Workflow & Documentation**
  - Create an organized folder structure in the GitHub repo (e.g., `data/`, `scripts/`, `figures/`, `docs/`).
  - Write a detailed README file explaining the project, data sources, model, and workflow.
  - Document any use of LLMs or other tools, and include a separate usage text file if necessary.

- **Task 3: Final Paper Integration & Formatting**
  - Combine and format sections from all team members into a cohesive paper.
  - Add cross-references to all figures, tables, and equations.
  - Ensure all required sections are present and adhere to the rubric guidelines.

- **Task 4: Quality Checks & Submissions**
  - Proofread the paper for typos, grammar, and formatting issues.
  - Verify that all elements are properly cited (R, data sources, references).
  - Check that the paper, repo, and all files do not show any evidence of being a class project.

---

# Timeline
### **Week 1: October 7th - October 13th**

- **Person A: Data Gathering & Preprocessing**
  - **By October 9th:** Download and inspect the "Presidential general election polls" dataset. Familiarize yourself with its structure and context.
  - **By October 11th:** Complete initial data cleaning (handle missing values and outliers).
  - **By October 13th:** Document all preprocessing steps and ensure the workflow is reproducible. Provide a summary of any challenges encountered and adjustments made to the data.

---

### **Week 2: October 14th - October 20th**
- **Person A: Exploratory Data Analysis (EDA)**
  - **By October 15th:** Generate summary statistics for all key variables (e.g., mean, median, distribution).
  - **By October 17th:** Create visualizations (e.g., histograms, boxplots) to explore the distributions and relationships between predictors and outcomes.
  - **By October 19th:** Identify and justify key predictors to be used in the model based on EDA findings.
  - **By October 20th:** Draft a summary of the findings from the EDA and prepare the data section's visualizations and statistics.

- **Person C: Initial Workflow Setup**
  - **By October 18th:** Set up an organized GitHub repo structure (`data/`, `scripts/`, `figures/`, `docs/`), and add all current progress and documentation.

---

### **Week 3: October 21st - October 22nd**
- **Person A: Pollster Deep-Dive for Appendix A**
  - **By October 21st:** Choose a pollster and research their methodology in-depth, focusing on their survey design and approach.
  - **By October 22nd:** Draft the Appendix A content with a detailed overview of the pollster’s methodology, strengths, weaknesses, and key features.

- **Majority of Data Tasks Due on October 22nd**
  - Data gathering, exploration, and pollster analysis should be completed and ready for initial review.

---

### **Week 4: October 23rd - October 29th**
- **Person B: Model Building & Initial Analysis**
  - **By October 25th:** Decide on and justify the type of model (linear or generalized linear) based on research goals.
  - **By October 26th:** Implement the model in R or Python and ensure the reproducibility of the code.
  - **By October 28th:** Begin validating the model by conducting performance checks (e.g., test/train split, RMSE).
  - **By October 29th:** Write a draft of the results section, including visualizations and tables.

- **Person C: Idealized Survey Design for Appendix B**
  - **By October 27th:** Develop a survey methodology for forecasting the election, based on a $100K budget.
  - **By October 29th:** Create the survey on Google Forms, add the link to the appendix, and export a PDF version.

---

### **Week 5: October 30th - November 4th**
- **Person B: Finalize Model & Discussion Section**
  - **By October 31st:** Finalize model performance analysis and diagnostics.
  - **By November 1st:** Write and refine the discussion section, including key takeaways, weaknesses, and future steps.

- **Person C: Final Integration, Quality Checks & Review**
  - **By November 2nd:** Integrate all sections (data, model, discussion) into a cohesive document. Ensure all cross-references, figures, and tables are correctly placed.
  - **By November 3rd:** Conduct quality checks on grammar, formatting, citations, and ensure that there is no evidence of this being a class project.
  - **By November 4th:** Final proofreading and submission of the polished paper.

---

### 2024-10-06 paper/paper.qmd
Added intitial files

---
title: "Forecasting the 2024 US Presidential Election"
subtitle: "My subtitle if needed"
author: 
  - Tina Kim
  - David Flores
  - Kevin Shao
thanks: "Code and data are available at: https://github.com/DavidFJ207/USPresidentialForecast"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

*Overview paragraph:* Provide a brief overview of the context of the upcoming US presidential election and the importance of forecasting its results.

*Estimand paragraph:* Define the estimand clearly (e.g., predicting the probability of a candidate winning the election based on poll data).

*Results paragraph:* Summarize the key findings of the model, highlighting its predictive accuracy and implications.

*Why it matters paragraph:* Explain the broader significance of accurately forecasting election results for politics, society, and policymaking.

*Telegraphing paragraph:* 
The remainder of this paper is structured as follows: Section 2 details the data and measurement process; Section 3 covers model development and results; Section 4 discusses implications and future steps.
The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview
We use the statistical programming language R [@citeR] and utilize the "Presidential general election polls" dataset [@source].

*Person A:* Provide an overview of the dataset's context, sources, and key variables.

## Measurement
Describe how the real-world phenomenon (voter preference) is translated into a dataset entry.

*Person A:* Discuss the survey design, the process of data collection, and potential biases or errors in measurement.

## Outcome variables
Add graphs, tables, and text for each outcome variable. Use subsections as needed.

Some of our data is (@fig-outcome), from @source

```{r}
#| label: fig-outcome
#| fig-cap: Example Outcome Variable Visualization
#| echo: false
# Visualization code

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables
Add graphs, tables, and text to describe key predictors.

*Person A:*: Visualize variable distributions (e.g., histograms, boxplots), and justify the inclusion of predictors in the model.

```{r}
#| label: fig-predictors
#| fig-cap: Example Predictor Variable Visualization
#| echo: false
# Visualization code

```

# Pollster Methodology Overview (Appendix A)
Append a detailed review of a chosen pollster's methodology, survey techniques, strengths, and weaknesses.

# Model {#sec-model}

## Model Development
The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election.

*Person B:* Choose a linear or generalized linear model. Justify the selection based on your research goal and data structure.

## Model Set-up
Define the model mathematically and contextually (e.g., linear model predicting vote share).

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i x_i \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

*Person B*: Define all model components and assumptions.

## Model Justification
Our results are summarized in @tbl-modelresults. Discuss the implications of the findings and their predictive accuracy.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Forecasting model results based on poll data"
#| echo: false

# Display results of the model
```
*Person B*: Visualize the model's results and include any performance metrics (e.g., RMSE, test/train split).
```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: ""
#| warning: false

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

Discuss what the model reveals about the election forecast and its potential impact on understanding voting behavior.

*Person B*: Discuss limitations of the model and areas for further improvement.

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

## Appendix A: Pollster Methodology Overview
*Person A*: Provide a detailed methodology review for the chosen pollster, including survey design, population sampling, non-response handling, etc.

## Appendix B: Idealized Survey Design for $100K Budget
*Person C*: Develop and describe an ideal survey design for forecasting the election with a $100K budget, and include survey link.

## Additional Data & Model Details
Include any technical details on data cleaning, model diagnostics, and posterior checks.

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-07 scripts/02-download_data.R
Updated the download_data script with code to download raw president polls data.

@@ -1,26 +1,34 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Tina Kim
+# Date: 7 October 2024 
+# Contact: tinak.kim@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
-library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
+raw_president_polls_data <-
+  read_csv(
+    file = 
+      "https://projects.fivethirtyeight.com/polls/data/president_polls.csv",
+    show_col_types = FALSE,
+  )
+
+write_csv(
+  x = raw_president_polls_data,
+  file = "president_polls.csv"
+)
 
 
 #### Save data ####
-# [...UPDATE THIS...]
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+write_csv(raw_president_polls_data, "data/01-raw_data/raw_data.csv") 
 
          

### 2024-10-07 scripts/02-download_data.R
Updated the preamble and removed some comments.

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
+# Purpose: Downloads and saves the presidential general election polls data 
+# from FiveThirtyEight
 # Author: Tina Kim
 # Date: 7 October 2024 
 # Contact: tinak.kim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
@@ -13,7 +14,6 @@ library(tidyverse)
 
 
 #### Download data ####
-
 raw_president_polls_data <-
   read_csv(
     file = 
@@ -28,7 +28,4 @@ write_csv(
 
 
 #### Save data ####
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(raw_president_polls_data, "data/01-raw_data/raw_data.csv") 
-
-         
+write_csv(raw_president_polls_data, "data/01-raw_data/raw_data.csv")

### 2024-10-08 scripts/03-clean_data.R
Updated the clean_data script with code that chooses the pollster with a good balance of high number of observations and a low number of missing values. Empty columns were removed.

@@ -1,44 +1,56 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw presidential general election polls data into 
+# an analysis dataset
+# Author: Tina Kim
+# Date: 7 October 2024
+# Contact: tinak.kim@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+# Find the top 3 most frequent pollster in the raw data
+top_pollsters <- raw_data %>%
+  count(pollster, sort = TRUE) %>%
+  top_n(3, n) %>%
+  pull(pollster)
+
+cleaned_data <- raw_data %>%
+  # Filter the raw data to only top 3 most frequent pollster names
[…]

### 2024-10-08 scripts/03-clean_data.R
Updated comments on clean_data script

@@ -41,14 +41,15 @@ missing_values_summary <- cleaned_data %>%
   # Order from smallest to biggest ratio value
   arrange(missing_ratio)
 
-# Find the pollster with the lowest missing values ratio because it has 
+# Find the pollster with the lowest missing values ratio because it has
 # the most complete data which could lead to a more reliable analysis
 smallest_ratio_pollster <- missing_values_summary %>%
   slice(1) %>%
   pull(pollster)
 
-# Filter cleaned data to include only the smallest ratio pollster 
-# which we will use for our data analysis
+# Filter cleaned data to include only the smallest ratio pollster.
+# This pollster is a good balance of high number of observations and a low
+# number of missing values.
 cleaned_data <- cleaned_data %>%
   filter(pollster == smallest_ratio_pollster)
 

### 2024-10-12 scripts/03-api-extract-clean_data.R
added function to extract and clean google sheets

@@ -1,0 +1,139 @@
+#### Preamble ####
+# Purpose: Extracts data from links into a summarize dataset
+# an analysis dataset
+# Author: Gadiel David Flores
+# Date: 7 October 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to cleaned poll data
+# Any other information needed? Intall googledrive and googleAuthR
+
+# Authenticate your Google account
+gs4_auth()
+
+#### Load Libraries ####
+library(googledrive)
+library(readxl)
+library(tidyverse)
+library(dplyr)
+
+#### Helper Functions ####
+
+# Function to download and read a Google Sheet
+download_and_read_google_sheet <- function(url) {
+  # Download the file from the Google Sheets URL
+  file <- drive_download(as_id(url), path = "temp_file.xlsx", overwrite = TRUE)
+  # Read the file
+  data <- read_excel(file$local_path)
+  # Return the data
+  return(data)
+}
+
+# Function to replace NA values with the previous non-NA value
+replace_na_with_previous <- function(df, column_index) {
+  last_non_na <- NA  # Variable to store the last non-NA value
+  
+  for (i in 1:nrow(df)) {
+    if (is.na(df[[column_index]][i])) {
+      df[[column_index]][i] <- last_non_na  # Replace NA with last non-NA value
+    } else {
+      last_non_na <- df[[column_index]][i]  # Update last non-NA value
+    }
+  }
+  
+  return(df)
+}
+
+
+#### Main Function to Process Google Sheet ####
+process_google_sheet <- function(sheet_url, output_path) {
+  # Step 1: Download and read the Google Sheet
+  sheet_data <- download_and_read_google_sheet(sheet_url)
+  
+  # Step 2: Define unwanted values to clean
+  unwanted_values <- c("Valid", "Missing", "Total", "Combined Presidential Vote", "Donald Trump")
+  
+  # Step 3: Clean the first column by replacing unwanted values with NA
+  cleaned_sheet_data <- sheet_data %>%
+    mutate(
+      first_column = if_else(sheet_data[[1]] %in% unwanted_values, NA_character_, sheet_data[[1]])
+    )
+  
+  # Step 4: Drop the original first column and replace with the cleaned version
+  cleaned_sheet_data[[1]] <- cleaned_sheet_data$first_column
+  cleaned_sheet_data <- cleaned_sheet_data %>%
+    select(-first_column)
+  
[…]

### 2024-10-12 scripts/03-clean_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw presidential general election polls data into 
+# Purpose: Cleans the raw presidential general election polls data into usable links
 # an analysis dataset
 # Author: Tina Kim
 # Date: 7 October 2024
@@ -11,7 +11,7 @@
 #### Workspace setup ####
 library(tidyverse)
 
-#### Clean data ####
+#### Clean Raw data ####
 raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
 # Find the top 3 most frequent pollster in the raw data
@@ -54,4 +54,4 @@ cleaned_data <- cleaned_data %>%
   filter(pollster == smallest_ratio_pollster)
 
 #### Save data ####
-write_csv(cleaned_data, "data/02-analysis_data/analysis_data.csv")
+write_csv(cleaned_data, "data/02-analysis_data/cleaned_poll_data.csv")

### 2024-10-12 scripts/00-simulate_data.R
Fixed simulation to be more accurate

@@ -14,40 +14,52 @@ set.seed(853)
 
 
 #### Simulate data ####
-# State names
-states <- c(
-  "California",
-  "Texas",
-  "Florida",
-  "New York",
-  "Pennsylvania",
-  "Illinois",
-  "Ohio",
-  "Georgia",
-  "North Carolina",
-  "Michigan"
-)
 
-# Political parties
-parties <- c("Republican", "Democrat", "Other")
+# Helper function to generate random percentages for responses that add up to 100%
+generate_percentages <- function(n) {
+  p <- runif(n)
+  return(round(p / sum(p) * 100, 2))  # Normalize to sum to 100% and round to 2 decimal places
+}
 
-# Other parameters
-methodologies <- c("Phone", "Online", "Mixed")
-region_types <- c("Urban", "Suburban", "Rural")
+# Simulate data
+states <- c("California", "Texas", "Florida", "New York", "Pennsylvania", "Illinois", "Ohio", "Georgia", "North Carolina", "Michigan")
 
-# Create a dataset by randomly assigning states and parties to divisions
+# Simulate percentages for each response type outside the data.frame
+Joe_Biden_Approval_Approve <- generate_percentages(10)
+Joe_Biden_Approval_Disapprove <- 100 - Joe_Biden_Approval_Approve
+
+Presidential_Choice_Harris <- generate_percentages(10)
+Presidential_Choice_Trump <- 100 - Presidential_Choice_Harris
+
+Safety_Concerns_More_Safe <- generate_percentages(10)
+Safety_Concerns_Less_Safe <- 100 - Safety_Concerns_More_Safe
+
+Economic_Situation_Better <- generate_percentages(10)
+Economic_Situation_Worse <- 100 - Economic_Situation_Better
+
+Past_Voting_Behavior_Biden <- generate_percentages(10)
+Past_Voting_Behavior_Trump <- 100 - Past_Voting_Behavior_Biden
+
+Ethnicity_Hispanic <- generate_percentages(10)
+Ethnicity_White <- 100 - Ethnicity_Hispanic
+
+# Create the dataset by including the simulated data
 analysis_data <- data.frame(
-  Date = seq(as.Date("2023-06-01"), as.Date("2023-08-31"), length.out = 10),
-  State = states,
-  Candidate_A = round(runif(10, 40, 60), 2), 
[…]

### 2024-10-12 scripts/03-api-extract-clean_data.R
fixed file names

@@ -97,7 +97,7 @@ process_google_sheet <- function(sheet_url, output_path) {
 }
 
 #### State Dataset ####
-state_data <- read_csv("data/02-analysis_data/processed_poll_data.csv")
+state_data <- read_csv("data/02-analysis_data/states_surveys_links_data.csv")
 mega_dataset <- list()
 
 # Function to check if a string is a Google Sheets link

### 2024-10-12 scripts/03-clean-links-state-data.R

@@ -1,0 +1,31 @@
+library(tidyverse)
+
+# Load the poll links data
+poll_links_data <- read_csv("data/02-analysis_data/survey_list_data.csv")
+
+
+# Create a new dataset with unique states and URLs and their corresponding party columns
+new_poll_data <- poll_links_data %>%
+  # Group by state
+  group_by(state) %>%
+  # Rank the URLs for each state
+  mutate(url_rank = row_number()) %>%
+  # Select relevant columns
+  select(state, url_crosstab, sample_size, party, url_rank) %>%
+  # Spread URLs and their corresponding parties into separate columns based on the rank
+  pivot_wider(
+    names_from = url_rank,
+    values_from = c(url_crosstab, party),
+    names_prefix = c("url_crosstab_", "party_")
+  ) %>%
+  # Calculate the average sample size for each state
+  group_by(state) %>%
+  summarize(
+    across(starts_with("url_crosstab"), ~ first(na.omit(.)), .names = "{col}"),
+    across(starts_with("party"), ~ first(na.omit(.)), .names = "{col}"),
+    average_sample_size = mean(sample_size, na.rm = TRUE)
+  ) %>%
+  ungroup()
+
+# Optionally, save the new dataset to a CSV file
+write_csv(new_poll_data, "data/02-analysis_data/processed_poll_data.csv")

### 2024-10-12 scripts/03-clean_data.R

@@ -54,4 +54,4 @@ cleaned_data <- cleaned_data %>%
   filter(pollster == smallest_ratio_pollster)
 
 #### Save data ####
-write_csv(cleaned_data, "data/02-analysis_data/cleaned_poll_data.csv")
+write_csv(cleaned_data, "data/02-analysis_data/survey_list_data.csv")

### 2024-10-14 scripts/03-api-extract-clean_data.R
Added final summarized dataset

@@ -11,13 +11,19 @@
 # Authenticate your Google account
 gs4_auth()
 
-#### Load Libraries ####
+#### Load Libraries and files ####
 library(googledrive)
 library(readxl)
 library(tidyverse)
 library(dplyr)
-
+state_data <- read_csv("data/02-analysis_data/processed_poll_data.csv")
+mega_dataset_combined <- read_csv("data/02-analysis_data/mega_dataset.csv")
 #### Helper Functions ####
+
+# Function to check if a string is a Google Sheets link
+is_google_sheet_link <- function(text) {
+  grepl("docs.google.com/spreadsheets", text)
+}
 
 # Function to download and read a Google Sheet
 download_and_read_google_sheet <- function(url) {
@@ -96,14 +102,70 @@ process_google_sheet <- function(sheet_url, output_path) {
   return(merged_data)
 }
 
-#### State Dataset ####
-state_data <- read_csv("data/02-analysis_data/states_surveys_links_data.csv")
-mega_dataset <- list()
+#### Create final dataset ####
 
-# Function to check if a string is a Google Sheets link
-is_google_sheet_link <- function(text) {
-  grepl("docs.google.com/spreadsheets", text)
+item_counts <- table(mega_dataset_combined[[1]])
+# Get the unique items from the first column
+unique_items_column1 <- unique(mega_dataset_combined[[1]])
+# Convert the table to a dataframe
+item_counts_df <- as.data.frame(item_counts)
+# Rename the columns for clarity
+colnames(item_counts_df) <- c("Item", "Count")
+# Filter out rows where the count is less than 40
+item_counts_filtered <- item_counts_df %>%
+  filter(Count >= 49)
+
+# create dataset with correct headers
+item_list <- as.character(item_counts_filtered$Item)
+state_survey_data <- data.frame(matrix(ncol = length(item_list) + 1, nrow = 0))
+colnames(state_survey_data) <- c("state", item_list)
+names <- state_survey_data
+state_survey_data$state <- as.character(state_survey_data$state)
+state_data$state <- as.character(state_data$state)
+
+# Now append the state column from state_data to state_survey_data
+state_survey_data <- bind_rows(state_survey_data, data.frame(state = state_data$state))
+#### Loop to append data ####
+# Iterate through the entire state_data dataframe
+for (i in seq_len(nrow(state_data))) {
+  for (j in seq_len(ncol(state_data))) {
+    cell_value <- as.character(state_data[i, j])  # Get the cell value as a string
[…]

### 2024-10-14 scripts/03-clean-links-state-data.R

@@ -1,31 +1,48 @@
+#### Preamble ####
+# Purpose: Organizes states poll links
+# an analysis dataset
+# Author: Gadiel David Flores Jimenez
+# Date: 7 October 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to cleaned poll data
+# Any other information needed? Intall googledrive and googleAuthR
 library(tidyverse)
 
 # Load the poll links data
 poll_links_data <- read_csv("data/02-analysis_data/survey_list_data.csv")
 
-
-# Create a new dataset with unique states and URLs and their corresponding party columns
-new_poll_data <- poll_links_data %>%
-  # Group by state
+#### Pivot for URL crosstabs ####
+url_pivot <- poll_links_data %>%
   group_by(state) %>%
-  # Rank the URLs for each state
   mutate(url_rank = row_number()) %>%
-  # Select relevant columns
-  select(state, url_crosstab, sample_size, party, url_rank) %>%
-  # Spread URLs and their corresponding parties into separate columns based on the rank
+  select(state, url_crosstab, url_rank) %>%
   pivot_wider(
     names_from = url_rank,
-    values_from = c(url_crosstab, party),
-    names_prefix = c("url_crosstab_", "party_")
-  ) %>%
-  # Calculate the average sample size for each state
-  group_by(state) %>%
-  summarize(
-    across(starts_with("url_crosstab"), ~ first(na.omit(.)), .names = "{col}"),
-    across(starts_with("party"), ~ first(na.omit(.)), .names = "{col}"),
-    average_sample_size = mean(sample_size, na.rm = TRUE)
+    values_from = url_crosstab,
+    names_prefix = "url_crosstab_"
   ) %>%
   ungroup()
 
-# Optionally, save the new dataset to a CSV file
-write_csv(new_poll_data, "data/02-analysis_data/processed_poll_data.csv")
+#### Shift non-NA values to the left in relevant columns ####
+# Define columns to shift
+columns_to_shift <- names(url_pivot)[!(names(url_pivot) %in% c("state"))]
+
+# Function to shift values to the left
+shift_left <- function(row) {
+  non_na_values <- row[!is.na(row)]  # Keep only non-NA values
+  c(non_na_values, rep(NA, length(row) - length(non_na_values)))  # Fill the remaining with NA
+}
+
+# Apply shift_left function
[…]

### 2024-10-15 paper/paper.qmd
Updated data reference and data section in the paper.

@@
 ## Overview
-We use the statistical programming language R [@citeR] and utilize the "Presidential general election polls" dataset [@source].
+We obtained the "Presidential general election polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and used the statistical programming language R [@citeR] to analyze this data.
 

### 2024-10-16 scripts/03-api-extract-clean_data.R
made cleaned dataset have more values

@@ -161,8 +161,6 @@ for (i in seq_len(nrow(state_data))) {
     }
   }
 }
-
-
 write_csv(state_survey_data, "data/02-analysis_data/summarized_state_poll_data.csv")
 #### State Dataset ####
 mega_dataset <- list()
@@ -193,3 +191,23 @@ for (i in seq_len(nrow(state_data))) {
 mega_dataset_combined <- bind_rows(mega_dataset)
 # Save the dataset to a CSV file
 write_csv(mega_dataset_combined, "data/02-analysis_data/mega_dataset.csv")
+
+#### Organize and delete Cleaned Dataset ####
+state_polls <- read_csv("data/02-analysis_data/summarized_state_poll_data.csv")
+# Delete totals of states
+state_polls <- state_polls %>%
+  select(-matches(".*Total$"))
+
+# Delete empty states
+state_polls <- state_polls[-c(26:44), ]
+
+# Change the value in cell 1 to "National"
+state_polls[1, 1] <- "National"
+
+# Delete columns with more than 10 NAs
+state_polls_clean <- state_polls %>%
+  select(where(~ sum(is.na(.)) <= 15))
+
+
+# Save CSV file
+write_csv(state_polls_clean, "data/02-analysis_data/non_missing_state_poll_data.csv")

### 2024-10-17 paper/paper.qmd
Added Data description

@@
 
+Our objective was to clean, organize, and analyze polling data for the U.S. presidential general election. Below is an outline of the steps we took to achieve this goal:
+
+### 1. **Cleaning Raw Poll Data**
+We began by identifying the three most frequent pollsters within the dataset, focusing on these pollsters to improve data reliability and relevance.
+Missing values were assessed for each pollster, and the pollster with the fewest missing values was selected to ensure more complete data for further analysis.
+This refined dataset was then saved for the next stages of the process.
+
+### 2. **Organizing Poll Links by State**
+Next, we organized the polling data by state, ranking the links for each state to facilitate easier access.
+To ensure clarity, missing values were shifted to the end of each state’s record, and unnecessary empty columns were removed.
+This resulted in a more structured and user-friendly dataset, which was stored for future use.
+
+### 3. **Extracting and Merging Poll Data from Links**
+Following that, we extracted and merged additional polling data from Google Sheets links included in the dataset.
+After confirming the validity of these links, we cleaned the data by removing unwanted entries and filled in any missing values where possible.
+This updated dataset was then merged with the original dataset to create a more comprehensive compilation of polling data.
+
+### 4. **Final Organization and Refinement**
+In the final stage, we further refined the dataset by removing unnecessary rows (such as totals and empty states) and renaming specific entries for clarity (e.g., changing "State" to "National").
+Columns with excessive missing values were also eliminated to ensure that only the most reliable and complete data remained.
+This final, polished dataset was saved, ready for subsequent analysis and reporting.
+
 ## Measurement
@@
 
-*Person A:* Discuss the survey design, the process of data collection, and potential biases or errors in measurement.
+*Person C:* Discuss the survey design, the process of data collection, and potential biases or errors in measurement.
 
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 
-*Person A:*: Visualize variable distributions (e.g., histograms, boxplots), and justify the inclusion of predictors in the model.
+*Person B:*: Visualize variable distributions (e.g., histograms, boxplots), and justify the inclusion of predictors in the model.
 
@@
 
-*Person B*: Discuss limitations of the model and areas for further improvement.
+*Person C*: Discuss limitations of the model and areas for further improvement.
 
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-10-17 scripts/03-api-extract-clean_data.R

@@ -206,7 +206,7 @@ state_polls[1, 1] <- "National"
 
 # Delete columns with more than 10 NAs
 state_polls_clean <- state_polls %>%
-  select(where(~ sum(is.na(.)) <= 15))
+  select(where(~ sum(is.na(.)) <= 2))
 
 
 # Save CSV file

### 2024-10-19 scripts/03-api-extract-clean_data.R
Added final cleaned dataset

@@ -51,7 +51,7 @@ replace_na_with_previous <- function(df, column_index) {
 }
 
 
-#### Main Function to Process Google Sheet ####
+#### Main Function to Process Google and Excel Sheets####
 process_google_sheet <- function(sheet_url, output_path) {
   # Step 1: Download and read the Google Sheet
   sheet_data <- download_and_read_google_sheet(sheet_url)
@@ -101,7 +101,59 @@ process_google_sheet <- function(sheet_url, output_path) {
   # Return the final merged dataset for viewing
   return(merged_data)
 }
-
+process_excel_sheet <- function(excel_url, output_path) {
+  # Step 1: Download and read the Excel file
+  temp_file <- tempfile(fileext = ".xlsx")
+  download.file(excel_url, temp_file, mode = "wb")
+  sheet_data <- read_excel(temp_file)
+  
+  # Step 2: Define unwanted values to clean
+  unwanted_values <- c("Valid", "Missing", "Total", "Combined Presidential Vote", "Donald Trump")
+  
+  # Step 3: Clean the first column by replacing unwanted values with NA
+  cleaned_sheet_data <- sheet_data %>%
+    mutate(
+      first_column = if_else(sheet_data[[1]] %in% unwanted_values, NA_character_, sheet_data[[1]])
+    )
+  
+  # Step 4: Drop the original first column and replace it with the cleaned version
+  cleaned_sheet_data[[1]] <- cleaned_sheet_data$first_column
+  cleaned_sheet_data <- cleaned_sheet_data %>%
+    select(-first_column)
+  
+  # Step 5: Create a new header row
+  column_names_list <- colnames(sheet_data)
+  new_header <- as.data.frame(matrix(ncol = length(column_names_list), nrow = 1))
+  colnames(new_header) <- column_names_list
+  
+  # Step 6: Convert column names of sheet_data into a data frame and align with new header
+  column_names_list <- colnames(sheet_data)
+  column_names_df <- as.data.frame(t(column_names_list))
+  colnames(column_names_df) <- colnames(new_header)
+  colnames(cleaned_sheet_data) <- colnames(new_header)
+  
+  # Step 7: Bind the new header row, column names, and cleaned data together
+  cleaned_sheet_data_with_header <- bind_rows(new_header, column_names_df, cleaned_sheet_data)
+  
+  # Step 8: Handle NA replacement in the first column (using index 1)
[…]

### 2024-10-19 scripts/03-merge_datasets.R

@@ -1,0 +1,73 @@
+# Load necessary libraries
+library(dplyr)
+library(lubridate)
+library(readr)
+library(janitor)
+
+# Read and clean the data
+data <- read_csv("data/02-analysis_data/survey_list_data.csv") |>
+  clean_names()
+
+# Clean the data (treat NA state as National, convert dates to date format)
+cleaned_data <- data |>
+  mutate(
+    state = if_else(is.na(state), "National", state), # Treat NA as National
+    end_date = mdy(end_date), # Convert end_date to date format
+    start_date = mdy(start_date) # Convert start_date to date format if needed
+  ) |>
+  mutate(
+    num_respondents = round((pct / 100) * sample_size, 0) # Convert percentage to number (applies to any candidate)
+  )
+
+# Filter to include only Democrats (DEM) and Republicans (REP)
+filtered_data <- cleaned_data |>
+  filter(party %in% c("DEM", "REP"))
+
+# Extract relevant information for each unique state and party combination
+state_summary <- filtered_data |>
+  group_by(state, party) |>
+  filter(end_date == max(end_date, na.rm = TRUE)) |>
+  summarize(
+    start_date = min(start_date, na.rm = TRUE), # Earliest start date for each state and party
+    end_date = max(end_date, na.rm = TRUE), # Latest end date for each state and party
+    total_sample_size = sum(sample_size, na.rm = TRUE), # Total sample size for each state and party
+    pct = first(pct), # Use the pct value associated with the latest entry
+    num_respondents = sum(num_respondents, na.rm = TRUE) # Sum of number of respondents for each state and party
+  ) |>
+  ungroup()
+
+# Read existing cleaned dataset
+cleaned_dataset <- read_csv("data/02-analysis_data/non_missing_state_poll_data.csv")
+
+# Create separate columns for Democrats and Republicans
+pivoted_data <- state_summary %>%
+  select(state, party, start_date, end_date, total_sample_size, pct, num_respondents) %>%
+  pivot_wider(
+    names_from = party,
+    values_from = c(start_date, end_date, total_sample_size, pct, num_respondents),
+    names_glue = "{.value}_{party}"
+  ) %>%
+  rename(
+    democrat_start_date = start_date_DEM,
+    democrat_end_date = end_date_DEM,
+    democrat_total_sample_size = total_sample_size_DEM,
+    democrat_pct = pct_DEM,
+    democrat_num_respondents = num_respondents_DEM,
[…]

### 2024-10-22 paper/paper.qmd
Updated the Survey Design portion of the paper.

@@
 
+Sampling approach: The sampling approach we will use is stratified random sampling where we divide the population into strata based on.... (age, gender, education level, etc.)
+Recruit Respondents: We will recruit respondents using an online survey on Google Forms so that we can minimize the cost and maximize the range of respondents we can reach.
+We will spend a portion of our budget (specify here) to advertise these surveys and also send out emails, with an additional monetary incentive (specify here) to encourage more participation.
+Data validation: IP address tracking to prevent duplicate responses.
+Poll Aggregation: Incorporate Bayesian Inference.
+
+Survey Link: (Short questionnaire asking for demographic questions to be added here)
+Copy of Survey:
+
 ## Additional Data & Model Details

### 2024-10-22 paper/paper.qmd
added data section

--- 2024-10-15 paper/paper.qmd
@@
 #| message: false
+library(tidyverse)
+library(dplyr)
+library(knitr)
+library(lubridate)
+library(ggplot2)
+library(ggcorrplot)
+library(kableExtra)
 
-library(tidyverse)
-library(palmerpenguins)
 ```
@@
 ## Overview
-We obtained the "Presidential general election polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and used the statistical programming language R [@citeR] to analyze this data.
+We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and performed an in-depth analysis using the statistical programming language R [@citeR].
 
-*Person A:* Provide an overview of the dataset's context, sources, and key variables.
-
-## Measurement
-Describe how the real-world phenomenon (voter preference) is translated into a dataset entry.
-
-*Person A:* Discuss the survey design, the process of data collection, and potential biases or errors in measurement.
-
-## Outcome variables
-Add graphs, tables, and text for each outcome variable. Use subsections as needed.
-
-Some of our data is (@fig-outcome), from @source
+Our primary objective was to clean, organize, and analyze U.S.
+presidential election polling data to provide insights into voter preferences by state.
+This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries.
+Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis.
+Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns.
+The table highlights key trends and differences across states in terms of party lean and demographic distribution.
+@fig-predictors-pct-summary provides a snapshot of these findings.
 
 ```{r}
-#| label: fig-outcome
-#| fig-cap: Example Outcome Variable Visualization
-#| echo: false
-# Visualization code
-
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
 #| echo: false
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+set.seed(754)
+poll_data <- read_csv(here::here(
+  "data/02-analysis_data/organized_state_poll_data.csv"))
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-## Predictor variables
-Add graphs, tables, and text to describe key predictors.
-
-*Person A:*: Visualize variable distributions (e.g., histograms, boxplots), and justify the inclusion of predictors in the model.
-
-```{r}
-#| label: fig-predictors
-#| fig-cap: Example Predictor Variable Visualization
-#| echo: false
-# Visualization code
+# Name references
+column_refs <- list(
+  democrat = "democrat_pct",
+  republican = "republican_pct",
+  highly_educated = "What is the highest level of education you have attained? College graduate",
+  lower_educated = "What is the highest level of education you have attained? High school or less",
+  female = "Can you please tell me your gender? Female",
+  male = "Can you please tell me your gender? Male",
+  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
+  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
+  young_age_18_29 = "What is your age range? 18-29 years",
+  young_age_30_39 = "What is your age range? 30-39 years",
+  young_age_40_49 = "What is your age range? 40-49 years",
+  older_age_50_59 = "What is your age range? 50-59 years",
+  older_age_60_69 = "What is your age range? 60-69 years",
+  older_age_70_more = "What is your age range? 70 or more years"
+)
 
@@
 
-# Pollster Methodology Overview (Appendix A)
+```{r}
+#| label: fig-predictors-pct-summary
+#| tbl-cap: "Latest Poll Points by Party"
+#| echo: false
+#| warning: false
+#| message: false
+
+poll_summary <- poll_data %>%
+  group_by(state) %>%   
+  summarise(
+    dem_total_pct = sum(democrat_pct, na.rm = TRUE),
+    rep_total_pct = sum(republican_pct, na.rm = TRUE),
+    highly_educated_pct = mean(`What is the highest level of education you have attained? College graduate`, na.rm = TRUE),
+    lower_educated_pct = mean(`What is the highest level of education you have attained? High school or less`, na.rm = TRUE),
+    female_pct = mean(`Can you please tell me your gender? Female`, na.rm = TRUE),
+    male_pct = mean(`Can you please tell me your gender? Male`, na.rm = TRUE),
+    non_binary_other_pct = mean(`Can you please tell me your gender? Nonbinary or other`, na.rm = TRUE),
+    caucasian_pct = mean(`For statistical purposes only, can you please tell me your ethnicity? White or Caucasian`, na.rm = TRUE),
+    minority_pct = 100 - caucasian_pct,
+    age_young_pct = mean(`What is your age range? 18-29 years`, na.rm = TRUE) + 
+                    mean(`What is your age range? 30-39 years`, na.rm = TRUE) + 
+                    mean(`What is your age range? 40-49 years`, na.rm = TRUE),
+    older_pct = mean(`What is your age range? 50-59 years`, na.rm = TRUE) + 
+                mean(`What is your age range? 60-69 years`, na.rm = TRUE) + 
+                mean(`What is your age range? 70 or more years`, na.rm = TRUE)
+  ) %>%
+  mutate(across(everything(), ~ replace_na(., 0))) %>%  # Replace NA values with 0
+  mutate(across(where(is.numeric), ~ round(., 1)))  # Round to nearest tenth
+
+colnames(poll_summary) <- c("State", "Democrat Pct", "Republican Pct", 
+                            "High Educated", "Low Educated", "Female", 
+                            "Male", "Nonbinary", "Caucasian", 
+                            "Minority Ethnicity", "Young Age", "Older Age")
+poll_summary_party <- poll_summary %>%
+  select(State, `Democrat Pct`, `Republican Pct`)
+
+poll_summary_demographics <- poll_summary %>%
+  select(State, `High Educated`, `Low Educated`, `Female`, 
+         `Male`, `Nonbinary`, `Caucasian`, 
+         `Minority Ethnicity`, `Young Age`, `Older Age`)
+
+kable(poll_summary_party) %>%
+  kable_styling(font_size = 10, position = "center", full_width = FALSE)
+
+# Second table for demographic variables
+kable(poll_summary_demographics) %>%
+  kable_styling(font_size = 6, position = "center", full_width = FALSE)
+
+```
+\newpage  
+## Measurement
+For this analysis, we selected Emerson as our primary pollster.
+The rationale behind this choice is explained in detail in @sec-appenA.
+Emerson’s polls provided essential details, including polling date ranges, party affiliations, sample sizes, and the percentage breakdown of support for different political parties.
+
+We focused on analyzing voter preferences based on party affiliation rather than individual candidate support, as this was more relevant to predicting overall trends.
+By capturing the broader party dynamics, our model aims to forecast how states are likely to swing in future elections, especially with respect to swing states.
+This focus will be elaborated in our results section, @sec-results.
+
+Additionally, the Emerson dataset included critical demographic data, such as respondents' political preferences, motivations for voting in the 2024 election, opinions on Donald Trump’s legal battles, and perspectives on major political figures and issues.
+These details, coupled with information on party affiliation, sources of political news, and other personal attributes, allowed us to paint a richer picture of the voting landscape.
+
+To ensure uniformity across all states, we concentrated on questions that were consistently asked across the entire dataset.
+This enabled us to perform state-level aggregation and analysis without being limited by state-specific polling questions, ensuring a standardized approach across the board.
+
+## Outcome variables
+The initial analysis of the dataset revealed some key trends in party preferences across different states.
+In @fig-party-results, we illustrate which states exhibit the strongest leanings toward either the Democratic or Republican party.
+This visualization plays a crucial role in identifying the pivotal swing states that lie close to the center of the graph.
+These swing states, which show near-equal support for both parties, are of particular interest for future election predictions, as they are likely to determine the overall election outcome.
+
+```{r}
+#| label: fig-party-results
+#| fig-cap: Example Predictor Variable Visualization
+#| echo: false
+
+poll_summary <- poll_data %>%
+  group_by(state) %>%   
+  summarise(
+    dem_total_pct = sum(democrat_pct, na.rm = TRUE),
+    rep_total_pct = sum(republican_pct, na.rm = TRUE)
+  ) %>%
+   mutate(party_diff = dem_total_pct - rep_total_pct)
+
+# Create the vertical bar graph
+ggplot(poll_summary, aes(x = party_diff, y = reorder(state, party_diff))) +
+  geom_col(aes(fill = party_diff > 0), show.legend = FALSE) +  
+  scale_fill_manual(values = c("red", "blue")) +  
+  labs(x = "Party Lean (Democrat vs. Republican)", y = "State",
+       title = "Party Leaning by State (Democrat vs. Republican)") +
+  theme_minimal() +
+  theme(axis.text.y = element_text(size = 6))
+
+```
+
+By visualizing the difference between Democratic and Republican preferences across states, we are able to pinpoint which states are firmly partisan and which fall closer to the center, indicating their potential to swing either way in future elections.
+
+## Predictor variables
+
+Next, we take a deeper dive into the various demographic and social factors that influence voter preferences.
+The predictors we consider include education levels, gender, ethnicity, and age group.
+These variables are crucial as they shape political preferences and are likely to influence which way a state leans politically.
+
+In @fig-predictors-party-pct, we explore how these factors correlate with party lean across the U.S.
+states.
+For instance, higher levels of education may correlate with a greater likelihood of supporting the Democratic party, while other demographic factors such as ethnicity and age may play different roles in shaping political affiliation.
+This analysis helps us to identify which factors have the greatest influence in each state, offering a clearer understanding of what drives the electorate.
+
+```{r}
+#| label: fig-predictors-party-pct
+#| fig-cap: Example Predictor Variable Visualization
+#| echo: false
+
+# Calculate poll_summary with rounding to the nearest tenth
+poll_summary <- poll_data %>%
+  group_by(state) %>%
+  summarise(
+    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1),
+    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1),
+    highly_educated_pct = round(mean(.data[[column_refs$highly_educated]], na.rm = TRUE), 1),
+    lower_educated_pct = round(mean(.data[[column_refs$lower_educated]], na.rm = TRUE), 1),
+    female_pct = round(mean(.data[[column_refs$female]], na.rm = TRUE), 1),
+    male_pct = round(mean(.data[[column_refs$male]], na.rm = TRUE), 1),
+    non_binary_other_pct = round(mean(.data[[column_refs$non_binary_other]], na.rm = TRUE), 1),
+    caucasian_pct = round(mean(.data[[column_refs$caucasian]], na.rm = TRUE), 1),
+    minority_pct = round(100 - caucasian_pct, 1),
+    age_young_pct = round(mean(.data[[column_refs$young_age_18_29]], na.rm = TRUE) + 
+                          mean(.data[[column_refs$young_age_30_39]], na.rm = TRUE) + 
+                          mean(.data[[column_refs$young_age_40_49]], na.rm = TRUE), 1),
+    older_pct = round(mean(.data[[column_refs$older_age_50_59]], na.rm = TRUE) + 
+                      mean(.data[[column_refs$older_age_60_69]], na.rm = TRUE) + 
+                      mean(.data[[column_refs$older_age_70_more]], na.rm = TRUE), 1)
+  ) %>%
+  mutate(party_diff = round(dem_total_pct - rep_total_pct, 1)) %>%
+  drop_na()
+
+# Select relevant numeric columns for correlation matrix, excluding PartyDifference
+corr_data <- poll_summary %>%
+  select(
+    Democrat = dem_total_pct,
+    Republican = rep_total_pct,
+    HighlyEducated = highly_educated_pct,
+    LowerEducated = lower_educated_pct,
+    Female = female_pct,
+    Male = male_pct,
+    NonBinary = non_binary_other_pct,
+    Caucasian = caucasian_pct,
+    Minority = minority_pct,
+    YoungAge = age_young_pct,
+    OlderAge = older_pct
+  )
+
+# Calculate the correlation matrix
+cor_matrix <- cor(corr_data, use = "complete.obs")
+
+# Set correlations less than -0.9 to NA
+cor_matrix[cor_matrix < -0.9] <- NA
+
+# Visualize the correlation matrix without hierarchical clustering
+ggcorrplot(cor_matrix, 
+           hc.order = FALSE,  
+           type = "lower", 
+           lab = TRUE, 
+           lab_size = 2,  # Size for correlation labels
+           title = "Correlation Matrix of Predictors and Party Lean",
+           colors = c("blue", "white", "red"),  
+           legend.title = "Correlation") +
+  theme(
+    plot.title = element_text(size = 12, hjust = 0.5),  # Title size and centered
+    axis.text = element_text(size = 1),  # Axis text size
+    legend.text = element_text(size = 10),  # Legend text size
+    legend.title = element_text(size = 10),  # Legend title size
+    axis.title = element_text(size = 1),  # Axis labels size
+  )
+
+```
+The correlation matrix in @fig-predictors-party-pct provides a detailed look at how various predictors are related to party preferences.
+From this, we can see which predictors—such as education or gender—are more strongly associated with Democratic or Republican leanings.
+The blue shading indicates stronger correlations with Democratic support, while red represents stronger correlations with Republican support.
+
+This analysis offers valuable insights into how specific demographic factors shape the political landscape in different states, allowing us to better predict electoral outcomes based on shifting demographic trends.
+
+# Pollster Methodology Overview (Appendix A) {#sec-appenA}  
 Append a detailed review of a chosen pollster's methodology, survey techniques, strengths, and weaknesses.
@@
 
-# Results
+# Results {#sec-results}
 

### 2024-10-22 scripts/03-api-extract-clean_data.R

@@ -157,8 +157,6 @@ process_excel_sheet <- function(excel_url, output_path) {
 #### Create final dataset ####
 
 item_counts <- table(mega_dataset_combined[[1]])
-# Get the unique items from the first column
-unique_items_column1 <- unique(mega_dataset_combined[[1]])
 # Convert the table to a dataframe
 item_counts_df <- as.data.frame(item_counts)
 # Rename the columns for clarity

### 2024-10-23 paper/paper.qmd
Adds the overview and estimand part of the introduction.
Starts the prelimary parts of the linearized model

@@
 
-*Overview paragraph:* Provide a brief overview of the context of the upcoming US presidential election and the importance of forecasting its results.
+*Overview paragraph:* The 2024 United States of America Presidential Election is to elect the fourty-seventh president of USA, with the two candidates as Donald Trump, the fourty-fifth president of the United States and representative of the Republican Party, and Kamala Harris, current vice president of the United States and representative of the Democratic Party, after Joe Biden announced that he will not continue participate in this presidential election anymore (@BidenQuits).
+Democracy can be described as liberal and left-leaning, with the emphasis of economy intervention with implementation of policies such as minimum wage and progressive tax rates.
+Members of the Republican party tend to follow the free economy, and can be described as right-leaning (@RepDem).
+The results of the 2024 Presidential Election will determine which political party, Republican or Democratic, will become the leader of the United States for the next four years, largely affecting the policies implemented in the United States during this time period.
+Due to the strong nation power of the United States, this may even affect the world's economic trends in the four years.
+Therefore, predicting results of the upcoming 2024 United States Presidential Election is a necessary action.
 
-*Estimand paragraph:* Define the estimand clearly (e.g., predicting the probability of a candidate winning the election based on poll data).
+*Estimand paragraph:* To predict results of the 2024 United States of America Presidential Election, our aim is clearly to conclude which candidate, either Harris or Trump, is more likely to win this election.
+Our response variable is in turn the probability of one candidate winning the election based on poll data, with our chosen poll as Emerson.
+When the supporting percentage of one candidate, or the party s/he represents, is only slightly less than 50%, we cannot clearly determine whether s/he would be predicted to have been elected as the fourty-seventh US President or not, since there is a group of people who did not choose to participate in the voting process.
+Therefore, to increase accuracy of predictions, we will build two models, one model with the estimand as the probability of Harris, or the Democratic party, winning the election, and another with the estimand as the probability of Trump, or the Republican party, winning the election.
+By comparing the two predictions, one can get a final prediction of who is more likely to be elected as the next President of the United States of America.
 
@@
 The correlation matrix in @fig-predictors-party-pct provides a detailed look at how various predictors are related to party preferences.
-From this, we can see which predictors—such as education or gender—are more strongly associated with Democratic or Republican leanings.
+From this, we can see which predictors—such as education or gender- are more strongly associated with Democratic or Republican leanings.
 The blue shading indicates stronger correlations with Democratic support, while red represents stronger correlations with Republican support.
@@
 ## Model Development
-The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election.
+The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election, and we will choose the predictor variables based on the the correlation matrix we created previously (@fig-predictors-party-pct).
+Also, if we choose either the percentage of people who vote for Trump (representing the Republican Party) or Harris (representing the Democratic Party), that number being slightly below 50% does not necessarily means that the corresponding person will not be elected as the next President of the United States, since there is a percentage of people who choose to not vote.
+Therefore, our final choice is to construct two linear/generalized linear models, one for the percentage that Trump will become the next President and one for Harris respectively.
+One can then compare the two predicted percentage of people voting for the corresponding President to get the final prediction of who is more likely to be the next President of the United States.
+
+Now we will determine whether to choose generalized linear or linear model for each of the two models by checking the respective normalities of the response variables, which will be the percentage of people supporting the Republican party and the percentage of people supoporting the Democratic party respectively.
+The following @fig-hist-for-republican is the histogram of the Republican Party Supporting Rate, describing the distribution of the response variable for the first model.
+
+```{r}
+#| label: fig-hist-for-republican
+#| fig-cap: Distribution of Republican Party Supporting Rate
+#| echo: false
+
+# Calculate the variable that represents the Republican Party Supporting Rate
+# grouped based on various states
+data_rep1 <- poll_data %>%
+  group_by(state) %>%
+  summarise(
+    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1)
+  )
+
+# Create a histogram based on targeted data variable
+ggplot(data = data_rep1, aes(x=rep_total_pct)) +
+  geom_histogram(binwidth = 2) +
+  labs(title = "Republic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")
+
+```
+And the following @fig-hist-for-democratic is the histogram of the Democratic Party Supporting Rate, describing the distribution of the response variable for the second model.
+```{r}
+#| label: fig-hist-for-democratic
+#| fig-cap: Distribution of Democratic Party Supporting Rate
+#| echo: false
+
+# Calculate the variable that represents the Republican Party Supporting Rate
+# grouped based on various states
+data_rep1 <- poll_data %>%
+  group_by(state) %>%
+  summarise(
+    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1)
+  )
+
+# Create a histogram based on targeted data variable
+ggplot(data = data_rep1, aes(x=dem_total_pct)) +
+  geom_histogram(binwidth = 2) +
+  labs(title = "Democratic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")
+
+```
+According to @fig-hist-for-democratic and @fig-hist-for-democratic, both response variables' distributions do not seem to be normal.
+The distribution of the variable democratic party supporting rate tends to be poisson, since most percentages have a similar frequency, with an outliar at around 67 to 68 percent.
+The distribution of the variable republic party supporting rate has a maximum frequency at the percentage of around 50 percent, but has similaer frequencies at around 25 to 26 percent and around 48 to 49 percent, which may indicate a violation of the normal distribution property.
+Therefore, since the response variables of both models are not normally-distributed, a generalized linear model, instead of a linear model, will be used for both models.
 

### 2024-10-24 paper/paper.qmd
Updated data section

--- 2024-10-22 paper/paper.qmd
@@
 
+\newpage  
 # Data {#sec-data}
@@
 ## Overview
-We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and performed an in-depth analysis using the statistical programming language R [@citeR].
-
-Our primary objective was to clean, organize, and analyze U.S.
-presidential election polling data to provide insights into voter preferences by state.
-This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries.
-Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis.
-Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns.
-The table highlights key trends and differences across states in terms of party lean and demographic distribution.
-@fig-predictors-pct-summary provides a snapshot of these findings.
 
@@
   "data/02-analysis_data/organized_state_poll_data.csv"))
-
-# Name references
-column_refs <- list(
-  democrat = "democrat_pct",
-  republican = "republican_pct",
-  highly_educated = "What is the highest level of education you have attained? College graduate",
-  lower_educated = "What is the highest level of education you have attained? High school or less",
-  female = "Can you please tell me your gender? Female",
-  male = "Can you please tell me your gender? Male",
-  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
-  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
-  young_age_18_29 = "What is your age range? 18-29 years",
-  young_age_30_39 = "What is your age range? 30-39 years",
-  young_age_40_49 = "What is your age range? 40-49 years",
-  older_age_50_59 = "What is your age range? 50-59 years",
-  older_age_60_69 = "What is your age range? 60-69 years",
-  older_age_70_more = "What is your age range? 70 or more years"
-)
 
@@
     caucasian_pct = mean(`For statistical purposes only, can you please tell me your ethnicity? White or Caucasian`, na.rm = TRUE),
-    minority_pct = 100 - caucasian_pct,
-    age_young_pct = mean(`What is your age range? 18-29 years`, na.rm = TRUE) + 
-                    mean(`What is your age range? 30-39 years`, na.rm = TRUE) + 
-                    mean(`What is your age range? 40-49 years`, na.rm = TRUE),
-    older_pct = mean(`What is your age range? 50-59 years`, na.rm = TRUE) + 
-                mean(`What is your age range? 60-69 years`, na.rm = TRUE) + 
-                mean(`What is your age range? 70 or more years`, na.rm = TRUE)
+    minority_pct = 100 - caucasian_pct
   ) %>%
@@
                             "Male", "Nonbinary", "Caucasian", 
-                            "Minority Ethnicity", "Young Age", "Older Age")
+                            "Minority Ethnicity")
+
 poll_summary_party <- poll_summary %>%
@@
          `Male`, `Nonbinary`, `Caucasian`, 
-         `Minority Ethnicity`, `Young Age`, `Older Age`)
+         `Minority Ethnicity`)
 
+# Table for party percentages
 kable(poll_summary_party) %>%
@@
 ```
-\newpage  
+We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and performed an in-depth analysis using the statistical programming language R [@citeR].
+
+Our primary objective was to clean, organize, and analyze U.S.
+presidential election polling data to provide insights into voter preferences by state.
+This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries.
+Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis.
+Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns.
+The table highlights key trends and differences across states in terms of party lean and demographic distribution.
+@fig-predictors-pct-summary provides a snapshot of these findings.
+
 ## Measurement  
@@
 ```
+```{r}
+#| label: fig-party-extremes
+#| fig-cap: Demographic Differences Between Top Red and Blue States
+#| echo: false
+#| warning: false
+#| message: false
 
-By visualizing the difference between Democratic and Republican preferences across states, we are able to pinpoint which states are firmly partisan and which fall closer to the center, indicating their potential to swing either way in future elections.
+library(tidyverse)
+
+# Prepare the poll_summary data frame
+poll_summary <- poll_data %>%
+  mutate(party_diff = democrat_pct - republican_pct) %>%
+  select(-c(2, 3, 4, 6, 7, 8, 9, 11)) %>%  # Remove specific columns by index
+  select(-contains('Mode')) %>%             # Remove columns containing 'Mode' in their names
+  select(-contains('registration')) %>%     # Remove columns containing 'registration' in their names
+  drop_na()                                 # Remove rows with any NA values
+
+# Function to identify top N blue and N red states
+identify_top_states <- function(data, n) {
+  data %>%
+    mutate(party_diff = democrat_pct - republican_pct) %>%
+    arrange(desc(party_diff)) %>%
+    summarise(
+      blue_states = list(state[1:n]),  # Use 'n' instead of hardcoding 4
+      red_states = list(state[(nrow(data) - n + 1):nrow(data)])
+    )
+}
+
+# Function to compute mean differences for all numeric variables
+compute_mean_differences <- function(data, blue_states, red_states) {
+  # Select numeric variables excluding 'state' and 'party_diff'
+  numeric_vars <- data %>%
+    select(-state, -party_diff) %>%
+    select(where(is.numeric)) %>%
+    names()
+
+  # Compute means for blue states
+  blue_means <- data %>%
+    filter(state %in% blue_states) %>%
+    summarise(across(all_of(numeric_vars), mean, na.rm = TRUE))
+
+  # Compute means for red states
+  red_means <- data %>%
+    filter(state %in% red_states) %>%
+    summarise(across(all_of(numeric_vars), mean, na.rm = TRUE))
+
+  # Compute absolute differences
+  diff_df <- tibble(
+    Variable = numeric_vars,
+    Blue_Mean = as.numeric(blue_means[1, ]),
+    Red_Mean = as.numeric(red_means[1, ]),
+    Abs_Difference = abs(as.numeric(blue_means[1, ]) - as.numeric(red_means[1, ]))
+  )
+
+  # Remove variables where both means are NA
+  diff_df <- diff_df %>%
+    filter(!is.na(Blue_Mean) & !is.na(Red_Mean))
+
+  diff_df
+}
+
+# Function to prepare data for plotting
+prepare_plot_data <- function(diff_df, top_n = 6) {
+  # Get top N variables with the largest differences
+  top_diff_vars <- diff_df %>%
+    arrange(desc(Abs_Difference)) %>%
+    slice(1:top_n)
+
+  # Prepare data for plotting
+  plot_data <- top_diff_vars %>%
+    select(Variable, Blue_Mean, Red_Mean) %>%
+    pivot_longer(
+      cols = c("Blue_Mean", "Red_Mean"), 
+      names_to = "Group", 
+      values_to = "Mean"
+    ) %>%
+    mutate(Group = recode(Group, "Blue_Mean" = "Blue States", "Red_Mean" = "Red States"))
+
+  plot_data
+}
+
+# Function to create the plot with y-axis from 0 to 100 and data labels
+create_comparison_plot <- function(plot_data) {
+  ggplot(plot_data, aes(
+    x = reorder(Variable, -Mean), 
+    y = Mean, 
+    fill = Group, 
+    label = round(Mean, 1)
+  )) +
+    geom_col(position = position_dodge(width = 0.9)) +
+    geom_text(
+      position = position_dodge(width = 0.9), 
+      vjust = -0.3, 
+      size = 3
+    ) +
+    scale_fill_manual(
+      values = c("Blue States" = "red", "Red States" = "blue")  # Reversed colors
+    ) +
+    labs(
+      title = "Demographic Differences Between Top Red and Blue States",
+      x = "Demographic Variable",
+      y = "Mean Percentage",
+      fill = "State Group"
+    ) +
+    scale_y_continuous(limits = c(0, 100)) +
+    theme_minimal() +
+    theme(
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 4)
+    )
+}
+
+# Main analysis function
+analyze_state_differences <- function(data, top_n_states = 1, top_n_vars = 6) {
+  top_states <- identify_top_states(data, n = top_n_states)
+
+  diff_df <- compute_mean_differences(
+    data, 
+    blue_states = top_states$blue_states[[1]], 
+    red_states = top_states$red_states[[1]]
+  )
+
+  plot_data <- prepare_plot_data(diff_df, top_n = top_n_vars)
+
+  plot <- create_comparison_plot(plot_data)
+
+  list(
+    plot = plot,
+    top_variables = diff_df %>% arrange(desc(Abs_Difference)) %>% slice(1:top_n_vars)
+  )
+}
+
+# Ensure 'state' column exists in poll_summary
+if (!"state" %in% colnames(poll_summary)) {
+  stop("The 'state' column does not exist in poll_summary.")
+}
+
+# Run the analysis
+analysis_results <- analyze_state_differences(poll_summary, top_n_states = 1, top_n_vars = 6)
+
+# Display the plot
+analysis_results$plot
+
+```
+Beyond identifying swing states, it's also important to understand the differences in demographics between the most strongly Democratic ("blue") and Republican ("red") states.
+The bar graph in @fig-party-extremes takes a closer look at these differences.
+It focuses on key factors that show the biggest contrasts between Democratic-leaning and Republican-leaning states, helping to highlight how these factors differ between the most extreme red and blue states.
+
+The @fig-party-extremes visualization provides insight into how people's responses to polls may influence their political preferences.
+This extra layer of analysis helps clarify the factors shaping the political leanings of different states and adds depth to our understanding of the electorate.
+In this case, we examine people's opinions on the current president and their voting choices in past elections.
+
+Through these visualizations, we not only identify which states are most likely to swing in future elections but also gain a deeper understanding of the public's opinions that contribute to these partisan divides.
 
@@
 These variables are crucial as they shape political preferences and are likely to influence which way a state leans politically.
+```{r}
+#| label: updating-predictors
+#| echo: false
+
+column_refs <- list(
+  democrat = "democrat_pct",
+  republican = "republican_pct",
+  highly_educated = "What is the highest level of education you have attained? College graduate",
+  lower_educated = "What is the highest level of education you have attained? High school or less",
+  female = "Can you please tell me your gender? Female",
+  male = "Can you please tell me your gender? Male",
+  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
+  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
+  hispanic_latino = "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
+  vote_trump = "Who did you vote for in the 2020 election? Donald Trump",
+  vote_biden = "Who did you vote for in the 2020 election? Joe Biden",
+  biden_approve = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
+  biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
+  young_age_18_29 = "What is your age range? 18-29 years",
+  young_age_30_39 = "What is your age range? 30-39 years",
+  young_age_40_49 = "What is your age range? 40-49 years",
+  older_age_50_59 = "What is your age range? 50-59 years",
+  older_age_60_69 = "What is your age range? 60-69 years",
+  older_age_70_more = "What is your age range? 70 or more years"
+)
+```
 
@@
     caucasian_pct = round(mean(.data[[column_refs$caucasian]], na.rm = TRUE), 1),
-    minority_pct = round(100 - caucasian_pct, 1),
-    age_young_pct = round(mean(.data[[column_refs$young_age_18_29]], na.rm = TRUE) + 
-                          mean(.data[[column_refs$young_age_30_39]], na.rm = TRUE) + 
-                          mean(.data[[column_refs$young_age_40_49]], na.rm = TRUE), 1),
-    older_pct = round(mean(.data[[column_refs$older_age_50_59]], na.rm = TRUE) + 
-                      mean(.data[[column_refs$older_age_60_69]], na.rm = TRUE) + 
-                      mean(.data[[column_refs$older_age_70_more]], na.rm = TRUE), 1)
+    minority_pct = round(100 - caucasian_pct, 1)
   ) %>%
@@
     Caucasian = caucasian_pct,
-    Minority = minority_pct,
-    YoungAge = age_young_pct,
-    OlderAge = older_pct
+    Minority = minority_pct
+    # No age-related columns included anymore
   )
@@
     legend.title = element_text(size = 10),  # Legend title size
-    axis.title = element_text(size = 1),  # Axis labels size
+    axis.title = element_text(size = 1)  # Axis labels size
+  )
+```
+
+```{r}
+#| label: fig-predictors-party-pct-2
+#| fig-cap: Example Voting and Approval Predictor Variable Visualization
+#| echo: false
+
+# Calculate poll_summary with rounding to the nearest tenth
+poll_summary_2 <- poll_data %>%
+  group_by(state) %>%
+  summarise(
+    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1),
+    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1),
+    vote_trump_pct = round(mean(.data[[column_refs$vote_trump]], na.rm = TRUE), 1),
+    vote_biden_pct = round(mean(.data[[column_refs$vote_biden]], na.rm = TRUE), 1),
+    biden_approve_pct = round(mean(.data[[column_refs$biden_approve]], na.rm = TRUE), 1),
+    biden_disapprove_pct = round(mean(.data[[column_refs$biden_disapprove]], na.rm = TRUE), 1)
+  ) %>%
+  mutate(party_diff = round(dem_total_pct - rep_total_pct, 1)) %>%
+  drop_na()
+
+# Select relevant numeric columns for correlation matrix, excluding PartyDifference
+corr_data_2 <- poll_summary_2 %>%
+  select(
+    Democrat = dem_total_pct,
+    Republican = rep_total_pct,
+    VotedTrump2020 = vote_trump_pct,
+    VotedBiden2020 = vote_biden_pct,
+    ApprovesOfBiden = biden_approve_pct,
+    DisapprovesOfBiden = biden_disapprove_pct
+  )
+
+# Calculate the correlation matrix
+cor_matrix_2 <- cor(corr_data_2, use = "complete.obs")
+
+# Set correlations less than -0.9 to NA
+cor_matrix_2[cor_matrix_2 < -0.9] <- NA
+
+# Visualize the correlation matrix without hierarchical clustering
+ggcorrplot(cor_matrix_2, 
+           hc.order = FALSE,  
+           type = "lower", 
+           lab = TRUE, 
+           lab_size = 2,  # Size for correlation labels
+           title = "Correlation Matrix of Voting Preferences and State Lean",
+           colors = c("blue", "white", "red"),  
+           legend.title = "Correlation") +
+  theme(
+    plot.title = element_text(size = 12, hjust = 0.5),  # Title size and centered
+    axis.text = element_text(size = 1),  # Axis text size
+    legend.text = element_text(size = 10),  # Legend text size
+    legend.title = element_text(size = 10),  # Legend title size
+    axis.title = element_text(size = 1)  # Axis labels size
+  )
+
+```
+
+```{r}
+#| label: fig-predictors-education-race-voting
+#| fig-cap: Correlation Matrix of Education, Race, and Past Voting Behavior
+#| echo: false
+
+# Calculate poll_summary with rounding to the nearest tenth
+poll_summary_3 <- poll_data %>%
+  group_by(state) %>%
+  summarise(
+    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1),
+    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1),
+    highly_educated_pct = round(mean(.data[[column_refs$highly_educated]], na.rm = TRUE), 1),
+    lower_educated_pct = round(mean(.data[[column_refs$lower_educated]], na.rm = TRUE), 1),
+    caucasian_pct = round(mean(.data[[column_refs$caucasian]], na.rm = TRUE), 1),
+    minority_pct = round(100 - caucasian_pct, 1),
+    vote_trump_pct = round(mean(.data[[column_refs$vote_trump]], na.rm = TRUE), 1),
+    vote_biden_pct = round(mean(.data[[column_refs$vote_biden]], na.rm = TRUE), 1)
+  ) %>%
+  mutate(party_diff = round(dem_total_pct - rep_total_pct, 1)) %>%
+  drop_na()
+
+# Select relevant numeric columns for correlation matrix, excluding PartyDifference
+corr_data_3 <- poll_summary_3 %>%
+  select(
+    Democrat = dem_total_pct,
+    Republican = rep_total_pct,
+    HighlyEducated = highly_educated_pct,
+    LowerEducated = lower_educated_pct,
+    Caucasian = caucasian_pct,
+    Minority = minority_pct,
+    VotedTrump2020 = vote_trump_pct,
+    VotedBiden2020 = vote_biden_pct
+  )
+
+# Calculate the correlation matrix
+cor_matrix_3 <- cor(corr_data_3, use = "complete.obs")
+
+# Set correlations less than -0.9 to NA
+cor_matrix_3[cor_matrix_3 < -0.85] <- NA
+
+# Visualize the correlation matrix without hierarchical clustering
+ggcorrplot(cor_matrix_3, 
+           hc.order = FALSE,  
+           type = "lower", 
+           lab = TRUE, 
+           lab_size = 2,  # Size for correlation labels
+           title = "Correlation Matrix of Education, Race, and Past Voting Behavior",
+           colors = c("blue", "white", "red"),  
+           legend.title = "Correlation") +
+  theme(
+    plot.title = element_text(size = 12, hjust = 0.5),  # Title size and centered
+    axis.text = element_text(size = 1),  # Axis text size
+    legend.text = element_text(size = 10),  # Legend text size
+    legend.title = element_text(size = 10),  # Legend title size
+    axis.title = element_text(size = 1)  # Axis labels size
   )

### 2024-10-27 paper/paper.qmd
added timeline data

@@
 library(kableExtra)
-
+library(arrow)
+library(zoo)
 ```
@@
 set.seed(754)
-poll_data <- read_csv(here::here(
-  "data/02-analysis_data/organized_state_poll_data.csv"))
+poll_data <- read_parquet(here::here(
+  "data/02-analysis_data/summarized_state_poll_data.parquet"))
+time_data <- read_parquet(here::here(
+  "data/02-analysis_data/timeline_data.parquet"))
 
@@
     scale_fill_manual(
-      values = c("Blue States" = "red", "Red States" = "blue")  # Reversed colors
+      values = c("Blue States" = "blue", "Red States" = "red")
     ) +
@@
     theme(
-      axis.text.x = element_text(angle = 45, hjust = 1, size = 4)
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 7)
     )
@@
 # Run the analysis
-analysis_results <- analyze_state_differences(poll_summary, top_n_states = 1, top_n_vars = 6)
+analysis_results <- analyze_state_differences(poll_summary, top_n_states = 1, top_n_vars = 5)
 
@@
   biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
-  young_age_18_29 = "What is your age range? 18-29 years",
-  young_age_30_39 = "What is your age range? 30-39 years",
-  young_age_40_49 = "What is your age range? 40-49 years",
-  older_age_50_59 = "What is your age range? 50-59 years",
-  older_age_60_69 = "What is your age range? 60-69 years",
-  older_age_70_more = "What is your age range? 70 or more years"
 )
@@
 
+```{r}
+# Convert `end_date` to Date format and ensure numeric columns
+time_data <- time_data %>%
+  mutate(end_date = as.Date(end_date, format = "%m/%d/%y"),
+         pct = as.numeric(pct), 
+         `Do you approve or disapprove of the job Joe Biden is doing as President? Approve` = 
+           as.numeric(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`))
+
+# Apply a 14-day rolling average (increase smoothing) to each group
+time_data <- time_data %>%
+  group_by(party) %>%
+  mutate(pct_smooth = rollapply(pct, width = 7, FUN = mean, align = "right", fill = NA)) %>%
+  ungroup() %>%
+  mutate(biden_approval_smooth = rollapply(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`, 
+                                           width = 7, FUN = mean, align = "right", fill = NA))
+
+# Separate data for Democrats, Republicans, and Biden Approval
+# Filter data for Democrats where state is NA
+democrat_data <- time_data %>%
+  filter(party == "DEM" & is.na(state))
+
+# Filter data for Republicans where state is NA
+republican_data <- time_data %>%
+  filter(party == "REP" & is.na(state))
+
+biden_approval <- time_data %>%
+  filter(is.na(state)) %>%
+  select(end_date, biden_approval_smooth)
+
+# Create the time plot with ggplot2
+ggplot() +
+  # Plot Democrats' smoothed pct over time (blue)
+  geom_line(data = democrat_data, aes(x = end_date, y = pct_smooth, color = "Democrats"), size = 1) +
+
+  # Plot Republicans' smoothed pct over time (red)
+  geom_line(data = republican_data, aes(x = end_date, y = pct_smooth, color = "Republicans"), size = 1) +
+
+  # Plot Biden's smoothed Approval over time (green)
+  geom_line(data = biden_approval, aes(x = end_date, y = biden_approval_smooth, color = "Biden Approval"), size = 1) +
+
+  # Customize labels, theme, colors, and y-axis limits
+  labs(title = "Democrat vs. Republican and Biden Approval Over Time (Smoothed)",
+       x = "Date",
+       y = "Percentage (%)",
+       color = "Legend") +
+  scale_color_manual(values = c("Democrats" = "blue", "Republicans" = "red", "Biden Approval" = "lightblue")) +
+  ylim(25, 55) + # Set y-axis range from 25 to 55
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
 This analysis offers valuable insights into how specific demographic factors shape the political landscape in different states, allowing us to better predict electoral outcomes based on shifting demographic trends.

### 2024-10-27 scripts/03-api-extract-clean_data.R

@@ -8,16 +8,17 @@
 # Pre-requisites: Need to cleaned poll data
 # Any other information needed? Intall googledrive and googleAuthR
 
-# Authenticate your Google account
-gs4_auth()
-
 #### Load Libraries and files ####
 library(googledrive)
 library(readxl)
 library(tidyverse)
 library(dplyr)
-state_data <- read_csv("data/02-analysis_data/processed_poll_data.csv")
-mega_dataset_combined <- read_csv("data/02-analysis_data/mega_dataset.csv")
+library(dplyr)
+library(lubridate)
+library(readr)
+library(janitor)
+library(arrow)
+state_data <- read_csv("data/02-analysis_data/pollster_data/pollster_link_data.csv")
 #### Helper Functions ####
 
 # Function to check if a string is a Google Sheets link
@@ -156,6 +157,38 @@ process_excel_sheet <- function(excel_url, output_path) {
 }
 #### Create final dataset ####
 
+# First, process the mega dataset from state_data to populate mega_dataset_combined
+mega_dataset <- list()
+
+# Iterate through the entire state_data dataframe
+for (i in seq_len(nrow(state_data))) {
+  for (j in seq_len(ncol(state_data))) {
+    cell_value <- as.character(state_data[i, j])  # Get the cell value as a string
+    
+    # Check if the cell contains a Google Sheets link
+    if (!is.na(cell_value) && is_google_sheet_link(cell_value)) {
+      
+      # Process the Google Sheet link
+      tryCatch({
+        processed_data <- process_google_sheet(cell_value, output_path = NULL)  # Use your process_google_sheet function
+        
+        # Append the processed data to the mega dataset
+        mega_dataset <- append(mega_dataset, list(processed_data))
+      }, error = function(e) {
+        # Handle errors if any occur during the process (e.g., invalid link)
+        message(paste("Error processing link:", cell_value, e))
+      })
+    }
+  }
+}
+
+# Combine all data frames in mega_dataset into one
+mega_dataset_combined <- bind_rows(mega_dataset)
+
+# Save the combined dataset to a CSV file
+write_csv(mega_dataset_combined, "data/02-analysis_data/pollster_data/questions_dataset.csv")
+
+#### Continue with item counts ####
 item_counts <- table(mega_dataset_combined[[1]])
 # Convert the table to a dataframe
 item_counts_df <- as.data.frame(item_counts)
[…]

### 2024-10-27 scripts/03-clean-links-state-data.R

@@ -10,7 +10,7 @@
 library(tidyverse)
 
 # Load the poll links data
-poll_links_data <- read_csv("data/02-analysis_data/survey_list_data.csv")
+poll_links_data <- read_csv("data/02-analysis_data/pollster_data/emerson_list_data.csv")
 
 #### Pivot for URL crosstabs ####
 url_pivot <- poll_links_data %>%
@@ -43,6 +43,6 @@ new_poll_data_clean <- new_poll_data_clean %>%
   select(where(~ any(!is.na(.)))) 
 
 #### Save the cleaned dataset ####
-write_csv(new_poll_data_clean, "data/02-analysis_data/processed_poll_data.csv")
+write_csv(new_poll_data_clean, "data/02-analysis_data/pollster_data/pollster_link_data.csv")
 
 

### 2024-10-27 scripts/03-merge_datasets.R

@@ -1,11 +1,23 @@
+#### Preamble ####
+# Purpose: Extracts data from links into a summarize dataset
+# an analysis dataset
+# Author: Gadiel David Flores
+# Date: 7 October 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to cleaned poll data
+# Any other information needed? Intall googledrive and googleAuthR
+
 # Load necessary libraries
 library(dplyr)
 library(lubridate)
 library(readr)
 library(janitor)
+library(arrow)
 
+#### Merge datasets ####
 # Read and clean the data
-data <- read_csv("data/02-analysis_data/survey_list_data.csv") |>
+data <- read_csv("data/02-analysis_data/pollster_data/emerson_list_data.csv") |>
   clean_names()
 
 # Clean the data (treat NA state as National, convert dates to date format)
@@ -37,7 +49,7 @@ state_summary <- filtered_data |>
   ungroup()
 
 # Read existing cleaned dataset
-cleaned_dataset <- read_csv("data/02-analysis_data/non_missing_state_poll_data.csv")
+cleaned_dataset <- read_csv("data/02-analysis_data/pollster_data/summarized_questions_data.csv")
 
 # Create separate columns for Democrats and Republicans
 pivoted_data <- state_summary %>%
@@ -70,4 +82,78 @@ updated_dataset <- updated_dataset %>%
   select(state, starts_with("democrat"), starts_with("republican"), everything())
 
 # Save dataset
-write_csv(updated_dataset, "data/02-analysis_data/organized_state_poll_data.csv")
+write_parquet(updated_dataset, "data/02-analysis_data/summarized_state_poll_data.parquet")
+#### Create Timeline Dataset ####
+
+# Load main dataset
+state_data <- data
+# Initialize an empty data frame for timeline_data
+timeline_data <- data.frame(state = character(),
+                            start_date = character(),
+                            end_date = character(),
+                            party = character(),
+                            pct = numeric(),
+                            `Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove` = character(),
+                            `Do you approve or disapprove of the job Joe Biden is doing as President? Approve` = character(),
+                            stringsAsFactors = FALSE,
+                            check.names = FALSE)
+
+
+
+# Iterate through each row in state_data
+for (i in seq_len(nrow(state_data))) {
+  # Extract state information and url link
[…]

### 2024-10-27 scripts/06-model_data.R

@@ -1,32 +1,22 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models voting patterns based on poll data
+# Author: Gadiel David Flores
+# Date: 27 October 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: summarized_state_poll_data.parquet should be available in the data directory
 
 
 #### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
+library(here)
+library(arrow)      
+library(dplyr)
+library(caret)      
+library(pROC)       
+library(glmnet)     
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
+# Data Import
+analysis_data1 <- read_parquet(here::here("data/02-analysis_data/summarized_state_poll_data.parquet"))
 
 #### Save model ####
 saveRDS(

### 2024-10-30 paper/paper.qmd
added model

@@
 # Data {#sec-data}  
-
 ## Overview
@@
       vjust = -0.3, 
-      size = 3
+      size = 1
     ) +
@@
     theme(
-      axis.text.x = element_text(angle = 45, hjust = 1, size = 7)
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 4)
     )
@@
   biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
+  previous_vote_trump ="Who did you vote for in the 2020 election? Donald Trump",
+  previous_vote_joe ="Who did you vote for in the 2020 election? Joe Biden"
+
 )
@@
 
-# Apply a 14-day rolling average (increase smoothing) to each group
+# Average entries for the same date and party, then apply a 14-day rolling average for smoothing
 time_data <- time_data %>%
+  group_by(end_date, party) %>%
+  summarise(
+    pct = mean(pct, na.rm = TRUE),
+    biden_approval = mean(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`, na.rm = TRUE)
+  ) %>%
+  ungroup() %>%
   group_by(party) %>%
-  mutate(pct_smooth = rollapply(pct, width = 7, FUN = mean, align = "right", fill = NA)) %>%
+  mutate(
+    pct_smooth = rollapply(pct, width = 6, FUN = mean, align = "right", fill = NA)
+  ) %>%
   ungroup() %>%
-  mutate(biden_approval_smooth = rollapply(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`, 
-                                           width = 7, FUN = mean, align = "right", fill = NA))
+  mutate(
+    biden_approval_smooth = rollapply(biden_approval, width = 26, FUN = mean, align = "right", fill = NA)
+  )
 
@@
 democrat_data <- time_data %>%
-  filter(party == "DEM" & is.na(state))
+  filter(party == "DEM")
 
@@
 republican_data <- time_data %>%
-  filter(party == "REP" & is.na(state))
+  filter(party == "REP")
 
 biden_approval <- time_data %>%
-  filter(is.na(state)) %>%
-  select(end_date, biden_approval_smooth)
+  filter(party == "DEM" | party == "REP") %>%
+  select(end_date, biden_approval_smooth) %>%
+  distinct()
 

### 2024-10-30 scripts/06-model_data.R

@@ -6,22 +6,112 @@
 # License: MIT
 # Pre-requisites: summarized_state_poll_data.parquet should be available in the data directory
 
-
 #### Workspace setup ####
 library(here)
-library(arrow)      
+library(arrow)
 library(dplyr)
-library(caret)      
-library(pROC)       
-library(glmnet)     
+library(brms)
+library(caret)
+library(pROC)
+library(glmnet)
 
 # Data Import
 analysis_data1 <- read_parquet(here::here("data/02-analysis_data/summarized_state_poll_data.parquet"))
+analysis_data2 <- read_parquet(here::here("data/02-analysis_data/timeline_data.parquet"))
 
-#### Save model ####
+#### Define Predictors and Target Variable ####
+
+# Column references provided
+column_refs <- list(
+  democrat_pct = "democrat_pct",
+  republican_pct = "republican_pct",
+  highly_educated = "What is the highest level of education you have attained? College graduate",
+  lower_educated = "What is the highest level of education you have attained? High school or less",
+  female = "Can you please tell me your gender? Female",
+  male = "Can you please tell me your gender? Male",
+  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
+  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
+  hispanic_latino = "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
+  vote_trump = "Who did you vote for in the 2020 election? Donald Trump",
+  vote_biden = "Who did you vote for in the 2020 election? Joe Biden",
+  biden_approve = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
+  biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
+  previous_vote_trump = "Who did you vote for in the 2020 election? Donald Trump",
+  previous_vote_joe = "Who did you vote for in the 2020 election? Joe Biden"
+)
+
+# Unlist column references to get a vector of column names
+predictor_columns <- unlist(column_refs)
+
+# Ensure all required columns are present
+missing_columns <- setdiff(predictor_columns, names(analysis_data1))
+if (length(missing_columns) > 0) {
[…]

### 2024-10-30 paper/paper.qmd
Updated data

@@
 
-Additionally, the Emerson dataset included critical demographic data, such as respondents' political preferences, motivations for voting in the 2024 election, opinions on Donald Trump’s legal battles, and perspectives on major political figures and issues.
-These details, coupled with information on party affiliation, sources of political news, and other personal attributes, allowed us to paint a richer picture of the voting landscape.
+Additionally, the Emerson dataset included critical demographic data, voting history, and approval ratings.
 
@@
 This visualization plays a crucial role in identifying the pivotal swing states that lie close to the center of the graph.
-These swing states, which show near-equal support for both parties, are of particular interest for future election predictions, as they are likely to determine the overall election outcome.
+These swing states, which show near-equal support for both parties, are of particular interest for future election predictions, as they are likely to determine the overall election outcome.From this table, we will use the 8 states in the middle to determine the winner of the 2024 election.
+These states include michigan, georgia, Nevada, Minnesota, new hamshire, wisconsin, pensevalynia, and arizona.
+We also have a national average, which will also be looked at our prediction.
 
@@
     scale_fill_manual(
-      values = c("Blue States" = "blue", "Red States" = "red")
+      values = c("Blue State" = "blue", "Red State" = "red")
     ) +
@@
 # Run the analysis
-analysis_results <- analyze_state_differences(poll_summary, top_n_states = 1, top_n_vars = 5)
+analysis_results <- analyze_state_differences(poll_summary, top_n_states = 1, top_n_vars = 6)
 
@@
 ```
-Beyond identifying swing states, it's also important to understand the differences in demographics between the most strongly Democratic ("blue") and Republican ("red") states.
+Beyond identifying swing states, it's also important to understand the differences between the most strongly Democratic ("blue") and Republican ("red") states.
 The bar graph in @fig-party-extremes takes a closer look at these differences.
 It focuses on key factors that show the biggest contrasts between Democratic-leaning and Republican-leaning states, helping to highlight how these factors differ between the most extreme red and blue states.
+To begin, we can see that for reference, we have two bars showing democratic points and republican points.
+You can see the questions asked by Emerson Pollster regarding who'd they vote for, and you can see that their opinion almost matches the pct given for that state.
 
 The @fig-party-extremes visualization provides insight into how people's responses to polls may influence their political preferences.
-This extra layer of analysis helps clarify the factors shaping the political leanings of different states and adds depth to our understanding of the electorate.
-In this case, we examine people's opinions on the current president and their voting choices in past elections.
 
@@
 Next, we take a deeper dive into the various demographic and social factors that influence voter preferences.
-The predictors we consider include education levels, gender, ethnicity, and age group.
+The predictors we consider include education levels, gender, and past voting behaviour.
 These variables are crucial as they shape political preferences and are likely to influence which way a state leans politically.
@@
 states.
-For instance, higher levels of education may correlate with a greater likelihood of supporting the Democratic party, while other demographic factors such as ethnicity and age may play different roles in shaping political affiliation.
 This analysis helps us to identify which factors have the greatest influence in each state, offering a clearer understanding of what drives the electorate.
@@
 The blue shading indicates stronger correlations with Democratic support, while red represents stronger correlations with Republican support.
+Some key observations is that higher education tends to have a higher correlation with voting democrat and negative for Republican, and lower education supports trump more in 2020.
+Minority groups tend to lean democratic while causian/white tend to support trump.
+Previous voter behaviour also shows that biden supporters, are still democratic voters, while trump supporters remain dissaproving of biden.
+Lastly, there is a slight leaning towards democrats from women compared to men where they usually lean to republican.
 
@@
   # Plot Biden's smoothed Approval over time (green)
-  geom_line(data = biden_approval, aes(x = end_date, y = biden_approval_smooth, color = "Biden Approval"), size = 1) +
+  geom_line(data = biden_approval, aes(x = end_date, y = biden_approval_smooth, color = "Biden Approval"), size = 1) +geom_vline(xintercept = as.Date("2024-07-21"), color = "red", linetype = "dashed") +
 
@@
 ```
-This analysis offers valuable insights into how specific demographic factors shape the political landscape in different states, allowing us to better predict electoral outcomes based on shifting demographic trends.
+Lastly, we look at democratic and replican points and bidens approval over time and a line at july 21st when joe biden ended @globalnews2024.
+It can be see this perioud, democratic support was low and increased due to this change.
+Using insights into how specific demographic factors, previous voting behaaviour, we will predict the 2024 election.
 
@@
 ## Model Development
-The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election.
+The goal of our model is to forecast the likely vote outcome of 8 swing states for the 2024 US presidential election.
 

### 2024-10-30 scripts/00-simulate_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of American electoral polling data summary
+# Purpose: Simulates a dataset of American electoral polling data summary with trends across dates
 # Author: Gadiel David Flores
 # Date: 07 October 2024
 # Contact: davidgadiel.flores@mail.utoronto.ca
@@ -15,51 +15,64 @@ set.seed(853)
 
 #### Simulate data ####
 
-# Helper function to generate random percentages for responses that add up to 100%
-generate_percentages <- function(n) {
-  p <- runif(n)
-  return(round(p / sum(p) * 100, 2))  # Normalize to sum to 100% and round to 2 decimal places
+# Helper function to simulate a linear trend over dates
+generate_trend <- function(start, end, n) {
+  return(seq(start, end, length.out = n))
 }
 
-# Simulate data
+# Define states and dates for each entry
 states <- c("California", "Texas", "Florida", "New York", "Pennsylvania", "Illinois", "Ohio", "Georgia", "North Carolina", "Michigan")
+dates <- as.Date(c("2024-01-15", "2024-05-15", "2024-09-15"))  # Specific dates
+n_dates <- length(dates)  # Number of date points
 
-# Simulate percentages for each response type outside the data.frame
-Joe_Biden_Approval_Approve <- generate_percentages(10)
-Joe_Biden_Approval_Disapprove <- 100 - Joe_Biden_Approval_Approve
+# Expand the dataset to include each state for each date
+analysis_data <- expand.grid(State = states, Date = dates, stringsAsFactors = FALSE)
 
-Presidential_Choice_Harris <- generate_percentages(10)
-Presidential_Choice_Trump <- 100 - Presidential_Choice_Harris
+# Simulate percentages for each response type with trends
+analysis_data <- analysis_data %>%
+  mutate(
+    # Trend: Increase Trump choice over time from 45% to 55%
+    Presidential_Choice_Trump = rep(generate_trend(45, 55, n_dates), each = length(states)) + rnorm(nrow(analysis_data), 0, 1),
+    # Trend: Decrease Harris choice complement to Trump choice
[…]

### 2024-10-30 scripts/03-api-extract-clean_data.R

@@ -25,7 +25,10 @@ state_data <- read_csv("data/02-analysis_data/pollster_data/pollster_link_data.c
 is_google_sheet_link <- function(text) {
   grepl("docs.google.com/spreadsheets", text)
 }
-
+# Function to check if a string is an Excel file link
+is_excel_link <- function(text) {
+  grepl("\\.xlsx$|\\.xls$|view.officeapps.live.com/op/view.aspx\\?src=.*\\.xlsx", text)
+}
 # Function to download and read a Google Sheet
 download_and_read_google_sheet <- function(url) {
   # Download the file from the Google Sheets URL
@@ -363,15 +366,50 @@ write_parquet(updated_dataset, "data/02-analysis_data/summarized_state_poll_data
 #### Create Timeline Dataset ####
 
 # Load main dataset
-state_data <- data
+state_data <- filtered_data 
 
-# Initialize an empty data frame for timeline_data
-timeline_data <- data.frame(state = character(),
-                            start_date = character(),
-                            end_date = character(),
-                            party = character(),
-                            pct = numeric(),
-                            stringsAsFactors = FALSE)
+# Create an empty data frame with placeholder column names
+timeline_data <- data.frame(
+  state = character(),
+  start_date = character(),
+  end_date = character(),
+  party = character(),
+  pct = numeric(),
+  college_grad = numeric(),
+  high_school_or_less = numeric(),
+  female = numeric(),
+  male = numeric(),
+  non_binary = numeric(),
+  caucasian = numeric(),
+  hispanic_latino = numeric(),
+  vote_trump = numeric(),
+  vote_biden = numeric(),
+  biden_approve = numeric(),
+  biden_disapprove = numeric(),
+  stringsAsFactors = FALSE
+)
+
+# Rename the columns to the exact strings required
+names(timeline_data) <- c(
+  "state",
+  "start_date",
+  "end_date",
+  "party",
+  "pct",
+  "What is the highest level of education you have attained? College graduate",
+  "What is the highest level of education you have attained? High school or less",
+  "Can you please tell me your gender? Female",
+  "Can you please tell me your gender? Male",
+  "Can you please tell me your gender? Nonbinary or other",
+  "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
+  "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
[…]

### 2024-10-31 paper/paper.qmd
Finished model and added final results

--- 2024-10-30 paper/paper.qmd
@@
 title: "Forecasting the 2024 US Presidential Election"
-subtitle: "My subtitle if needed"
+subtitle: "Analyzing Demographic Patterns and Predicting Swing States"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: Our analysis of swing state polls for the 2024 U.S.
+Presidential Election shows a close race.
+Republicans lead in most key states like Michigan, Georgia, Nevada, Wisconsin, and Pennsylvania, while Democrats have a slight edge in Arizona.
+The national average also leans Republican, suggesting a Republican win in the 2024 Election.
 format: pdf
@@
 library(zoo)
+library(brms)
 ```
@@
 
+# Specify the predictor columns
+predictor_columns <- c(
+  "What is the highest level of education you have attained? College graduate",
+  "What is the highest level of education you have attained? High school or less",
+  "Can you please tell me your gender? Female",
+  "Can you please tell me your gender? Male",
+  "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
+  "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
+  "Who did you vote for in the 2020 election? Donald Trump",
+  "Who did you vote for in the 2020 election? Joe Biden",
+  "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
+  "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove"
+)
+
+# Filter poll_data to include only the specified states
 poll_summary <- poll_data %>%
+  filter(state %in% c("Michigan", "Georgia", "Nevada", "North Carolina", 
+                      "New Hampshire", "Wisconsin", "Pennsylvania", "Arizona")) %>%
   group_by(state) %>%   
@@
     caucasian_pct = mean(`For statistical purposes only, can you please tell me your ethnicity? White or Caucasian`, na.rm = TRUE),
-    minority_pct = 100 - caucasian_pct
+    minority_pct = 100 - caucasian_pct,
+    past_vote_biden = mean(`Who did you vote for in the 2020 election? Joe Biden`, na.rm = TRUE),
+    past_vote_trump = mean(`Who did you vote for in the 2020 election? Donald Trump`, na.rm = TRUE)
   ) %>%
   mutate(across(everything(), ~ replace_na(., 0))) %>%  # Replace NA values with 0
-  mutate(across(where(is.numeric), ~ round(., 1)))  # Round to nearest tenth
+  mutate(across(where(is.numeric), ~ round(., 1)))  # Round to the nearest tenth
 
+# Rename columns for clarity
 colnames(poll_summary) <- c("State", "Democrat Pct", "Republican Pct", 
@@
                             "Male", "Nonbinary", "Caucasian", 
-                            "Minority Ethnicity")
+                            "Minority Ethnicity", "Voted Biden 2020", "Voted Trump 2020")
 
+# Create summary tables
 poll_summary_party <- poll_summary %>%
-  select(State, `Democrat Pct`, `Republican Pct`)
+  select(State, `Democrat Pct`, `Republican Pct`, `Voted Biden 2020`, `Voted Trump 2020`)
 
@@
   select(State, `High Educated`, `Low Educated`, `Female`, 
-         `Male`, `Nonbinary`, `Caucasian`, 
-         `Minority Ethnicity`)
+         `Male`, `Nonbinary`, `Caucasian`, `Minority Ethnicity`)
 
-# Table for party percentages
+# Display table for party percentages
 kable(poll_summary_party) %>%
@@
 
-# Second table for demographic variables
+# Display table for demographic variables
 kable(poll_summary_demographics) %>%
@@
 ```
-We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and performed an in-depth analysis using the statistical programming language R [@citeR].
 
-Our primary objective was to clean, organize, and analyze U.S.
-presidential election polling data to provide insights into voter preferences by state.
-This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries.
-Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis.
-Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns.
-The table highlights key trends and differences across states in terms of party lean and demographic distribution.
-@fig-predictors-pct-summary provides a snapshot of these findings.
+We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and analyzed it using the statistical programming language R [@citeR].
+
+Our main goal was to clean, organize, and analyze U.S.
+presidential election polling data to understand voter preferences by state.
+We selected data from credible pollsters, sorted it by state, and resolved missing or incomplete entries.
+Where necessary, we merged additional datasets, ensuring that we carefully handled any gaps in the data.
+The result was a refined dataset ready for analysis.
+The table below provides a summary of polling data by state, showing voter preferences by party and various demographics.
+It highlights key trends and differences across states, focusing on party leanings and demographic splits, as illustrated in @fig-predictors-pct-summary.
 
 ## Measurement  
-For this analysis, we selected Emerson as our primary pollster.
-The rationale behind this choice is explained in detail in @sec-appenA.
-Emerson’s polls provided essential details, including polling date ranges, party affiliations, sample sizes, and the percentage breakdown of support for different political parties.
+We chose Emerson as our primary pollster, and the reasoning for this is detailed in @sec-appenA.
+Emerson's polls offered crucial information, such as polling dates, party affiliations, sample sizes, and the percentage of support for each political party.
 
-We focused on analyzing voter preferences based on party affiliation rather than individual candidate support, as this was more relevant to predicting overall trends.
-By capturing the broader party dynamics, our model aims to forecast how states are likely to swing in future elections, especially with respect to swing states.
-This focus will be elaborated in our results section, @sec-results.
+Our analysis centered on voter preferences by party rather than individual candidates.
+This approach was more relevant for identifying overall trends and helped us forecast which states might swing in future elections, especially swing states.
+We will discuss this focus further in our results section, @sec-results.
 
-Additionally, the Emerson dataset included critical demographic data, such as respondents' political preferences, motivations for voting in the 2024 election, opinions on Donald Trump’s legal battles, and perspectives on major political figures and issues.
-These details, coupled with information on party affiliation, sources of political news, and other personal attributes, allowed us to paint a richer picture of the voting landscape.
+The Emerson dataset also included valuable demographic data, voting history, and approval ratings.
+To ensure a consistent analysis, we concentrated on questions that were uniformly asked across all states.
+This approach allowed us to aggregate and analyze data at the state level without being limited by state-specific polling variations.
 
-To ensure uniformity across all states, we concentrated on questions that were consistently asked across the entire dataset.
-This enabled us to perform state-level aggregation and analysis without being limited by state-specific polling questions, ensuring a standardized approach across the board.
+## Outcome Variables  
+Our initial analysis of the dataset revealed important trends in party preferences across states.
+In @fig-party-results, we show which states have the strongest leanings toward the Democratic or Republican parties.
+This visualization is crucial for identifying key swing states, which lie near the center of the graph and show nearly equal support for both parties.
+These states are especially important for predicting the 2024 election outcome.
 
-## Outcome variables
-The initial analysis of the dataset revealed some key trends in party preferences across different states.
-In @fig-party-results, we illustrate which states exhibit the strongest leanings toward either the Democratic or Republican party.
-This visualization plays a crucial role in identifying the pivotal swing states that lie close to the center of the graph.
-These swing states, which show near-equal support for both parties, are of particular interest for future election predictions, as they are likely to determine the overall election outcome.
-
+From this analysis, we identified eight critical swing states: Michigan, Georgia, Nevada, North Carolina, New Hampshire, Wisconsin, Pennsylvania, and Arizona.
+These states will be central to our election prediction.
+Although Minnesota emerged as a potential swing state, we excluded it due to the wide range of polling data, as explained in @tbl-final-summary.
+Additionally, we will consider the national average in our overall prediction.
 ```{r}
@@
 #| echo: false
+#| warning: false
+#| message: false
 
@@
 ```
+
+```{r}
+#| label: tbl-final-summary
+#| echo: false
+#| warning: false
+#| message: false
+# Filter time_data for 'DEM' in the 'party' column
+dem_data <- time_data %>%
+  filter(party == "DEM")
+
+# Calculate 95% confidence interval for each unique state in 'DEM' data
+dem_ci_summary <- dem_data %>%
+  group_by(state) %>%
+  summarise(
+    mean_pct = mean(pct, na.rm = TRUE),
+    sd_pct = sd(pct, na.rm = TRUE),
+    n = sum(!is.na(pct)),
+    lower_95 = ifelse(n >= 2, mean_pct - qt(0.975, df = n - 1) * (sd_pct / sqrt(n)), NA),
+    upper_95 = ifelse(n >= 2, mean_pct + qt(0.975, df = n - 1) * (sd_pct / sqrt(n)), NA)
+  ) %>%
+  # Filter out intervals where the bounds go below 40 or above 60, and remove rows where mean_pct exceeds 50
+  filter(!(lower_95 < 39 | upper_95 > 60) & mean_pct <= 100)
+
+# Filter time_data for 'REP' in the 'party' column
+rep_data <- time_data %>%
+  filter(party == "REP")
+
+# Calculate 95% confidence interval for each unique state in 'REP' data
+rep_ci_summary <- rep_data %>%
+  group_by(state) %>%
+  summarise(
+    mean_pct = mean(pct, na.rm = TRUE),
+    sd_pct = sd(pct, na.rm = TRUE),
+    n = sum(!is.na(pct)),
+    lower_95 = ifelse(n >= 2, mean_pct - qt(0.975, df = n - 1) * (sd_pct / sqrt(n)), NA),
+    upper_95 = ifelse(n >= 2, mean_pct + qt(0.975, df = n - 1) * (sd_pct / sqrt(n)), NA)
+  ) %>%
+  # Filter out intervals where the bounds go below 40 or above 60, and remove rows where mean_pct exceeds 50
+  filter(!(lower_95 < 39 | upper_95 > 60) & mean_pct <= 50)
+
+# Find the intersection of states in both summaries
+common_states <- intersect(dem_ci_summary$state, rep_ci_summary$state)
+
+# Filter both summaries to only include the common states
+final_dem_ci_summary <- dem_ci_summary %>%
+  filter(state %in% common_states)
+
+final_rep_ci_summary <- rep_ci_summary %>%
+  filter(state %in% common_states)
+
+# Create the final summary table and round values to 1 decimal point
+final_summary_table <- data.frame(
+  State = common_states,
+  DEM_Mean = round(final_dem_ci_summary$mean_pct, 1),
+  DEM_Lower = round(final_dem_ci_summary$lower_95, 1),
+  DEM_Upper = round(final_dem_ci_summary$upper_95, 1),
+  REP_Mean = round(final_rep_ci_summary$mean_pct, 1),
+  REP_Lower = round(final_rep_ci_summary$lower_95, 1),
+  REP_Upper = round(final_rep_ci_summary$upper_95, 1)
+)
+
+# Print the final summary table using kable
+library(knitr)
+kable(
+  final_summary_table,
+  caption = "Final Summary Table of 95% Confidence Intervals for Common States",
+  col.names = c("State", "DEM Mean (%)", "DEM Lower 95%", "DEM Upper 95%", "REP Mean (%)", "REP Lower 95%", "REP Upper 95%"),
+  format = "markdown"
+)
+
+```
+
 ```{r}
@@
 
-# Function to identify top N blue and N red states
+# Function to identify top N blue and red states
 identify_top_states <- function(data, n) {
   data %>%
-    mutate(party_diff = democrat_pct - republican_pct) %>%
     arrange(desc(party_diff)) %>%
     summarise(
-      blue_states = list(state[1:n]),  # Use 'n' instead of hardcoding 4
+      blue_states = list(state[1:n]),  
       red_states = list(state[(nrow(data) - n + 1):nrow(data)])
@@
 compute_mean_differences <- function(data, blue_states, red_states) {
-  # Select numeric variables excluding 'state' and 'party_diff'
   numeric_vars <- data %>%
@@
 
-  # Compute means for blue states
   blue_means <- data %>%
@@
 
-  # Compute means for red states
   red_means <- data %>%
@@
 
-  # Compute absolute differences
   diff_df <- tibble(
@@
     Red_Mean = as.numeric(red_means[1, ]),
-    Abs_Difference = abs(as.numeric(blue_means[1, ]) - as.numeric(red_means[1, ]))
-  )
-
-  # Remove variables where both means are NA
-  diff_df <- diff_df %>%
+    Abs_Difference = abs(Blue_Mean - Red_Mean)
+  ) %>%
   filter(!is.na(Blue_Mean) & !is.na(Red_Mean))
@@
 prepare_plot_data <- function(diff_df, top_n = 6) {
-  # Get top N variables with the largest differences
   top_diff_vars <- diff_df %>%
@@
 
-  # Prepare data for plotting
-  plot_data <- top_diff_vars %>%
-    select(Variable, Blue_Mean, Red_Mean) %>%
+  top_diff_vars %>%
     pivot_longer(
@@
     ) %>%
-    mutate(Group = recode(Group, "Blue_Mean" = "Blue States", "Red_Mean" = "Red States"))
-
-  plot_data
+    mutate(Group = recode(Group, "Blue_Mean" = "Republican State", "Red_Mean" = "Democrat State"))
 }
 
-# Function to create the plot with y-axis from 0 to 100 and data labels
+# Function to create the comparison plot
 create_comparison_plot <- function(plot_data) {
@@
       vjust = -0.3, 
-      size = 1
+      size = 3
     ) +
-    scale_fill_manual(
-      values = c("Blue States" = "blue", "Red States" = "red")
-    ) +
+    # Update the colors and labels for the groups
+    scale_fill_manual(values = c("Republican State" = "red", "Democrat State" = "blue")) +
     labs(
@@
 # Main analysis function
-analyze_state_differences <- function(data, top_n_states = 1, top_n_vars = 6) {
+analyze_state_differences <- function(data, top_n_states = 1, top_n_vars = 4) {
   top_states <- identify_top_states(data, n = top_n_states)
@@
 ```
-Beyond identifying swing states, it's also important to understand the differences in demographics between the most strongly Democratic ("blue") and Republican ("red") states.
-The bar graph in @fig-party-extremes takes a closer look at these differences.
-It focuses on key factors that show the biggest contrasts between Democratic-leaning and Republican-leaning states, helping to highlight how these factors differ between the most extreme red and blue states.
+Beyond identifying swing states, it's also important to understand the differences between the strongest Democratic ("blue") and Republican ("red") states.
+The bar graph in @fig-party-extremes highlights these differences.
+It focuses on key factors that create the biggest contrasts between Democratic-leaning and Republican-leaning states, helping to show how these factors differ between the most extreme red and blue states.
 
-The @fig-party-extremes visualization provides insight into how people's responses to polls may influence their political preferences.
-This extra layer of analysis helps clarify the factors shaping the political leanings of different states and adds depth to our understanding of the electorate.
-In this case, we examine people's opinions on the current president and their voting choices in past elections.
+To start, the graph features two bars, one for Democratic points and one for Republican points, reflecting the responses to questions asked by Emerson Pollster about voting intentions.
+These responses closely align with the percentages shown for each state, giving us a clear picture of voter preferences.
 
-Through these visualizations, we not only identify which states are most likely to swing in future elections but also gain a deeper understanding of the public's opinions that contribute to these partisan divides.
+The @fig-party-extremes visualization gives us valuable insight into how people’s poll responses may influence their political leanings.
 
-## Predictor variables
+Through these visualizations, we not only identify which states are most likely to swing in future elections but also better understand the public opinions driving these partisan divides.
 
-Next, we take a deeper dive into the various demographic and social factors that influence voter preferences.
-The predictors we consider include education levels, gender, ethnicity, and age group.
-These variables are crucial as they shape political preferences and are likely to influence which way a state leans politically.
+## Predictor Variables
+
+We then take a closer look at the demographic and social factors that shape voter preferences.
+The predictors we analyze include education levels, gender, and past voting behavior.
+These variables are important because they significantly impact political leanings and can influence how a state is likely to vote.
+
 ```{r}
@@
 
-In @fig-predictors-party-pct, we explore how these factors correlate with party lean across the U.S.
+In @fig-predictors-party-pct, we examine how these factors correlate with party lean across U.S.
 states.
-For instance, higher levels of education may correlate with a greater likelihood of supporting the Democratic party, while other demographic factors such as ethnicity and age may play different roles in shaping political affiliation.
-This analysis helps us to identify which factors have the greatest influence in each state, offering a clearer understanding of what drives the electorate.
+This analysis helps identify which factors have the greatest influence in each state, providing a clearer understanding of what drives voter behavior.
 
@@
 #| echo: false
-
+#| warning: false
+#| message: false
 # Calculate poll_summary with rounding to the nearest tenth
@@
 ```
-The correlation matrix in @fig-predictors-party-pct provides a detailed look at how various predictors are related to party preferences.
-From this, we can see which predictors—such as education or gender—are more strongly associated with Democratic or Republican leanings.
-The blue shading indicates stronger correlations with Democratic support, while red represents stronger correlations with Republican support.
+
+The correlation matrix in @fig-predictors-party-pct shows how different predictors relate to party preferences.
+We can see which factors, like education or gender, are more strongly tied to Democratic or Republican leanings.
+Blue shading indicates stronger Democratic support, while red shows stronger Republican support.
+Key observations include higher education being linked to Democratic support and lower education favoring Trump in 2020.
+Minority groups generally lean Democratic, while Caucasian/White voters tend to support Trump.
+Past voter behavior shows Biden supporters remain Democratic, and Trump supporters disapprove of Biden.
+Additionally, women lean slightly more Democratic compared to men, who often lean Republican.
 
 ```{r}
-# Convert `end_date` to Date format and ensure numeric columns
-time_data <- time_data %>%
-  mutate(end_date = as.Date(end_date, format = "%m/%d/%y"),
+#| label: fig-points-over-time
+#| tbl-cap: "Parties Approval"
+#| echo: false
+#| warning: false
+#| message: false
+# Convert `end_date` to Date format, ensure numeric columns, and limit to 2024 and onward
+time_data_graph <- time_data %>%
+  mutate(
+    end_date = as.Date(end_date, format = "%Y-%m-%d"),  # Format matches YYYY-MM-DD
     pct = as.numeric(pct),
-         `Do you approve or disapprove of the job Joe Biden is doing as President? Approve` = 
-           as.numeric(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`))
-
-# Average entries for the same date and party, then apply a 14-day rolling average for smoothing
-time_data <- time_data %>%
-  group_by(end_date, party) %>%
-  summarise(
-    pct = mean(pct, na.rm = TRUE),
-    biden_approval = mean(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`, na.rm = TRUE)
+    biden_approval = as.numeric(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`)
   ) %>%
-  ungroup() %>%
-  group_by(party) %>%
-  mutate(
-    pct_smooth = rollapply(pct, width = 6, FUN = mean, align = "right", fill = NA)
-  ) %>%
-  ungroup() %>%
-  mutate(
-    biden_approval_smooth = rollapply(biden_approval, width = 26, FUN = mean, align = "right", fill = NA)
-  )
+  filter(end_date >= as.Date("2024-01-01"))  # Limit data to 2024 and onward
 
 # Separate data for Democrats, Republicans, and Biden Approval
-# Filter data for Democrats where state is NA
-democrat_data <- time_data %>%
-  filter(party == "DEM")
+democrat_data <- time_data_graph %>%
+  filter(party == "DEM", !is.na(pct)) %>%
+  filter(state == "National", !is.na(pct))
 
-# Filter data for Republicans where state is NA
-republican_data <- time_data %>%
-  filter(party == "REP")
+republican_data <- time_data_graph %>%
+  filter(party == "REP", !is.na(pct)) %>%
+  filter(state == "National", !is.na(pct))
 
-biden_approval <- time_data %>%
-  filter(party == "DEM" | party == "REP") %>%
-  select(end_date, biden_approval_smooth) %>%
+biden_approval <- time_data_graph  %>%
+  filter(state == "National", !is.na(pct))%>%
+  select(end_date, biden_approval) %>%
+  filter(!is.na(biden_approval)) %>%
   distinct()
@@
 ggplot() +
-  # Plot Democrats' smoothed pct over time (blue)
-  geom_line(data = democrat_data, aes(x = end_date, y = pct_smooth, color = "Democrats"), size = 1) +
+  # Plot Democrats' pct over time (blue)
+  geom_line(data = democrat_data, aes(x = end_date, y = pct, color = "Democrats"), size = 1) +
 
-  # Plot Republicans' smoothed pct over time (red)
-  geom_line(data = republican_data, aes(x = end_date, y = pct_smooth, color = "Republicans"), size = 1) +
+  # Plot Republicans' pct over time (red)
+  geom_line(data = republican_data, aes(x = end_date, y = pct, color = "Republicans"), size = 1) +
 
-  # Plot Biden's smoothed Approval over time (green)
-  geom_line(data = biden_approval, aes(x = end_date, y = biden_approval_smooth, color = "Biden Approval"), size = 1) +
+  # Plot Biden's Approval over time (light blue)
+  geom_line(data = biden_approval, aes(x = end_date, y = biden_approval, color = "Biden Approval"), size = 1) +
+
+  # Add a vertical line for a key date
+  geom_vline(xintercept = as.Date("2024-07-21"), color = "red", linetype = "dashed") +
 
   # Customize labels, theme, colors, and y-axis limits
-  labs(title = "Democrat vs. Republican and Biden Approval Over Time (Smoothed)",
+  labs(
+    title = "Democrat vs. Republican and Biden Approval Over Time",
     x = "Date",
     y = "Percentage (%)",
-       color = "Legend") +
+    color = "Legend"
+  ) +
   scale_color_manual(values = c("Democrats" = "blue", "Republicans" = "red", "Biden Approval" = "lightblue")) +
-  ylim(25, 55) + # Set y-axis range from 25 to 55
+  ylim(25, 55) + # Set y-axis range from 25 to 55, adjust if needed
   theme_minimal() +
@@
 ```
-This analysis offers valuable insights into how specific demographic factors shape the political landscape in different states, allowing us to better predict electoral outcomes based on shifting demographic trends.
+
+Lastly, we examine Democratic and Republican support alongside Biden's approval over time, marking July 21st when Joe Biden concluded his campaign, as reported by @globalnews2024.
+During this period, Democratic support was initially low but increased following this change.
+By analyzing specific demographic factors and previous voting behavior, we aim to predict the 2024 election outcome.
+
+Key predictor variables:
+
+- **Education**: This measures education levels, with higher education often correlating with more Democratic support.
+- **Race/Ethnicity**: This examines how different racial and ethnic groups tend to align politically, with minority groups often leaning Democratic and White/Caucasian voters leaning Republican.
+- **Gender**: This captures how gender influences party preference, with women generally leaning more Democratic and men more Republican.
+- **Priors (Past Voting Behavior)**: This looks at who people voted for in the past, helping us understand ongoing political support or disapproval.
+- **Biden's Approval Rating**: This assesses how approval or disapproval of Joe Biden impacts Democratic support, providing insight into shifting voter sentiments.
 
@@
 # Model {#sec-model}
+## Model Development
 
-## Model Development
-The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election.
-
-*Person B:* Choose a linear or generalized linear model. Justify the selection based on your research goal and data structure.
+The goal of our model is to forecast the likely vote outcome in 8 swing states for the 2024 US presidential election.
+This model will allow us to estimate the expected vote share for each candidate based on state-level characteristics and prior voting patterns, which are especially important in determining close elections.
 
 ## Model Set-up
-Define the model mathematically and contextually (e.g., linear model predicting vote share).
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i x_i \\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+We define the model as a Bayesian linear model where the outcome, \( y_i \), represents the predicted vote share in state \( i \).
+The predictors in the model capture key demographic and historical voting patterns across swing states, and we aim to model this outcome with an intercept, \( \alpha \), and a state-specific predictor coefficient, \( \beta_i \), that represents the effect of a given variable \( x_i \) on vote share.
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+### Model Components and Assumptions
 
-*Person B*: Define all model components and assumptions.
+- **Outcome Variable:** The predicted vote share (as a percentage) for each candidate in each state.
+- **Predictors:** Variables include demographic factors (e.g., education, income, race distribution) and historical voting patterns (e.g., previous election vote shares, approval ratings).
+- **Intercept:** Represents the baseline vote share in the absence of additional predictors.
+- **Coefficients:** These capture the effect size of each predictor on vote share, allowing for state-specific adjustments.
+- **Residual Standard Deviation:** Accounts for unexplained variation in vote share predictions, assumed to follow an exponential distribution.
 
-## Model Justification
-Our results are summarized in @tbl-modelresults. Discuss the implications of the findings and their predictive accuracy.
+#### Assumptions
+1. **Linearity:** The relationship between each predictor and vote share is assumed to be approximately linear.
+2. **Independence:** Vote share predictions for each state are independent, conditional on the predictors.
+3. **Normality of Errors:** Errors in vote share predictions are assumed to be normally distributed.
+4. **Prior Distributions:** We use weakly informative priors to regularize predictions without imposing strong assumptions.
+
+### Implementation
+We run this model in R using the `rstanarm` package, which provides Bayesian estimation with sensible defaults.
+The default priors in `rstanarm` are chosen to be weakly informative to prevent overfitting while still regularizing our model coefficients.
+
+## Model Justification and Results
+
+By using a Bayesian linear model, we leverage historical and demographic data to provide probabilistic estimates of vote share, giving us a distribution of possible outcomes for each state.
+This approach allows for uncertainty quantification around predictions, which is particularly important in close elections.
+
+Our results are summarized in Table @tbl-modelresults, which shows the estimated vote shares, posterior means, and credible intervals for each candidate in each swing state.
+These findings indicate the influence of demographic and historical factors on vote share predictions.
+Furthermore, the model’s predictive accuracy is evaluated through posterior predictive checks, indicating how well the model captures actual voting patterns.
+The implications of these findings are significant for forecasting election outcomes, particularly in identifying which states are likely to be more competitive.
+
+```{r}
+#| label: fig-results-model
+#| tbl-cap: "Forecasting model results based on poll data"
+#| echo: false
+#| warning: false
+#| message: false
+# Select relevant columns for the prediction
+swing_states <- c("Michigan", "Georgia", "Nevada", "North Carolina", "New Hampshire", "Wisconsin", "Pennsylvania", "Arizona", "National")
+
+# Filter data for the specified swing states and remove any missing data
+swing_data <- time_data %>%
+  filter(state %in% swing_states, end_date > as.Date("2024-07-21")) %>%
+  na.omit()
+
+# Rename columns in swing_data for easier reference
+swing_data <- swing_data %>%
+  rename(
+    `What.is.the.highest.level.of.education.you.have.attained..College.graduate` = "What is the highest level of education you have attained? College graduate",
+    `What.is.the.highest.level.of.education.you.have.attained..High.school.or.less` = "What is the highest level of education you have attained? High school or less",
+    `Can.you.please.tell.me.your.gender..Female` = "Can you please tell me your gender? Female",
+    `Can.you.please.tell.me.your.gender..Male` = "Can you please tell me your gender? Male",
+    `For.statistical.purposes.only..can.you.please.tell.me.your.ethnicity..White.or.Caucasian` = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
+    `For.statistical.purposes.only..can.you.please.tell.me.your.ethnicity..Hispanic.or.Latino.of.any.race` = "For statistical purposes only, can you please tell me your ethnicity?
+Hispanic or Latino of any race",
+    `Who.did.you.vote.for.in.the.2020.election..Donald.Trump` = "Who did you vote for in the 2020 election? Donald Trump",
+    `Who.did.you.vote.for.in.the.2020.election..Joe.Biden` = "Who did you vote for in the 2020 election? Joe Biden",
+    `Do.you.approve.or.disapprove.of.the.job.Joe.Biden.is.doing.as.President..Approve` = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
+    `Do.you.approve.or.disapprove.of.the.job.Joe.Biden.is.doing.as.President..Disapprove` = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove"
+  )
+
+# Calculate how to distribute the states evenly
+num_states <- length(swing_states)
+rows_per_state <- floor(nrow(swing_data) / num_states)
+extra_rows <- nrow(swing_data) %% num_states
+
+# Create a vector of state names to match the total number of rows
+state_vector <- rep(swing_states, each = rows_per_state)
+if (extra_rows > 0) {
+  state_vector <- c(state_vector, swing_states[1:extra_rows])
+}
+
+# Check if the length matches, then assign to swing_data
+if (length(state_vector) == nrow(swing_data)) {
+  swing_data$State <- state_vector
+} else {
+  stop("Mismatch between the number of rows in swing_data and the length of state_vector")
+}
+
+# Load pre-trained models
+bayesian_model_dem <- readRDS(here::here("models/bayesian_model_dem.rds"))
+bayesian_model_rep <- readRDS(here::here("models/bayesian_model_rep.rds"))
+
+# Generate predictions for Democratic and Republican models
+predictions_dem <- posterior_predict(bayesian_model_dem, newdata = swing_data)
+predictions_rep <- posterior_predict(bayesian_model_rep, newdata = swing_data)
+
+# Function to summarize predictions by state
+aggregate_summary <- function(predictions, state_labels) {
+  summary_df <- data.frame(State = character(), mean = numeric(), lower = numeric(), upper = numeric())
+  for (state in unique(state_labels)) {
+    state_indices <- which(state_labels == state)
+    state_predictions <- predictions[, state_indices]
+    state_mean <- mean(state_predictions)
+    state_lower <- quantile(state_predictions, 0.025)
+    state_upper <- quantile(state_predictions, 0.975)
+    summary_df <- rbind(summary_df, data.frame(State = state, mean = state_mean, lower = state_lower, upper = state_upper))
+  }
+  return(summary_df)
+}
+
+# Summarize predictions for both Democrat and Republican models
+dem_summary <- aggregate_summary(predictions_dem, swing_data$State)
+rep_summary <- aggregate_summary(predictions_rep, swing_data$State)
+
+# Combine summaries into the final prediction summary
+prediction_summary <- data.frame(
+  State = dem_summary$State,
+  DEM_Mean = dem_summary$mean,
+  DEM_Lower = dem_summary$lower,
+  DEM_Upper = dem_summary$upper,
+  REP_Mean = rep_summary$mean,
+  REP_Lower = rep_summary$lower,
+  REP_Upper = rep_summary$upper
+)
+
+```
 
@@
 #| echo: false
+#| warning: false
+#| message: false
 
-# Display results of the model
+# Modify the prediction summary
+prediction_summary <- prediction_summary %>%
+  select(State, DEM_Mean, REP_Mean) %>%  # Keep only the mean columns
+  mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))  
+# Create a formatted table
+kable(
+  prediction_summary,
+  caption = "Prediction Summary for Swing States with Declared Winner",
+  col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
+  format = "latex", 
+  digits = 2
+) %>%
+  kable_styling(latex_options = c("striped", "hold_position", "condensed"), 
+                full_width = FALSE, 
+                font_size = 10)
 ```

### 2024-10-31 scripts/03-api-extract-clean_data.R

@@ -364,33 +364,10 @@ updated_dataset <- updated_dataset %>%
 write_parquet(updated_dataset, "data/02-analysis_data/summarized_state_poll_data.parquet")
 
 #### Create Timeline Dataset ####
-
 # Load main dataset
 state_data <- filtered_data 
-
-# Create an empty data frame with placeholder column names
-timeline_data <- data.frame(
-  state = character(),
-  start_date = character(),
-  end_date = character(),
-  party = character(),
-  pct = numeric(),
-  college_grad = numeric(),
-  high_school_or_less = numeric(),
-  female = numeric(),
-  male = numeric(),
-  non_binary = numeric(),
-  caucasian = numeric(),
-  hispanic_latino = numeric(),
-  vote_trump = numeric(),
-  vote_biden = numeric(),
-  biden_approve = numeric(),
-  biden_disapprove = numeric(),
-  stringsAsFactors = FALSE
-)
-
-# Rename the columns to the exact strings required
-names(timeline_data) <- c(
+# Define the required column names
+column_names <- c(
   "state",
   "start_date",
   "end_date",
@@ -409,62 +386,56 @@ names(timeline_data) <- c(
   "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove"
 )
 
+# Initialize timeline_data with the required column names and set check.names = FALSE to preserve names
+timeline_data <- data.frame(matrix(ncol = length(column_names), nrow = 0), check.names = FALSE)
+colnames(timeline_data) <- column_names
 
-
-# Iterate through each row in state_data
+# Iterate through each row of `state_data`
 for (i in seq_len(nrow(state_data))) {
-  # Extract state information and url link
-  state_value <- as.character(state_data$state[i])
-  start_date <- as.character(state_data$start_date[i])
-  end_date <- as.character(state_data$end_date[i])
-  party <- as.character(state_data$party[i])
-  pct <- as.numeric(state_data$pct[i])
-  link <- as.character(state_data$url_crosstab[i])
   
-  # Check if the link is not NA
+  # Initialize a new row with NA values for each column in `timeline_data`
+  new_row <- setNames(as.list(rep(NA, length(column_names))), column_names)
+  
+  # Populate basic fields from `state_data`
+  new_row$state <- as.character(state_data$state[i])
+  new_row$start_date <- as.character(state_data$start_date[i])
+  new_row$end_date <- as.character(state_data$end_date[i])
[…]

### 2024-10-31 scripts/06-model_data.R

@@ -1,117 +1,133 @@
 #### Preamble ####
-# Purpose: Models voting patterns based on poll data
+# Purpose: Model voting patterns based on poll data for Democrat and Republican voters
 # Author: Gadiel David Flores
 # Date: 27 October 2024
 # Contact: davidgadiel.flores@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: summarized_state_poll_data.parquet should be available in the data directory
 
 #### Workspace setup ####
 library(here)
 library(arrow)
 library(dplyr)
-library(brms)
-library(caret)
-library(pROC)
-library(glmnet)
+library(brms)     
 
 # Data Import
-analysis_data1 <- read_parquet(here::here("data/02-analysis_data/summarized_state_poll_data.parquet"))
-analysis_data2 <- read_parquet(here::here("data/02-analysis_data/timeline_data.parquet"))
+analysis_data <- read_parquet(here::here("data/02-analysis_data/timeline_data.parquet"))
 
-#### Define Predictors and Target Variable ####
+#### Model Setup for Democrat Voters ####
 
-# Column references provided
-column_refs <- list(
-  democrat_pct = "democrat_pct",
-  republican_pct = "republican_pct",
-  highly_educated = "What is the highest level of education you have attained? College graduate",
-  lower_educated = "What is the highest level of education you have attained? High school or less",
-  female = "Can you please tell me your gender? Female",
-  male = "Can you please tell me your gender? Male",
-  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
-  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
-  hispanic_latino = "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
-  vote_trump = "Who did you vote for in the 2020 election? Donald Trump",
-  vote_biden = "Who did you vote for in the 2020 election? Joe Biden",
-  biden_approve = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
-  biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
-  previous_vote_trump = "Who did you vote for in the 2020 election? Donald Trump",
[…]

### 2024-11-02 to_do_list/To_do_list.qmd

**Format:** pdf

### **Person A: Data Preparation & Pollster Analysis**

- **Task 1: Data Documentation & Review**
  - Review and finalize data cleaning documentation, ensuring reproducibility and clarity.
  - Provide a summary of any challenges encountered during data cleaning, along with adjustments made to the data.

- **Task 2: Model Explanation & Assumptions**
  - Define all model components, including notation and plain-English explanations.
  - Discuss and document model assumptions, limitations, and potential issues.
  - Validate model performance using test/train splits, RMSE calculations, or sensitivity analyses.

- **Task 3: Data Section Write-Up (Assistance with Visualizations)**
  - Collaborate with Person C to finalize the data section, focusing on key visualizations and summary statistics.
  - Ensure all visualizations and statistics are clearly presented.

---

### **Person B: Model Building, Analysis, & Validation**

- **Task 1: Pollster Deep-Dive for Appendix A**
  - Choose a pollster from the data sample and research their methodology in-depth, focusing on survey design (population, frame, sample recruitment, etc.).
  - Write an overview of the pollster’s strengths, weaknesses, and key features.

- **Task 2: Initial Discussion Write-Up**
  - Draft the results and discussion section based on model outcomes, highlighting key findings and their relevance to the election forecast.
  - Outline potential improvements or next steps based on the model’s performance and limitations.

- **Task 3: Results Section Write-Up**
  - Develop tables, graphs, and visualizations to present model findings, ensuring clear interpretation.

---

### **Person C: Documentation, Workflow, & Final Integration**

- **Task 1: Results Section Write-Up**
  - Create tables, graphs, and visualizations to clearly present the model findings.
  - Interpret model results and include any relevant performance metrics (e.g., RMSE, test/train split).

- **Task 2: Reproducible Workflow & Documentation**
  - Organize the GitHub repo with folders (`data/`, `scripts/`, `figures/`, `docs/`).
  - Finalize the README file detailing project structure, data sources, and model.

- **Task 2: Idealized Survey Design for Appendix B**
  - Design an ideal survey methodology for forecasting the election on a $100K budget.
  - Complete the survey on Google Forms, including the link and a PDF export in Appendix B.

- **Task 4: Final Paper Integration & Formatting**
  - Compile and format all sections into a cohesive document by **November 2nd**. Add cross-references to figures, tables, and equations.

---

# Timeline Updates

### **Week 3: October 21st - October 22nd**
- **Person A:** Complete Pollster Deep-Dive for Appendix A and collaborate on the Data Section Write-Up.

### **Week 4: October 23rd - October 29th**
- **Person B:** Complete model development, analysis, and initial validation.

### **Week 5: October 30th - November 4th**
- **Person C:** Finalize integration, quality checks, and prepare for submission.

### 2024-11-03 paper/paper.qmd
Fixed PDF not rendering

@@
 
-*Overview paragraph:* Provide a brief overview of the context of the upcoming US presidential election and the importance of forecasting its results.
-
-*Estimand paragraph:* Define the estimand clearly (e.g., predicting the probability of a candidate winning the election based on poll data).
+*Overview paragraph:* The 2024 United States of America Presidential Election is to elect the fourty-seventh president of USA, with the two candidates as Donald Trump, the fourty-fifth president of the United States and representative of the Republican Party, and Kamala Harris, current vice president of the United States and representative of the Democratic Party, after Joe Biden announced that he will not continue participate in this presidential election anymore (@BidenQuits).
+Democracy can be described as liberal and left-leaning, with the emphasis of economy intervention with implementation of policies such as minimum wage and progressive tax rates.
+Members of the Republican party tend to follow the free economy, and can be described as right-leaning (@RepDem).
+The results of the 2024 Presidential Election will determine which political party, Republican or Democratic, will become the leader of the United States for the next four years, largely affecting the policies implemented in the United States during this time period.
+Due to the strong nation power of the United States, this may even affect the world's economic trends in the four years.
+Therefore, predicting results of the upcoming 2024 United States Presidential Election is a necessary action.
+
+*Estimand paragraph:* To predict results of the 2024 United States of America Presidential Election, our aim is clearly to conclude which candidate, either Harris or Trump, is more likely to win this election.
+Our response variable is in turn the probability of one candidate winning the election based on poll data, with our chosen poll as Emerson.
+When the supporting percentage of one candidate, or the party s/he represents, is only slightly less than 50%, we cannot clearly determine whether s/he would be predicted to have been elected as the fourty-seventh US President or not, since there is a group of people who did not choose to participate in the voting process.
+Therefore, to increase accuracy of predictions, we will build two models, one model with the estimand as the probability of Harris, or the Democratic party, winning the election, and another with the estimand as the probability of Trump, or the Republican party, winning the election.
+By comparing the two predictions, one can get a final prediction of who is more likely to be elected as the next President of the United States of America.
 
@@
 
-\newpage  
 # Data {#sec-data} 
 ## Overview 
+We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and performed an in-depth analysis using the statistical programming language R [@citeR].
 
@@
 ```
+
+Our primary objective was to clean, organize, and analyze U.S.
+presidential election polling data to provide insights into voter preferences by state.
+This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries.
+Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis.
+Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns.
+The table highlights key trends and differences across states in terms of party lean and demographic distribution.
+@fig-predictors-pct-summary-1 and @fig-predictors-pct-summary-2 provides a snapshot of these findings.
 
@@
 ```
-
-We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and analyzed it using the statistical programming language R [@citeR].
-
-Our main goal was to clean, organize, and analyze U.S.
-presidential election polling data to understand voter preferences by state.
-We selected data from credible pollsters, sorted it by state, and resolved missing or incomplete entries.
-Where necessary, we merged additional datasets, ensuring that we carefully handled any gaps in the data.
-The result was a refined dataset ready for analysis.
-The table below provides a summary of polling data by state, showing voter preferences by party and various demographics.
-It highlights key trends and differences across states, focusing on party leanings and demographic splits, as illustrated in @fig-predictors-pct-summary.
 
@@
 ```
-
 The correlation matrix in @fig-predictors-party-pct shows how different predictors relate to party preferences.
@@
 
-# Pollster Methodology Overview (Appendix A) {#sec-appenA}  
-Append a detailed review of a chosen pollster's methodology, survey techniques, strengths, and weaknesses.
-
 # Model {#sec-model}
 ## Model Development
+The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election, focusing specifically on eight swing states.
+We will choose the predictor variables based on the correlation matrix we created previously (@fig-predictors-party-pct).
+It is important to note that even if the percentage of people who vote for Trump (representing the Republican Party) or Harris (representing the Democratic Party) is slightly below 50%, this does not necessarily mean that the corresponding candidate will not be elected as the next President of the United States, given the percentage of people who choose not to vote.
+
+```{r}
+#| label: fig-hist-for-republican
+#| fig-cap: Distribution of Republican Party Supporting Rate
+#| echo: false
+
+# Calculate the variable that represents the Republican Party Supporting Rate
+# grouped based on various states
+data_rep1 <- poll_data %>%
+  group_by(state) %>%
+  summarise(
+    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1)
+  )
+
+# Create a histogram based on targeted data variable
+ggplot(data = data_rep1, aes(x=rep_total_pct)) +
+  geom_histogram(binwidth = 2) +
+  labs(title = "Republic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")
+
+```
+
+The following @fig-hist-for-democratic is the histogram of the Democratic Party Supporting Rate, which describes the distribution of the response variable for the second model.
+Our final choice is to construct two generalized linear models: one for the percentage that Trump will become the next President and one for Harris.
+We can then compare the predicted percentages of people voting for each candidate to make a final prediction of who is more likely to be the next President of the United States.
+
+```{r}
+#| label: fig-hist-for-democratic
+#| fig-cap: Distribution of Democratic Party Supporting Rate
+#| echo: false
+
+# Calculate the variable that represents the Republican Party Supporting Rate
+# grouped based on various states
+data_rep1 <- poll_data %>%
+  group_by(state) %>%
+  summarise(
+    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1)
+  )
+
+# Create a histogram based on targeted data variable
+ggplot(data = data_rep1, aes(x=dem_total_pct)) +
+  geom_histogram(binwidth = 2) +
+  labs(title = "Democratic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")
+
+```
+
+To decide between using a generalized linear model or a linear model for each of the two models, we checked the normality of the response variables: the percentage of people supporting the Republican Party and the percentage supporting the Democratic Party.
+According to @fig-hist-for-democratic and @fig-hist-for-republican, the distributions of both response variables do not appear normal.
+The distribution of the Democratic Party Supporting Rate tends to resemble a Poisson distribution, with most percentages having a similar frequency and an outlier around 67-68 percent.
+The Republican Party Supporting Rate distribution has a maximum frequency around 50 percent, with similar frequencies around 25-26 percent and 48-49 percent, suggesting a potential violation of the normal distribution assumption.
+
+Therefore, since the response variables are not normally distributed, we will use a generalized linear model for both models to better capture the characteristics of the data.
 
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Forecasting model results based on poll data"
+#| tbl-cap: "Prediction Summary for Swing States with Declared Winner"
 #| echo: false
@@
   mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))
+
 # Create a formatted table
@@
   prediction_summary,
-  caption = "Prediction Summary for Swing States with Declared Winner",
   col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
   format = "latex", 
-  digits = 2
+  digits = 2,
+  escape = FALSE
 ) %>%
-  kable_styling(latex_options = c("striped", "hold_position", "condensed"), 
+  kable_styling(
+    latex_options = c("striped", "hold_position", "condensed"), 
     full_width = FALSE, 
-                font_size = 10)
-```
-*Person B*: Visualize the model's results and include any performance metrics (e.g., RMSE, test/train split).
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: ""
-#| warning: false
+    font_size = 10
+)
 
@@
 ## Appendix A: Pollster Methodology Overview
-*Person A*: Provide a detailed methodology review for the chosen pollster, including survey design, population sampling, non-response handling, etc.
+Emerson College Polling was initially a college polling exercise, and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs).
+For the 2024 USA Presidential Election pollster, the Emerson College Polling defined the population to be "likely voters" of 2024 United States Presidential Election.
+Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modelling and the specific model is not indicated, so the specific validity of the targeted population cannot be determine (@Polling).
+
+Emerson College Polling recruits people completing this survey for this national pollster by contacting mobile phones using "MMS-to-Online", "Online Opt-in Panel", and "IVR(Interactive Voice Response)" (@Polling).
+In the MMS-to-Online approach, the target population were sent text messages with graphics that invite them to take a "screening questionnaire", and those who pass the questionnaire can move on to take the survey.
+The selected respondents were based on "state voter files provided by Aristotle." The Online Opt-in Panel approach involves respondents from the targeted population being invited to finish the survey by means of an online opt-in panel "provided by CINT" (@AboutUs).
+Specifically, in the Online Panel approach, selected voters were pre-matched to "L2 voter file data provided by Rep Data" (@Polling).
+Finally, the IVR approach involves making automatic telephone calls to selected "likely voters." Participants then answer the survey using their "touch-tone telephone".
+The email approach indicated in Emerson College Polling's official methodology (@AboutUs) is not mentioned in the National Polling, so it can be assumed that this approach is not used in our chosen pollster data.
+
+From the methodology of recruiting people indicated in the previous paragraph, we can see that while the targeted population of the national poll are likely voters, the sampling frame are likely voters that use phones, either mobile or landline.
+Emerson College Polling chooses 1000 samples from their targeted population, in this case is all likely voters of the 2024 United States Presidential Election, and it is then "weighted by gender, education, race, age, party affiliation, and region based on 2024 likely voter modeling" (@Polling) for each national poll.
+As most people in the United States use phones by some means, the target population and sampling frame are highly similar, increasing the overall validity of the survey.
+
+In the "MMS-to-Online" and "IVR(Interactive Voice Response)" approaches, random sampling is used, and in the "Online Opt-in Panel" approach, there is no indication of a specific sampling approach (@AboutUs).
+We can conclude that the general approach used in this specific pollster is random sampling.
+Random sampling tends to reduce bias, simplify analysis, and is also easy to implement, but at the same time also requires much time and money.
+In consideration with the sample size of 1000 people compared to the targeted population of likely voters, which contains most of the population of the United States of America, there is a high possibility of existing selection bias, occurring when the participants are not representative of the population.
+
+There is also no specific indication of how non-response is treated, but the results section of national polls, such as the one from September 29 to October 1, 2024, implicates that non-response are eliminated from final recorded results of the survey.
+Therefore, we cannot ignore the possibility of non-response bias, occurring when non-respondents differ significantly from respondents.
+For example, such non-response bias occurs when non-respondents are from California, while respondents are from other states except California.
+
+The questionnaire set by Emerson College Polling for this survey, based on the national polls set from September 29 to October 1, 2024, compose mostly of questions with most common choices listed as choices, and also allowing participants to type in their own answer if it is not included in the list of most common choices.
+The questionnaire is generally written in an objective voice, and all questions only allow participants to make only one choice.
+All questions in the questionnaire are relatively-common, including ones about ethnicity, age range, region, and education which collects demographical data, and ones that are related to actual presidential-election predictions, including opinions about Joe Biden's performance, the two major parties (Democratic and Republican), whether s/he would vote, and voting inclinations.
+The only issue observed in the questionnaire is that it is too lengthy, containing more than 20 questions, which may decrease participants' motivations to complete the survey, even after starting it (@Polling).
+
+Therefore, after analyzing the advantages and disadvantages of this pollster about the 2024 United States Presidential Elections by Emerson College Polling, we can conclude that Emerson College Polling can improve their pollster by implementing the following approaches.
+Emerson can consider increasing the sample size in order to make their survey more representative of the targeted population, explicitly describing their definition of the targeted population "likely voters" and the model used to determine characteristics of the targeted population.
+Emerson can also consider changing from simple random to stratified sampling, based on different states or other demographical features based on their targeted population model, making their sample more representative of the population, decrease the length of the questionnaire, and also explicitly explain their treatment of non-response, which will increase the validity of the study.
+Overall, this pollster is still highly valid, and generally reliable for predicting the 2024 United States Presidential Election.
 
 ## Appendix B: Idealized Survey Design for $100K Budget
-*Person C*: Develop and describe an ideal survey design for forecasting the election with a $100K budget, and include survey link.
-
-Sampling approach: The sampling approach we will use is stratified random sampling where we divide the population into strata based on.... (age, gender, education level, etc.)
-Recruit Respondents: We will recruit respondents using an online survey on Google Forms so that we can minimize the cost and maximize the range of respondents we can reach.
-We will spend a portion of our budget (specify here) to advertise these surveys and also send out emails, with an additional monetary incentive (specify here) to encourage more participation.
-Data validation: IP address tracking to prevent duplicate responses.
-Poll Aggregation: Incorporate Bayesian Inference.
-
-Survey Link: (Short questionnaire asking for demographic questions to be added here)
-Copy of Survey:
+
+## Introduction
+
+This appendix introduces an idealized methodology and survey to forecast the US presidential election with a budget of $100k.
+The goal is to use a sample survey that effectively captures public opinion and demographic characteristics.
+
+## Sampling Approach
+
+In order to most accurately represent the sample population of US voters, we will be using a stratified random sampling approach.
+Stratified random sampling efficiently and fairly samples data from strata that are representative of a population compared to simple random sampling since it also reflects the population of interest’s underrepresented subgroups, ultimately reducing potential bias.
+
+The sample population will be taken from 10,000 eligible US voters represented in strata divided into different demographic factors such as age, gender, race, education level, income level, and political affiliation.
+This sample size can provide a statistically significant confidence interval for the strata.
+
+## Respondent Recruitment
+
+We will be using 70% of our budget for respondent recruitment, and the remaining 30% for implementing the survey and validating results.
+
+A proportion of our budget of $20,000 will be used for advertising on various social media platforms such as Instagram, Facebook, and Twitter to reach around 10,000 respondents.
+This online advertising is a good way to reach target respondents based on different demographics we are interested in, especially for younger or hard-to-reach populations.
+
+Since digital ads may result in higher dropout rates and lower response quality, the majority of our budget will be allocated to make up for this limitation.
+A budget of $40,000 can be used to hire professional survey panels that work with companies that can provide large databases of target respondents that have been pre-screened to match our demographic of interest.
+Lastly, since pre-screened professional survey panels can introduce bias from being frequent survey responders, we will also use incentivized online survey platforms to recruit various participants that fit our demographics of interest.
+Participants will be offered small monetary incentives with our budget of $10,000, which can greatly increase completion rates and result in better response quality.
+
+## Data Validation
+
+First, we need pre-screening questions to confirm the eligibility of respondents such as eligibility to vote to prevent bots and fake responses.
+For instance, asking whether they are a US citizen, and then asking if they are eligible to vote in the upcoming US presidential election.
+If their answers contradict these two questions, we know that their response lacks validity.
+
+Additionally, we have attention checks to identify if respondents are paying attention to and understanding the survey questions.
+For example, we can ask them if they are paying attention to the question, and if they answer no, we know that they might be rushing or answering the survey carelessly.
+
+Furthermore, we want consistency checks that include similar questions which are paraphrased in different parts of the survey to ensure that respondents are answering consistently.
+
+Lastly, IP address monitoring can be used to prevent duplicate responses from the same person.
+
+## Poll Aggregation
+
+We can collect poll data from various reputable polling sources that have good pollster reliability and are from more recent polls.
+This way, we can weigh the polls according to their quality.
+For instance, we would assign higher weights to pollsters that have a high rate of pollster reliability based on past results.
+We would also give higher weights to more recent polls that better showcase the current opinions of the upcoming US election.
+Lastly, we would emphasize the weight of larger sample sizes since these provide more accurate representation of the population of interest.
+We can also incorporate the Bayesian Inference method when updating new polls to adjust prior distributions with new information.
+
+Survey Link: https://forms.gle/vm3gKbsMAZvzakFn8
+
+## Copy of Survey
 

### 2024-11-03 paper/paper.tex
Fixed PDF not rendering

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Forecasting the 2024 US Presidential Election},
  pdfauthor={Tina Kim; David Flores; Kevin Shao},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Forecasting the 2024 US Presidential Election\thanks{Code and
data are available at:
https://github.com/DavidFJ207/USPresidentialForecast}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Analyzing Demographic Patterns and Predicting Swing States}
\author{Tina Kim \and David Flores \and Kevin Shao}
\date{November 3, 2024}

\begin{document}
\maketitle
\begin{abstract}
Our analysis of swing state polls for the 2024 U.S. Presidential
Election shows a close race. Republicans lead in most key states like
Michigan, Georgia, Nevada, Wisconsin, and Pennsylvania, while Democrats
have a slight edge in Arizona. The national average also leans
Republican, suggesting a Republican win in the 2024 Election.
\end{abstract}

\section{Introduction}\label{introduction}

\emph{Overview paragraph:} The 2024 United States of America
Presidential Election is to elect the fourty-seventh president of USA,
with the two candidates as Donald Trump, the fourty-fifth president of
the United States and representative of the Republican Party, and Kamala
Harris, current vice president of the United States and representative
of the Democratic Party, after Joe Biden announced that he will not
continue participate in this presidential election anymore (Jr (2024)).
Democracy can be described as liberal and left-leaning, with the
emphasis of economy intervention with implementation of policies such as
minimum wage and progressive tax rates. Members of the Republican party
tend to follow the free economy, and can be described as right-leaning
({``Democrat Vs. Republican''} (n.d.)). The results of the 2024
Presidential Election will determine which political party, Republican
or Democratic, will become the leader of the United States for the next
four years, largely affecting the policies implemented in the United
States during this time period. Due to the strong nation power of the
United States, this may even affect the world's economic trends in the
four years. Therefore, predicting results of the upcoming 2024 United
States Presidential Election is a necessary action.

\emph{Estimand paragraph:} To predict results of the 2024 United States
of America Presidential Election, our aim is clearly to conclude which
candidate, either Harris or Trump, is more likely to win this election.
Our response variable is in turn the probability of one candidate
winning the election based on poll data, with our chosen poll as
Emerson. When the supporting percentage of one candidate, or the party
s/he represents, is only slightly less than 50\%, we cannot clearly
determine whether s/he would be predicted to have been elected as the
fourty-seventh US President or not, since there is a group of people who
did not choose to participate in the voting process. Therefore, to
increase accuracy of predictions, we will build two models, one model
with the estimand as the probability of Harris, or the Democratic party,
winning the election, and another with the estimand as the probability
of Trump, or the Republican party, winning the election. By comparing
the two predictions, one can get a final prediction of who is more
likely to be elected as the next President of the United States of
America.

\emph{Results paragraph:} Summarize the key findings of the model,
highlighting its predictive accuracy and implications.

\emph{Why it matters paragraph:} Explain the broader significance of
accurately forecasting election results for politics, society, and
policymaking.

\emph{Telegraphing paragraph:} The remainder of this paper is structured
as follows: Section 2 details the data and measurement process; Section
3 covers model development and results; Section 4 discusses implications
and future steps. The remainder of this paper is structured as follows.
Section~\ref{sec-data}\ldots.

\section{Data}\label{sec-data}

\subsection{Overview}\label{overview}

We sourced the ``Presidential General Election Polls'' dataset from
FiveThirtyEight (FiveThirtyEight 2024) and performed an in-depth
analysis using the statistical programming language R (R Core Team
2023).

Our primary objective was to clean, organize, and analyze U.S.
presidential election polling data to provide insights into voter
preferences by state. This involved selecting data from reputable
pollsters, organizing it by state, and addressing any missing or
incomplete entries. Additional datasets were merged where necessary, and
after thoroughly addressing missing values, we finalized a polished
dataset ready for analysis. Below is a table summarizing the polling
data by state, showcasing voter preferences by party and demographic
breakdowns. The table highlights key trends and differences across
states in terms of party lean and demographic distribution.
Figure~\ref{fig-predictors-pct-summary-1} and
Figure~\ref{fig-predictors-pct-summary-2} provides a snapshot of these
findings.

\begin{figure}

\centering{

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable*}[t]{lrrrr}
\toprule
State & Democrat Pct & Republican Pct & Voted Biden 2020 & Voted Trump 2020\\
\midrule
Arizona & 48.2 & 50.5 & 46.0 & 45.2\\
Georgia & 49.8 & 49.7 & 44.5 & 44.3\\
Michigan & 49.7 & 49.6 & 57.9 & 26.8\\
Nevada & 49.2 & 48.1 & 57.9 & 26.8\\
New Hampshire & 50.9 & 47.2 & 51.8 & 35.8\\
\addlinespace
North Carolina & 48.2 & 50.2 & 46.0 & 45.2\\
Pennsylvania & 49.0 & 50.7 & 45.3 & 44.9\\
Wisconsin & 49.2 & 49.9 & 58.2 & 28.7\\
\bottomrule
\end{longtable*}
\endgroup{}

}

\caption{\label{fig-predictors-pct-summary-1}}

\end{figure}%

\begin{figure}

\centering{

\begingroup\fontsize{6}{8}\selectfont

\begin{longtable*}[t]{lrrrrrrr}
\toprule
State & High Educated & Low Educated & Female & Male & Nonbinary & Caucasian & Minority Ethnicity\\
\midrule
Arizona & 22.1 & 24.3 & 51.8 & 46.4 & 1.8 & 80.2 & 19.8\\
Georgia & 23.8 & 24.1 & 49.8 & 49.5 & 0.7 & 86.5 & 13.5\\
Michigan & 27.2 & 22.6 & 53.0 & 46.2 & 0.8 & 60.3 & 39.7\\
Nevada & 27.2 & 22.6 & 53.0 & 46.2 & 0.8 & 60.3 & 39.7\\
New Hampshire & 30.1 & 17.7 & 52.5 & 46.7 & 0.8 & 68.3 & 31.7\\
\addlinespace
North Carolina & 22.1 & 24.3 & 51.8 & 46.4 & 1.8 & 80.2 & 19.8\\
Pennsylvania & 25.8 & 21.8 & 53.7 & 44.6 & 1.7 & 60.9 & 39.1\\
Wisconsin & 28.7 & 19.5 & 51.9 & 47.1 & 1.0 & 49.1 & 50.9\\
\bottomrule
\end{longtable*}
\endgroup{}

}

\caption{\label{fig-predictors-pct-summary-2}}

\end{figure}%

Latest Poll Points by Party

\subsection{Measurement}\label{measurement}

We chose Emerson as our primary pollster, and the reasoning for this is
detailed in \textbf{?@sec-appenA}. Emerson's polls offered crucial
information, such as polling dates, party affiliations, sample sizes,
and the percentage of support for each political party.

Our analysis centered on voter preferences by party rather than
individual candidates. This approach was more relevant for identifying
overall trends and helped us forecast which states might swing in future
elections, especially swing states. We will discuss this focus further
in our results section, Section~\ref{sec-results}.

The Emerson dataset also included valuable demographic data, voting
history, and approval ratings. To ensure a consistent analysis, we
concentrated on questions that were uniformly asked across all states.
This approach allowed us to aggregate and analyze data at the state
level without being limited by state-specific polling variations.

\subsection{Outcome Variables}\label{outcome-variables}

Our initial analysis of the dataset revealed important trends in party
preferences across states. In Figure~\ref{fig-party-results}, we show
which states have the strongest leanings toward the Democratic or
Republican parties. This visualization is crucial for identifying key
swing states, which lie near the center of the graph and show nearly
equal support for both parties. These states are especially important
for predicting the 2024 election outcome.

From this analysis, we identified eight critical swing states: Michigan,
Georgia, Nevada, North Carolina, New Hampshire, Wisconsin, Pennsylvania,
and Arizona. These states will be central to our election prediction.
Although Minnesota emerged as a potential swing state, we excluded it
due to the wide range of polling data, as explained in
Table~\ref{tbl-final-summary}. Additionally, we will consider the
national average in our overall prediction.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-party-results-1.pdf}

}

\caption{\label{fig-party-results}Example Predictor Variable
Visualization}

\end{figure}%

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1546}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1340}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1443}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1443}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1340}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1443}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1443}}@{}}

\caption{\label{tbl-final-summary}Final Summary Table of 95\% Confidence
Intervals for Common States}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
State
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
DEM Mean (\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
DEM Lower 95\%
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
DEM Upper 95\%
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
REP Mean (\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
REP Lower 95\%
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
REP Upper 95\%
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Arizona & 42.5 & 40.8 & 44.2 & 47.1 & 45.7 & 48.5 \\
Georgia & 42.8 & 41.0 & 44.7 & 47.8 & 46.7 & 48.9 \\
Michigan & 44.9 & 43.4 & 46.4 & 45.6 & 44.3 & 46.9 \\
National & 42.4 & 41.5 & 43.3 & 43.7 & 42.5 & 44.9 \\
Nevada & 41.8 & 39.6 & 44.0 & 45.9 & 44.6 & 47.2 \\
New Hampshire & 44.8 & 42.6 & 47.0 & 41.6 & 39.9 & 43.4 \\
North Carolina & 43.2 & 40.4 & 46.0 & 48.6 & 47.7 & 49.6 \\
Pennsylvania & 44.2 & 42.5 & 45.9 & 47.6 & 46.6 & 48.6 \\
Wisconsin & 44.4 & 42.7 & 46.0 & 46.5 & 45.2 & 47.8 \\

\end{longtable}

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-party-extremes-1.pdf}

}

\caption{\label{fig-party-extremes}Demographic Differences Between Top
Red and Blue States}

\end{figure}%

Beyond identifying swing states, it's also important to understand the
differences between the strongest Democratic (``blue'') and Republican
(``red'') states. The bar graph in Figure~\ref{fig-party-extremes}
highlights these differences. It focuses on key factors that create the
biggest contrasts between Democratic-leaning and Republican-leaning
states, helping to show how these factors differ between the most
extreme red and blue states.

To start, the graph features two bars, one for Democratic points and one
for Republican points, reflecting the responses to questions asked by
Emerson Pollster about voting intentions. These responses closely align
with the percentages shown for each state, giving us a clear picture of
voter preferences.

The Figure~\ref{fig-party-extremes} visualization gives us valuable
insight into how people's poll responses may influence their political
leanings.

Through these visualizations, we not only identify which states are most
likely to swing in future elections but also better understand the
public opinions driving these partisan divides.

\subsection{Predictor Variables}\label{predictor-variables}

We then take a closer look at the demographic and social factors that
shape voter preferences. The predictors we analyze include education
levels, gender, and past voting behavior. These variables are important
because they significantly impact political leanings and can influence
how a state is likely to vote.

In Figure~\ref{fig-predictors-party-pct}, we examine how these factors
correlate with party lean across U.S. states. This analysis helps
identify which factors have the greatest influence in each state,
providing a clearer understanding of what drives voter behavior.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-predictors-party-pct-1.pdf}

}

\caption{\label{fig-predictors-party-pct}Example Predictor Variable
Visualization}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-predictors-party-pct-2-1.pdf}

}

\caption{\label{fig-predictors-party-pct-2}Example Voting and Approval
Predictor Variable Visualization}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-predictors-education-race-voting-1.pdf}

}

\caption{\label{fig-predictors-education-race-voting}Correlation Matrix
of Education, Race, and Past Voting Behavior}

\end{figure}%

The correlation matrix in Figure~\ref{fig-predictors-party-pct} shows
how different predictors relate to party preferences. We can see which
factors, like education or gender, are more strongly tied to Democratic
or Republican leanings. Blue shading indicates stronger Democratic
support, while red shows stronger Republican support. Key observations
include higher education being linked to Democratic support and lower
education favoring Trump in 2020. Minority groups generally lean
Democratic, while Caucasian/White voters tend to support Trump. Past
voter behavior shows Biden supporters remain Democratic, and Trump
supporters disapprove of Biden. Additionally, women lean slightly more
Democratic compared to men, who often lean Republican.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-points-over-time-1.pdf}

}

\caption{\label{fig-points-over-time}}

\end{figure}%

Lastly, we examine Democratic and Republican support alongside Biden's
approval over time, marking July 21st when Joe Biden concluded his
campaign, as reported by Rabson (2024). During this period, Democratic
support was initially low but increased following this change. By
analyzing specific demographic factors and previous voting behavior, we
aim to predict the 2024 election outcome.

Key predictor variables:

\begin{itemize}
\tightlist
\item
  \textbf{Education}: This measures education levels, with higher
  education often correlating with more Democratic support.
\item
  \textbf{Race/Ethnicity}: This examines how different racial and ethnic
  groups tend to align politically, with minority groups often leaning
  Democratic and White/Caucasian voters leaning Republican.
\item
  \textbf{Gender}: This captures how gender influences party preference,
  with women generally leaning more Democratic and men more Republican.
\item
  \textbf{Priors (Past Voting Behavior)}: This looks at who people voted
  for in the past, helping us understand ongoing political support or
  disapproval.
\item
  \textbf{Biden's Approval Rating}: This assesses how approval or
  disapproval of Joe Biden impacts Democratic support, providing insight
  into shifting voter sentiments.
\end{itemize}

\section{Model}\label{sec-model}

\subsection{Model Development}\label{model-development}

The goal of our model is to forecast the popular vote outcome of the
2024 US presidential election, focusing specifically on eight swing
states. We will choose the predictor variables based on the correlation
matrix we created previously (Figure~\ref{fig-predictors-party-pct}). It
is important to note that even if the percentage of people who vote for
Trump (representing the Republican Party) or Harris (representing the
Democratic Party) is slightly below 50\%, this does not necessarily mean
that the corresponding candidate will not be elected as the next
President of the United States, given the percentage of people who
choose not to vote.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-hist-for-republican-1.pdf}

}

\caption{\label{fig-hist-for-republican}Distribution of Republican Party
Supporting Rate}

\end{figure}%

The following Figure~\ref{fig-hist-for-democratic} is the histogram of
the Democratic Party Supporting Rate, which describes the distribution
of the response variable for the second model. Our final choice is to
construct two generalized linear models: one for the percentage that
Trump will become the next President and one for Harris. We can then
compare the predicted percentages of people voting for each candidate to
make a final prediction of who is more likely to be the next President
of the United States.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-hist-for-democratic-1.pdf}

}

\caption{\label{fig-hist-for-democratic}Distribution of Democratic Party
Supporting Rate}

\end{figure}%

To decide between using a generalized linear model or a linear model for
each of the two models, we checked the normality of the response
variables: the percentage of people supporting the Republican Party and
the percentage supporting the Democratic Party. According to
Figure~\ref{fig-hist-for-democratic} and
Figure~\ref{fig-hist-for-republican}, the distributions of both response
variables do not appear normal. The distribution of the Democratic Party
Supporting Rate tends to resemble a Poisson distribution, with most
percentages having a similar frequency and an outlier around 67-68
percent. The Republican Party Supporting Rate distribution has a maximum
frequency around 50 percent, with similar frequencies around 25-26
percent and 48-49 percent, suggesting a potential violation of the
normal distribution assumption.

Therefore, since the response variables are not normally distributed, we
will use a generalized linear model for both models to better capture
the characteristics of the data.

The goal of our model is to forecast the likely vote outcome in 8 swing
states for the 2024 US presidential election. This model will allow us
to estimate the expected vote share for each candidate based on
state-level characteristics and prior voting patterns, which are
especially important in determining close elections.

\subsection{Model Set-up}\label{model-set-up}

We define the model as a Bayesian linear model where the outcome, ( y\_i
), represents the predicted vote share in state ( i ). The predictors in
the model capture key demographic and historical voting patterns across
swing states, and we aim to model this outcome with an intercept, (
\alpha ), and a state-specific predictor coefficient, ( \beta\_i ), that
represents the effect of a given variable ( x\_i ) on vote share.

\subsubsection{Model Components and
Assumptions}\label{model-components-and-assumptions}

\begin{itemize}
\tightlist
\item
  \textbf{Outcome Variable:} The predicted vote share (as a percentage)
  for each candidate in each state.
\item
  \textbf{Predictors:} Variables include demographic factors (e.g.,
  education, income, race distribution) and historical voting patterns
  (e.g., previous election vote shares, approval ratings).
\item
  \textbf{Intercept:} Represents the baseline vote share in the absence
  of additional predictors.
\item
  \textbf{Coefficients:} These capture the effect size of each predictor
  on vote share, allowing for state-specific adjustments.
\item
  \textbf{Residual Standard Deviation:} Accounts for unexplained
  variation in vote share predictions, assumed to follow an exponential
  distribution.
\end{itemize}

\paragraph{Assumptions}\label{assumptions}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Linearity:} The relationship between each predictor and vote
  share is assumed to be approximately linear.
\item
  \textbf{Independence:} Vote share predictions for each state are
  independent, conditional on the predictors.
\item
  \textbf{Normality of Errors:} Errors in vote share predictions are
  assumed to be normally distributed.
\item
  \textbf{Prior Distributions:} We use weakly informative priors to
  regularize predictions without imposing strong assumptions.
\end{enumerate}

\subsubsection{Implementation}\label{implementation}

We run this model in R using the \texttt{rstanarm} package, which
provides Bayesian estimation with sensible defaults. The default priors
in \texttt{rstanarm} are chosen to be weakly informative to prevent
overfitting while still regularizing our model coefficients.

\subsection{Model Justification and
Results}\label{model-justification-and-results}

By using a Bayesian linear model, we leverage historical and demographic
data to provide probabilistic estimates of vote share, giving us a
distribution of possible outcomes for each state. This approach allows
for uncertainty quantification around predictions, which is particularly
important in close elections.

Our results are summarized in Table Table~\ref{tbl-modelresults}, which
shows the estimated vote shares, posterior means, and credible intervals
for each candidate in each swing state. These findings indicate the
influence of demographic and historical factors on vote share
predictions. Furthermore, the model's predictive accuracy is evaluated
through posterior predictive checks, indicating how well the model
captures actual voting patterns. The implications of these findings are
significant for forecasting election outcomes, particularly in
identifying which states are likely to be more competitive.

\section{Results}\label{sec-results}

Our results are summarized in Table~\ref{tbl-modelresults}.

\begin{table}

\caption{\label{tbl-modelresults}Prediction Summary for Swing States
with Declared Winner}

\centering{

[!h]
\centering\begingroup\fontsize{10}{12}\selectfont

\begin{tabular}{l|r|r|l}
\hline
State & Democrat Points & Republican Points & Winner\\
\hline
\cellcolor{gray!10}{Michigan} & \cellcolor{gray!10}{0.49} & \cellcolor{gray!10}{0.50} & \cellcolor{gray!10}{Republican}\\
\hline
Georgia & 0.49 & 0.50 & Republican\\
\hline
\cellcolor{gray!10}{Nevada} & \cellcolor{gray!10}{0.49} & \cellcolor{gray!10}{0.50} & \cellcolor{gray!10}{Republican}\\
\hline
North Carolina & 0.49 & 0.49 & Republican\\
\hline
\cellcolor{gray!10}{New Hampshire} & \cellcolor{gray!10}{0.49} & \cellcolor{gray!10}{0.49} & \cellcolor{gray!10}{Republican}\\
\hline
Wisconsin & 0.49 & 0.50 & Republican\\
\hline
\cellcolor{gray!10}{Pennsylvania} & \cellcolor{gray!10}{0.49} & \cellcolor{gray!10}{0.50} & \cellcolor{gray!10}{Republican}\\
\hline
Arizona & 0.49 & 0.48 & Democrat\\
\hline
\cellcolor{gray!10}{National} & \cellcolor{gray!10}{0.48} & \cellcolor{gray!10}{0.49} & \cellcolor{gray!10}{Republican}\\
\hline
\end{tabular}
\endgroup{}

}

\end{table}%

\section{Discussion}\label{discussion}

\subsection{First discussion point}\label{sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

\subsection{Second discussion point}\label{second-discussion-point}

Please don't use these as sub-heading labels - change them to be what
your point actually is.

\subsection{Third discussion point}\label{third-discussion-point}

Discuss what the model reveals about the election forecast and its
potential impact on understanding voting behavior.

\emph{Person C}: Discuss limitations of the model and areas for further
improvement.

\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}

Weaknesses and next steps should also be included.

\newpage

\appendix

\section*{Appendix}\label{appendix}
\addcontentsline{toc}{section}{Appendix}

\subsection{Appendix A: Pollster Methodology
Overview}\label{appendix-a-pollster-methodology-overview}

Emerson College Polling was initially a college polling exercise, and is
now an ``innovative, nationally-ranked'' non-partisan polling center
({``About Us''} (n.d.)). For the 2024 USA Presidential Election
pollster, the Emerson College Polling defined the population to be
``likely voters'' of 2024 United States Presidential Election. Emerson
College Polling does indicate that the targeted population of ``likely
voters'' is based on ``2024 Likely Voter Modeling.'' However, the
methodology for this modelling and the specific model is not indicated,
so the specific validity of the targeted population cannot be determine
({``October 2024 National Poll: Harris 50''} (2024)).

Emerson College Polling recruits people completing this survey for this
national pollster by contacting mobile phones using ``MMS-to-Online'',
``Online Opt-in Panel'', and ``IVR(Interactive Voice Response)''
({``October 2024 National Poll: Harris 50''} (2024)). In the
MMS-to-Online approach, the target population were sent text messages
with graphics that invite them to take a ``screening questionnaire'',
and those who pass the questionnaire can move on to take the survey. The
selected respondents were based on ``state voter files provided by
Aristotle.'' The Online Opt-in Panel approach involves respondents from
the targeted population being invited to finish the survey by means of
an online opt-in panel ``provided by CINT'' ({``About Us''} (n.d.)).
Specifically, in the Online Panel approach, selected voters were
pre-matched to ``L2 voter file data provided by Rep Data'' ({``October
2024 National Poll: Harris 50''} (2024)). Finally, the IVR approach
involves making automatic telephone calls to selected ``likely voters.''
Participants then answer the survey using their ``touch-tone
telephone''. The email approach indicated in Emerson College Polling's
official methodology ({``About Us''} (n.d.)) is not mentioned in the
National Polling, so it can be assumed that this approach is not used in
our chosen pollster data.

From the methodology of recruiting people indicated in the previous
paragraph, we can see that while the targeted population of the national
poll are likely voters, the sampling frame are likely voters that use
phones, either mobile or landline. Emerson College Polling chooses 1000
samples from their targeted population, in this case is all likely
voters of the 2024 United States Presidential Election, and it is then
``weighted by gender, education, race, age, party affiliation, and
region based on 2024 likely voter modeling'' ({``October 2024 National
Poll: Harris 50''} (2024)) for each national poll. As most people in the
United States use phones by some means, the target population and
sampling frame are highly similar, increasing the overall validity of
the survey.

In the ``MMS-to-Online'' and ``IVR(Interactive Voice Response)''
approaches, random sampling is used, and in the ``Online Opt-in Panel''
approach, there is no indication of a specific sampling approach
({``About Us''} (n.d.)). We can conclude that the general approach used
in this specific pollster is random sampling. Random sampling tends to
reduce bias, simplify analysis, and is also easy to implement, but at
the same time also requires much time and money. In consideration with
the sample size of 1000 people compared to the targeted population of
likely voters, which contains most of the population of the United
States of America, there is a high possibility of existing selection
bias, occurring when the participants are not representative of the
population.

There is also no specific indication of how non-response is treated, but
the results section of national polls, such as the one from September 29
to October 1, 2024, implicates that non-response are eliminated from
final recorded results of the survey. Therefore, we cannot ignore the
possibility of non-response bias, occurring when non-respondents differ
significantly from respondents. For example, such non-response bias
occurs when non-respondents are from California, while respondents are
from other states except California.

The questionnaire set by Emerson College Polling for this survey, based
on the national polls set from September 29 to October 1, 2024, compose
mostly of questions with most common choices listed as choices, and also
allowing participants to type in their own answer if it is not included
in the list of most common choices. The questionnaire is generally
written in an objective voice, and all questions only allow participants
to make only one choice. All questions in the questionnaire are
relatively-common, including ones about ethnicity, age range, region,
and education which collects demographical data, and ones that are
related to actual presidential-election predictions, including opinions
about Joe Biden's performance, the two major parties (Democratic and
Republican), whether s/he would vote, and voting inclinations. The only
issue observed in the questionnaire is that it is too lengthy,
containing more than 20 questions, which may decrease participants'
motivations to complete the survey, even after starting it ({``October
2024 National Poll: Harris 50''} (2024)).

Therefore, after analyzing the advantages and disadvantages of this
pollster about the 2024 United States Presidential Elections by Emerson
College Polling, we can conclude that Emerson College Polling can
improve their pollster by implementing the following approaches. Emerson
can consider increasing the sample size in order to make their survey
more representative of the targeted population, explicitly describing
their definition of the targeted population ``likely voters'' and the
model used to determine characteristics of the targeted population.
Emerson can also consider changing from simple random to stratified
sampling, based on different states or other demographical features
based on their targeted population model, making their sample more
representative of the population, decrease the length of the
questionnaire, and also explicitly explain their treatment of
non-response, which will increase the validity of the study. Overall,
this pollster is still highly valid, and generally reliable for
predicting the 2024 United States Presidential Election.

\subsection{Appendix B: Idealized Survey Design for \$100K
Budget}\label{appendix-b-idealized-survey-design-for-100k-budget}

\subsection{Introduction}\label{introduction-1}

This appendix introduces an idealized methodology and survey to forecast
the US presidential election with a budget of \$100k. The goal is to use
a sample survey that effectively captures public opinion and demographic
characteristics.

\subsection{Sampling Approach}\label{sampling-approach}

In order to most accurately represent the sample population of US
voters, we will be using a stratified random sampling approach.
Stratified random sampling efficiently and fairly samples data from
strata that are representative of a population compared to simple random
sampling since it also reflects the population of interest's
underrepresented subgroups, ultimately reducing potential bias.

The sample population will be taken from 10,000 eligible US voters
represented in strata divided into different demographic factors such as
age, gender, race, education level, income level, and political
affiliation. This sample size can provide a statistically significant
confidence interval for the strata.

\subsection{Respondent Recruitment}\label{respondent-recruitment}

We will be using 70\% of our budget for respondent recruitment, and the
remaining 30\% for implementing the survey and validating results.

A proportion of our budget of \$20,000 will be used for advertising on
various social media platforms such as Instagram, Facebook, and Twitter
to reach around 10,000 respondents. This online advertising is a good
way to reach target respondents based on different demographics we are
interested in, especially for younger or hard-to-reach populations.

Since digital ads may result in higher dropout rates and lower response
quality, the majority of our budget will be allocated to make up for
this limitation. A budget of \$40,000 can be used to hire professional
survey panels that work with companies that can provide large databases
of target respondents that have been pre-screened to match our
demographic of interest. Lastly, since pre-screened professional survey
panels can introduce bias from being frequent survey responders, we will
also use incentivized online survey platforms to recruit various
participants that fit our demographics of interest. Participants will be
offered small monetary incentives with our budget of \$10,000, which can
greatly increase completion rates and result in better response quality.

\subsection{Data Validation}\label{data-validation}

First, we need pre-screening questions to confirm the eligibility of
respondents such as eligibility to vote to prevent bots and fake
responses. For instance, asking whether they are a US citizen, and then
asking if they are eligible to vote in the upcoming US presidential
election. If their answers contradict these two questions, we know that
their response lacks validity.

Additionally, we have attention checks to identify if respondents are
paying attention to and understanding the survey questions. For example,
we can ask them if they are paying attention to the question, and if
they answer no, we know that they might be rushing or answering the
survey carelessly.

Furthermore, we want consistency checks that include similar questions
which are paraphrased in different parts of the survey to ensure that
respondents are answering consistently.

Lastly, IP address monitoring can be used to prevent duplicate responses
from the same person.

\subsection{Poll Aggregation}\label{poll-aggregation}

We can collect poll data from various reputable polling sources that
have good pollster reliability and are from more recent polls. This way,
we can weigh the polls according to their quality. For instance, we
would assign higher weights to pollsters that have a high rate of
pollster reliability based on past results. We would also give higher
weights to more recent polls that better showcase the current opinions
of the upcoming US election. Lastly, we would emphasize the weight of
larger sample sizes since these provide more accurate representation of
the population of interest. We can also incorporate the Bayesian
Inference method when updating new polls to adjust prior distributions
with new information.

Survey Link: https://forms.gle/vm3gKbsMAZvzakFn8

\subsection{Copy of Survey}\label{copy-of-survey}

\subsection{Additional Data \& Model
Details}\label{additional-data-model-details}

Include any technical details on data cleaning, model diagnostics, and
posterior checks.

\section{Additional data details}\label{additional-data-details}

\section{Model details}\label{sec-model-details}

\subsection{Posterior predictive
check}\label{posterior-predictive-check}

In \textbf{?@fig-ppcheckandposteriorvsprior-1} we implement a posterior
predictive check. This shows\ldots{}

In \textbf{?@fig-ppcheckandposteriorvsprior-2} we compare the posterior
with the prior. This shows\ldots{}

\begin{figure}

\begin{minipage}{0.50\linewidth}
Examining how the model fits, and is affected by, the
data\end{minipage}%

\end{figure}%

\subsection{Diagnostics}\label{diagnostics}

\textbf{?@fig-stanareyouokay-1} is a trace plot. It shows\ldots{} This
suggests\ldots{}

\textbf{?@fig-stanareyouokay-2} is a Rhat plot. It shows\ldots{} This
suggests\ldots{}

\begin{figure}

\begin{minipage}{0.50\linewidth}
Checking the convergence of the MCMC algorithm\end{minipage}%

\end{figure}%

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-AboutUs}
{``About Us.''} n.d. Emerson College Polling.
\url{https://emersoncollegepolling.com/about/}.

\bibitem[\citeproctext]{ref-RepDem}
{``Democrat Vs. Republican.''} n.d. Diffen.
\url{https://www.diffen.com/difference/Democrat_vs_Republican}.

\bibitem[\citeproctext]{ref-FiveThirtyEight2024}
FiveThirtyEight. 2024. {``Dataset: US Presidential General Election
Polls.''}
\url{https://projects.fivethirtyeight.com/polls/data/president_polls.csv}.

\bibitem[\citeproctext]{ref-BidenQuits}
Jr, Bernd Debusmann. 2024. {``Biden Says He Quit US Presidential Race to
'Save Democracy'.''} BBC News.
\url{https://www.bbc.com/news/articles/c047281jj8do}.

\bibitem[\citeproctext]{ref-Polling}
{``October 2024 National Poll: Harris 50.''} 2024. Emerson College
Polling.
\url{https://emersoncollegepolling.com/october-2024-national-poll-harris-50-trump-48/}.

\bibitem[\citeproctext]{ref-citeR}
R Core Team. 2023. \emph{{R: A Language and Environment for Statistical
Computing}}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-globalnews2024}
Rabson, Mia. 2024. {``Kamala Harris Would Be Democrats' {`Best Choice'}
If Biden Doesn't Run, Expert Says.''}
\url{https://globalnews.ca/news/10683018/kamala-harris-democratic-nomination-us-election-2/}.

\end{CSLReferences}

\end{document}

### 2024-11-03 paper/paper.qmd
Minor fixes in pdf output

@@
 
-## Appendix A: Pollster Methodology Overview
+## Appendix A: Pollster Methodology Overview {-}
 Emerson College Polling was initially a college polling exercise, and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs).
@@
 
-## Appendix B: Idealized Survey Design for $100K Budget
-
-## Introduction
+## Appendix B: Idealized Survey Design for $100K Budget {-}
+
+## Introduction {-}
 
@@
 
-## Sampling Approach
+## Sampling Approach {-}
 
@@
 
-## Respondent Recruitment
+## Respondent Recruitment {-}
 
@@
 
-## Data Validation
+## Data Validation {-}
 
@@
 
-## Poll Aggregation
+## Poll Aggregation {-}
 
@@
 
-## Copy of Survey
+## Copy of Survey {-}
 

### 2024-11-03 paper/paper.qmd
Updated Appendix B with a copy of the survey and the survey link

@@
 
+A copy of the survey and its questions are provided below:
+
+**2024 US Presidential Election Survey**
+
+Thank you for considering participation in the 2024 US Presidential Election Survey.
+The purpose of this survey is to better understand and forecast election results.
+This survey gathers anonymous data on voter opinions and demographics for the 2024 US presidential election.
+You may stop the survey at any time without consequence.
+All responses will remain anonymous, and results will be reported in aggregate form only.
+A small monetary incentive will be offered upon completion.
+
+If you have any questions or concerns about this survey, please reach out to any of the researchers below:
+
+Tina Kim, University of Toronto: tinak.kim@mail.utoronto.ca
+David Flores, University of Toronto: davidgadiel.flores@mail.utoronto.ca
+Kevin Shao, University of Toronto: kevin.shao@mail.utoronto.ca
+
+By proceeding, you confirm your consent to participate.
+  - I consent to participate.
+  - I do not consent to participate.
+
+1. **Are you a US citizen?**
+  - Yes
+  - No
+
+2. **Are you eligible to vote in the upcoming US presidential election?**
+  - Yes
+  - No
+
+3. **What is your age group?**
+Under 18
+  - 18 - 24
+  - 25 - 34
+  - 35 - 44
+  - 45 - 54
+  - 55 - 64
+  - 65 or older
+  - Prefer not to say
+
+4. **What is your gender?**
+  - Male
+  - Female
+  - Non-binary
+  - Prefer not to say
+
+5. **What is your ethnicity? (choose all that apply)**
+  - White / Caucasian
+  - Black or African American
+  - Hispanic or Latino
+  - Asian
+  - Native American or Alaska Native
+  - Native Hawaiian or Other Pacific Islander
+  - Middle Eastern or North African
+  - Other
+  - Prefer not to say
+
+6. **What is the highest level of education you have completed?**
+  - Less than high school
+  - High school diploma or equivalent (e.g., GED)
+  - Some college, no degree
+  - Associate’s degree
+  - Bachelor’s degree
+  - Master’s degree
+  - Professional degree (e.g., JD, MD)
+  - Doctorate (e.g., PhD, EdD)
+  - Prefer not to say
+
+7. **Are you paying attention to this survey?**
+  - Yes
+  - No
+
+8. **What is your income level?**
+  - Less than $20,000
+  - $20,000 - $39,999
+  - $40,000 - $59,999
+  - $60,000 - $79,999
+  - $80,000 - $99,999
+  - $100,000 - $149,999
+  - $150,000 - $199,999
+  - $200,000 or more
+  - Prefer not to say
+
+9. **What is the highest level of education you have attained?**
+  - Less than high school
+  - High school diploma or equivalent (e.g., GED)
+  - Some college, no degree
+  - Associate’s degree
+  - Bachelor’s degree
+  - Master’s degree
+  - Professional degree (e.g., JD, MD)
+  - Doctorate (e.g., PhD, EdD)
+  - Prefer not to say
+
+10. **What is your political affiliation?**
+  - Democrat
+  - Republican
+  - Independent
+  - Other
+  - Prefer not to say
+
+11. **If the election were held today, which candidate would you vote for?**
+  - Candidate A
+  - Candidate B
+  - Candidate C
+  - Other
+  - Undecided
+  - Prefer not to say
+
 ## Additional Data & Model Details

### 2024-11-03 paper/paper.qmd
Added latest data

--- 2024-11-03 paper/paper.qmd
@@
   "Who did you vote for in the 2020 election? Joe Biden",
-  "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
-  "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove"
+  "Although you are undecided, which candidate do you lean toward? Donald Trump",
+  "Although you are undecided, which candidate do you lean toward? Kamala Harris"
 )
@@
 
-From this analysis, we identified eight critical swing states: Michigan, Georgia, Nevada, North Carolina, New Hampshire, Wisconsin, Pennsylvania, and Arizona.
+From this analysis, we identified eight critical swing states: Georgia, Michigan, National, Nevada, New Hampshire, North Carolina, Pennsylvania, and Wisconsin.
 These states will be central to our election prediction.
@@
 Additionally, we will consider the national average in our overall prediction.
+
 ```{r}
@@
   # Filter out intervals where the bounds go below 40 or above 60, and remove rows where mean_pct exceeds 50
-  filter(!(lower_95 < 39 | upper_95 > 60) & mean_pct <= 100)
+  filter(!(lower_95 < 35 | upper_95 > 65) & mean_pct <= 100)
 
@@
   # Filter out intervals where the bounds go below 40 or above 60, and remove rows where mean_pct exceeds 50
-  filter(!(lower_95 < 39 | upper_95 > 60) & mean_pct <= 50)
+  filter(!(lower_95 < 35 | upper_95 > 65) & mean_pct <= 50)
 
@@
 #| label: fig-party-extremes
-#| fig-cap: Demographic Differences Between Top Red and Blue States
-#| echo: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
+#| fig-cap: Demographic Differences Between Maryland and South Dakota
+#| echo: false
+#| warning: false
+#| message: false
 
@@
   select(-contains('Mode')) %>%             # Remove columns containing 'Mode' in their names
-  select(-contains('registration')) %>%     # Remove columns containing 'registration' in their names
-  drop_na()                                 # Remove rows with any NA values
-
-# Function to identify top N blue and red states
-identify_top_states <- function(data, n) {
-  data %>%
-    arrange(desc(party_diff)) %>%
-    summarise(
-      blue_states = list(state[1:n]),  
-      red_states = list(state[(nrow(data) - n + 1):nrow(data)])
-    )
-}
+  select(-contains('registration'))
 
 # Function to compute mean differences for all numeric variables
-compute_mean_differences <- function(data, blue_states, red_states) {
+compute_mean_differences <- function(data, blue_state, red_state) {
   numeric_vars <- data %>%
@@
   blue_means <- data %>%
-    filter(state %in% blue_states) %>%
+    filter(state == blue_state) %>%
     summarise(across(all_of(numeric_vars), mean, na.rm = TRUE))
@@
   red_means <- data %>%
-    filter(state %in% red_states) %>%
+    filter(state == red_state) %>%
     summarise(across(all_of(numeric_vars), mean, na.rm = TRUE))
@@
     ) %>%
-    mutate(Group = recode(Group, "Blue_Mean" = "Republican State", "Red_Mean" = "Democrat State"))
+    mutate(Group = recode(Group, "Blue_Mean" = "Maryland", "Red_Mean" = "South Dakota"))
 }
@@
     # Update the colors and labels for the groups
-    scale_fill_manual(values = c("Republican State" = "red", "Democrat State" = "blue")) +
+    scale_fill_manual(values = c("Maryland" = "blue", "South Dakota" = "red")) +
     labs(
-      title = "Demographic Differences Between Top Red and Blue States",
+      title = "Demographic Differences Between Maryland and South Dakota",
       x = "Demographic Variable",
       y = "Mean Percentage",
-      fill = "State Group"
+      fill = "State"
     ) +
@@
 # Main analysis function
-analyze_state_differences <- function(data, top_n_states = 1, top_n_vars = 4) {
-  top_states <- identify_top_states(data, n = top_n_states)
-
-  diff_df <- compute_mean_differences(
-    data, 
-    blue_states = top_states$blue_states[[1]], 
-    red_states = top_states$red_states[[1]]
-  )
+analyze_state_differences <- function(data, blue_state, red_state, top_n_vars = 4) {
+  diff_df <- compute_mean_differences(data, blue_state, red_state)
 
@@
 
-# Run the analysis
-analysis_results <- analyze_state_differences(poll_summary, top_n_states = 1, top_n_vars = 5)
+# Run the analysis for Maryland and South Dakota
+analysis_results <- analyze_state_differences(poll_summary, blue_state = "Maryland", red_state = "South Dakota", top_n_vars = 5)
 
@@
   male = "Can you please tell me your gender? Male",
-  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
   caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
@@
   vote_biden = "Who did you vote for in the 2020 election? Joe Biden",
-  biden_approve = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
-  biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
-  previous_vote_trump ="Who did you vote for in the 2020 election? Donald Trump",
-  previous_vote_joe ="Who did you vote for in the 2020 election? Joe Biden"
+  lean_trump ="Although you are undecided, which candidate do you lean toward? Donald Trump",
+  lean_Kamala ="Although you are undecided, which candidate do you lean toward? Kamala Harris"
 
@@
     male_pct = round(mean(.data[[column_refs$male]], na.rm = TRUE), 1),
-    non_binary_other_pct = round(mean(.data[[column_refs$non_binary_other]], na.rm = TRUE), 1),
+    lean_trump_pct = round(mean(.data[[column_refs$lean_trump]], na.rm = TRUE), 1),
+    lean_kamala_pct = round(mean(.data[[column_refs$lean_Kamala]], na.rm = TRUE), 1),
     caucasian_pct = round(mean(.data[[column_refs$caucasian]], na.rm = TRUE), 1),
@@
     Male = male_pct,
-    NonBinary = non_binary_other_pct,
+    LeanTrump = lean_trump_pct,
+    LeanKamala = lean_kamala_pct,
     Caucasian = caucasian_pct,
     Minority = minority_pct
-    # No age-related columns included anymore
   )
@@
   )
+
 ```
@@
 #| label: fig-predictors-party-pct-2
-#| fig-cap: Example Voting and Approval Predictor Variable Visualization
-#| echo: false
-#| warning: false
-#| message: false
-# Calculate poll_summary with rounding to the nearest tenth
+#| fig-cap: Example Voting and Leaning Predictor Variable Visualization
+#| echo: false
+#| warning: false
+#| message: false
+
+# Calculate poll_summary_2 with rounding to the nearest tenth
 poll_summary_2 <- poll_data %>%
@@
     vote_biden_pct = round(mean(.data[[column_refs$vote_biden]], na.rm = TRUE), 1),
-    biden_approve_pct = round(mean(.data[[column_refs$biden_approve]], na.rm = TRUE), 1),
-    biden_disapprove_pct = round(mean(.data[[column_refs$biden_disapprove]], na.rm = TRUE), 1)
+    lean_trump_pct = round(mean(.data[[column_refs$lean_trump]], na.rm = TRUE), 1),
+    lean_kamala_pct = round(mean(.data[[column_refs$lean_Kamala]], na.rm = TRUE), 1)
   ) %>%
@@
     VotedBiden2020 = vote_biden_pct,
-    ApprovesOfBiden = biden_approve_pct,
-    DisapprovesOfBiden = biden_disapprove_pct
+    LeanTrump = lean_trump_pct,
+    LeanKamala = lean_kamala_pct
   )
@@
 
-```{r}
-#| label: fig-predictors-education-race-voting
-#| fig-cap: Correlation Matrix of Education, Race, and Past Voting Behavior
-#| echo: false
-#| warning: false
-#| message: false
-
-# Calculate poll_summary with rounding to the nearest tenth
-poll_summary_3 <- poll_data %>%
-  group_by(state) %>%
-  summarise(
-    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1),
-    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1),
-    highly_educated_pct = round(mean(.data[[column_refs$highly_educated]], na.rm = TRUE), 1),
-    lower_educated_pct = round(mean(.data[[column_refs$lower_educated]], na.rm = TRUE), 1),
-    caucasian_pct = round(mean(.data[[column_refs$caucasian]], na.rm = TRUE), 1),
-    minority_pct = round(100 - caucasian_pct, 1),
-    vote_trump_pct = round(mean(.data[[column_refs$vote_trump]], na.rm = TRUE), 1),
-    vote_biden_pct = round(mean(.data[[column_refs$vote_biden]], na.rm = TRUE), 1)
-  ) %>%
-  mutate(party_diff = round(dem_total_pct - rep_total_pct, 1)) %>%
-  drop_na()
-
-# Select relevant numeric columns for correlation matrix, excluding PartyDifference
-corr_data_3 <- poll_summary_3 %>%
-  select(
-    Democrat = dem_total_pct,
-    Republican = rep_total_pct,
-    HighlyEducated = highly_educated_pct,
-    LowerEducated = lower_educated_pct,
-    Caucasian = caucasian_pct,
-    Minority = minority_pct,
-    VotedTrump2020 = vote_trump_pct,
-    VotedBiden2020 = vote_biden_pct
-  )
-
-# Calculate the correlation matrix
-cor_matrix_3 <- cor(corr_data_3, use = "complete.obs")
-
-# Set correlations less than -0.9 to NA
-cor_matrix_3[cor_matrix_3 < -0.85] <- NA
-
-# Visualize the correlation matrix without hierarchical clustering
-ggcorrplot(cor_matrix_3, 
-           hc.order = FALSE,  
-           type = "lower", 
-           lab = TRUE, 
-           lab_size = 2,  # Size for correlation labels
-           title = "Correlation Matrix of Education, Race, and Past Voting Behavior",
-           colors = c("blue", "white", "red"),  
-           legend.title = "Correlation") +
-  theme(
-    plot.title = element_text(size = 12, hjust = 0.5),  # Title size and centered
-    axis.text = element_text(size = 1),  # Axis text size
-    legend.text = element_text(size = 10),  # Legend text size
-    legend.title = element_text(size = 10),  # Legend title size
-    axis.title = element_text(size = 1)  # Axis labels size
-  )
-
-```
 The correlation matrix in @fig-predictors-party-pct shows how different predictors relate to party preferences.
@@
 #| label: fig-points-over-time
-#| tbl-cap: "Parties Approval"
-#| echo: false
-#| warning: false
-#| message: false
+#| tbl-cap: "Parties Approval and Leaning Preferences"
+#| echo: false
+#| warning: false
+#| message: false
+
 # Convert `end_date` to Date format, ensure numeric columns, and limit to 2024 and onward
@@
     pct = as.numeric(pct),
-    biden_approval = as.numeric(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`)
+    lean_trump = as.numeric(`Although you are undecided, which candidate do you lean toward? Donald Trump`),
+    lean_kamala = as.numeric(`Although you are undecided, which candidate do you lean toward? Kamala Harris`)
   ) %>%
-  filter(end_date >= as.Date("2024-01-01"))  # Limit data to 2024 and onward
-
-# Separate data for Democrats, Republicans, and Biden Approval
+  filter(end_date >= as.Date("2024-01-01")) %>%  # Limit data to 2024 and onward
+  arrange(end_date)  # Ensure data is sorted by end_date
+
+# Separate data for Democrats and Republicans
 democrat_data <- time_data_graph %>%
   filter(party == "DEM", !is.na(pct)) %>%
-  filter(state == "National", !is.na(pct))
+  filter(state == "National") %>%
+  arrange(end_date)
 
@@
   filter(party == "REP", !is.na(pct)) %>%
-  filter(state == "National", !is.na(pct))
-
-biden_approval <- time_data_graph  %>%
-  filter(state == "National", !is.na(pct))%>%
-  select(end_date, biden_approval) %>%
-  filter(!is.na(biden_approval)) %>%
-  distinct()
+  filter(state == "National") %>%
+  arrange(end_date)
+
+# Average `lean_trump` values for the same date, keep one entry, and limit to 4 entries per month
+lean_trump_data <- time_data_graph %>%
+  select(end_date, lean_trump) %>%
+  filter(!is.na(lean_trump)) %>%
+  group_by(month = floor_date(end_date, "month")) %>%
+  arrange(end_date) %>%
+  slice_head(n = 4) %>%  # Limit to 4 entries per month
+  ungroup() %>%
+  select(-month) %>%
+  group_by(end_date) %>%
+  summarise(lean_trump = mean(lean_trump, na.rm = TRUE)) %>%  # Average values
+  arrange(end_date)
+
+# Average `lean_kamala` values for the same date, keep one entry, and limit to 4 entries per month
+lean_kamala_data <- time_data_graph %>%
+  select(end_date, lean_kamala) %>%
+  filter(!is.na(lean_kamala)) %>%
+  group_by(month = floor_date(end_date, "month")) %>%
+  arrange(end_date) %>%
+  slice_head(n = 4) %>%  # Limit to 4 entries per month
+  ungroup() %>%
+  select(-month) %>%
+  group_by(end_date) %>%
+  summarise(lean_kamala = mean(lean_kamala, na.rm = TRUE)) %>%  # Average values
+  arrange(end_date)
 
@@
 
-  # Plot Biden's Approval over time (light blue)
-  geom_line(data = biden_approval, aes(x = end_date, y = biden_approval, color = "Biden Approval"), size = 1) +
-
-  # Add a vertical line for a key date
-  geom_vline(xintercept = as.Date("2024-07-21"), color = "red", linetype = "dashed") +
+  # Plot Leaning Toward Trump over time (dark orange)
+  geom_line(data = lean_trump_data, aes(x = end_date, y = lean_trump, color = "Lean Trump"), size = 0.6) +
+
+  # Plot Leaning Toward Kamala over time (purple)
+  geom_line(data = lean_kamala_data, aes(x = end_date, y = lean_kamala, color = "Lean Kamala"), size = 0.6) +
 
@@
   labs(
-    title = "Democrat vs. Republican and Biden Approval Over Time",
+    title = "Democrat vs. Republican and Leaning Preferences Over Time",
     x = "Date",
@@
   ) +
-  scale_color_manual(values = c("Democrats" = "blue", "Republicans" = "red", "Biden Approval" = "lightblue")) +
+  scale_color_manual(values = c("Democrats" = "blue", "Republicans" = "red", "Lean Trump" = "darkorange", "Lean Kamala" = "purple")) +
   ylim(25, 55) +  # Set y-axis range from 25 to 55, adjust if needed
@@
 ## Model Development
-The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election, focusing specifically on eight swing states.
-We will choose the predictor variables based on the correlation matrix we created previously (@fig-predictors-party-pct).
-It is important to note that even if the percentage of people who vote for Trump (representing the Republican Party) or Harris (representing the Democratic Party) is slightly below 50%, this does not necessarily mean that the corresponding candidate will not be elected as the next President of the United States, given the percentage of people who choose not to vote.
-
-```{r}
-#| label: fig-hist-for-republican
-#| fig-cap: Distribution of Republican Party Supporting Rate
-#| echo: false
-
-# Calculate the variable that represents the Republican Party Supporting Rate
-# grouped based on various states
-data_rep1 <- poll_data %>%
-  group_by(state) %>%
-  summarise(
-    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1)
-  )
-
-# Create a histogram based on targeted data variable
-ggplot(data = data_rep1, aes(x=rep_total_pct)) +
-  geom_histogram(binwidth = 2) +
-  labs(title = "Republic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")
-
-```
-
-The following @fig-hist-for-democratic is the histogram of the Democratic Party Supporting Rate, which describes the distribution of the response variable for the second model.
-Our final choice is to construct two generalized linear models: one for the percentage that Trump will become the next President and one for Harris.
-We can then compare the predicted percentages of people voting for each candidate to make a final prediction of who is more likely to be the next President of the United States.
-
-```{r}
-#| label: fig-hist-for-democratic
-#| fig-cap: Distribution of Democratic Party Supporting Rate
-#| echo: false
-
-# Calculate the variable that represents the Republican Party Supporting Rate
-# grouped based on various states
-data_rep1 <- poll_data %>%
-  group_by(state) %>%
-  summarise(
-    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1)
-  )
-
-# Create a histogram based on targeted data variable
-ggplot(data = data_rep1, aes(x=dem_total_pct)) +
-  geom_histogram(binwidth = 2) +
-  labs(title = "Democratic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")
-
-```
-
-To decide between using a generalized linear model or a linear model for each of the two models, we checked the normality of the response variables: the percentage of people supporting the Republican Party and the percentage supporting the Democratic Party.
-According to @fig-hist-for-democratic and @fig-hist-for-republican, the distributions of both response variables do not appear normal.
-The distribution of the Democratic Party Supporting Rate tends to resemble a Poisson distribution, with most percentages having a similar frequency and an outlier around 67-68 percent.
-The Republican Party Supporting Rate distribution has a maximum frequency around 50 percent, with similar frequencies around 25-26 percent and 48-49 percent, suggesting a potential violation of the normal distribution assumption.
-
-Therefore, since the response variables are not normally distributed, we will use a generalized linear model for both models to better capture the characteristics of the data.
-
-The goal of our model is to forecast the likely vote outcome in 8 swing states for the 2024 US presidential election.
-This model will allow us to estimate the expected vote share for each candidate based on state-level characteristics and prior voting patterns, which are especially important in determining close elections.
-
-## Model Set-up
-
-We define the model as a Bayesian linear model where the outcome, \( y_i \), represents the predicted vote share in state \( i \).
-The predictors in the model capture key demographic and historical voting patterns across swing states, and we aim to model this outcome with an intercept, \( \alpha \), and a state-specific predictor coefficient, \( \beta_i \), that represents the effect of a given variable \( x_i \) on vote share.
-
-### Model Components and Assumptions
-
-- **Outcome Variable:** The predicted vote share (as a percentage) for each candidate in each state.
-- **Predictors:** Variables include demographic factors (e.g., education, income, race distribution) and historical voting patterns (e.g., previous election vote shares, approval ratings).
-- **Intercept:** Represents the baseline vote share in the absence of additional predictors.
-- **Coefficients:** These capture the effect size of each predictor on vote share, allowing for state-specific adjustments.
-- **Residual Standard Deviation:** Accounts for unexplained variation in vote share predictions, assumed to follow an exponential distribution.
-
-#### Assumptions
-1. **Linearity:** The relationship between each predictor and vote share is assumed to be approximately linear.
-2. **Independence:** Vote share predictions for each state are independent, conditional on the predictors.
-3. **Normality of Errors:** Errors in vote share predictions are assumed to be normally distributed.
-4. **Prior Distributions:** We use weakly informative priors to regularize predictions without imposing strong assumptions.
-
-### Implementation
-We run this model in R using the `rstanarm` package, which provides Bayesian estimation with sensible defaults.
-The default priors in `rstanarm` are chosen to be weakly informative to prevent overfitting while still regularizing our model coefficients.
-
-## Model Justification and Results
-
-By using a Bayesian linear model, we leverage historical and demographic data to provide probabilistic estimates of vote share, giving us a distribution of possible outcomes for each state.
-This approach allows for uncertainty quantification around predictions, which is particularly important in close elections.
-
-Our results are summarized in Table @tbl-modelresults, which shows the estimated vote shares, posterior means, and credible intervals for each candidate in each swing state.
-These findings indicate the influence of demographic and historical factors on vote share predictions.
-Furthermore, the model’s predictive accuracy is evaluated through posterior predictive checks, indicating how well the model captures actual voting patterns.
-The implications of these findings are significant for forecasting election outcomes, particularly in identifying which states are likely to be more competitive.
+
+### Forecasting Vote Outcomes in Key Swing States
+
+Our model aims to forecast the voting outcome of the 2024 U.S.
+presidential election, focusing specifically on eight pivotal swing states.
+We carefully selected predictor variables based on a correlation matrix we generated previously (see Figure @fig-predictors-party-pct).
+
+### Model Design and Approach
+
+We constructed two separate Bayesian generalized linear models to estimate the percentage likelihood of Trump and Harris winning in each swing state.
+Our models look at demographic and historical data that let us to generate probabilistic predictions for both candidates.
+These models provide insights into which candidate is more likely to secure a win, even in closely contested states.
+
+### Key Components of Our Models
+
+- **Outcome Variables**: The percentage of the popular vote each candidate is expected to receive.
+- **Predictor Variables**: We include factors such as education level (e.g., college graduate vs.
+high school or less), gender distribution (male vs.
+female), and ethnic demographics (White or Caucasian vs.
+Hispanic or Latino).
+Additionally, historical voting patterns (e.g., 2020 voting preferences) and current leanings toward Trump or Harris are crucial predictors.
+- **Model Structure**: Each model incorporates an intercept to represent baseline support and predictor coefficients to quantify the effect of each variable on the vote share.
+We use a Beta distribution to account for the bounded nature of percentage outcomes, ensuring our predictions remain between 0 and 100%.
+
+### Assumptions Underlying Our Models
+
+1. **Linearity**: The relationship between each predictor and vote share is assumed to be approximately linear.
+2. **Independence**: We assume that vote share predictions for each state are independent, conditional on the included predictors.
+3.
+**Non-Normality**: Given the non-normal distribution of our response variables, using a generalized linear model is more appropriate.
+Our analysis revealed that the distribution of Democratic support resembles a Poisson-like distribution, while Republican support shows multimodal characteristics.
+4. **Priors**: We applied weakly informative priors to regularize our predictions, with specific priors for those who voted for Trump and Biden in 2020, aligning with historical voting trends.
+
+### Bayesian Modeling Implementation
+
+We implemented our models using the `brms` package in R, which allows for flexible Bayesian regression modeling.
+The models were fit with four chains, running 6,000 iterations each, to ensure convergence and reliable posterior estimates.
+Sensible default priors were adjusted to reflect our prior knowledge about voting behavior, and we conducted posterior predictive checks to assess the models' performance.
+
+### Results and Interpretation
+
+The results, summarized in Table @tbl-modelresults, highlight the predicted vote shares for each candidate in the swing states.
+For each state, we report the posterior mean, as well as 95% credible intervals, to capture the range of plausible outcomes.
+Our findings underscore the impact of demographic and historical factors on voting behavior, providing a nuanced understanding of which states are likely to be battlegrounds in the upcoming election.
+
+By using a Bayesian approach, we not only make point estimates but also quantify the uncertainty surrounding these predictions.
+This is crucial for election forecasting, where even small shifts in voter behavior can have significant implications.
+Our analysis will inform campaign strategies and identify states where mobilizing specific demographics could be pivotal.
 
@@
 #| message: false
-# Select relevant columns for the prediction
+# Define the swing states of interest
 swing_states <- c("Michigan", "Georgia", "Nevada", "North Carolina", "New Hampshire", "Wisconsin", "Pennsylvania", "Arizona", "National")
@@
     `Who.did.you.vote.for.in.the.2020.election..Joe.Biden` = "Who did you vote for in the 2020 election? Joe Biden",
-    `Do.you.approve.or.disapprove.of.the.job.Joe.Biden.is.doing.as.President..Approve` = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
-    `Do.you.approve.or.disapprove.of.the.job.Joe.Biden.is.doing.as.President..Disapprove` = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove"
+    `Although.you.are.undecided..which.candidate.do.you.lean.toward..Donald.Trump` = "Although you are undecided, which candidate do you lean toward? Donald Trump",
+    `Although.you.are.undecided..which.candidate.do.you.lean.toward..Kamala.Harris` = "Although you are undecided, which candidate do you lean toward? Kamala Harris"
   )
@@
 # Generate predictions for Democratic and Republican models
-predictions_dem <- posterior_predict(bayesian_model_dem, newdata = swing_data)
-predictions_rep <- posterior_predict(bayesian_model_rep, newdata = swing_data)
+#predictions_dem <- posterior_predict(bayesian_model_dem, newdata = swing_data)
+#predictions_rep <- posterior_predict(bayesian_model_rep, newdata = swing_data)
 
 # Function to summarize predictions by state
-aggregate_summary <- function(predictions, state_labels) {
-  summary_df <- data.frame(State = character(), mean = numeric(), lower = numeric(), upper = numeric())
-  for (state in unique(state_labels)) {
-    state_indices <- which(state_labels == state)
-    state_predictions <- predictions[, state_indices]
-    state_mean <- mean(state_predictions)
-    state_lower <- quantile(state_predictions, 0.025)
-    state_upper <- quantile(state_predictions, 0.975)
-    summary_df <- rbind(summary_df, data.frame(State = state, mean = state_mean, lower = state_lower, upper = state_upper))
-  }
-  return(summary_df)
-}
+#aggregate_summary <- function(predictions, state_labels) {
+  #summary_df <- data.frame(State = character(), mean = numeric(), lower = numeric(), upper = numeric())
+  #for (state in unique(state_labels)) {
+    #state_indices <- which(state_labels == state)
+    #state_predictions <- predictions[, state_indices]
+    #state_mean <- mean(state_predictions)
+    #state_lower <- quantile(state_predictions, 0.025)
+    #state_upper <- quantile(state_predictions, 0.975)
+    #summary_df <- rbind(summary_df, data.frame(State = state, mean = state_mean, lower = state_lower, upper = state_upper))
+  #}
+  #return(summary_df)
+#}
 
 # Summarize predictions for both Democrat and Republican models
-dem_summary <- aggregate_summary(predictions_dem, swing_data$State)
-rep_summary <- aggregate_summary(predictions_rep, swing_data$State)
+#dem_summary <- aggregate_summary(predictions_dem, swing_data$State)
+#rep_summary <- aggregate_summary(predictions_rep, swing_data$State)
 
 # Combine summaries into the final prediction summary
-prediction_summary <- data.frame(
-  State = dem_summary$State,
-  DEM_Mean = dem_summary$mean,
-  DEM_Lower = dem_summary$lower,
-  DEM_Upper = dem_summary$upper,
-  REP_Mean = rep_summary$mean,
-  REP_Lower = rep_summary$lower,
-  REP_Upper = rep_summary$upper
-)
+#prediction_summary <- data.frame(
+  #State = dem_summary$State,
+  #DEM_Mean = dem_summary$mean,
+  #DEM_Lower = dem_summary$lower,
+  #DEM_Upper = dem_summary$upper,
+  #REP_Mean = rep_summary$mean,
+  #REP_Lower = rep_summary$lower,
+  #REP_Upper = rep_summary$upper
+#)
 
@@
 # Modify the prediction summary
-prediction_summary <- prediction_summary %>%
-  select(State, DEM_Mean, REP_Mean) %>%  # Keep only the mean columns
-  mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))
+#prediction_summary <- prediction_summary %>%
+  #select(State, DEM_Mean, REP_Mean) %>%  # Keep only the mean columns
+  #mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))
 
 # Create a formatted table
-kable(
-  prediction_summary,
-  col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
-  format = "latex", 
-  digits = 2,
-  escape = FALSE
-) %>%
-  kable_styling(
-    latex_options = c("striped", "hold_position", "condensed"), 
-    full_width = FALSE, 
-    font_size = 10
-)
+#kable(
+  #prediction_summary,
+  #col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
+  #format = "latex", 
+  #digits = 2,
+  #escape = FALSE
+#) %>%
+  #kable_styling(
+    #latex_options = c("striped", "hold_position", "condensed"), 
+    #full_width = FALSE, 
+    #font_size = 10
+#)
 

### 2024-11-03 scripts/01-test_simulated_data.R

@@ -1,89 +1,100 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the American Election
+# Author: Gadiel David Flores
+# Date: October 29, 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+# - The `tidyverse` package must be installed and loaded
+# - 00-simulate_data.R must have been run
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+sim_data <- read_csv(here::here("data/00-simulated_data/simulated_poll_data.csv"))
 
 # Test if the data was successfully loaded
-if (exists("analysis_data")) {
+if (exists("sim_data")) {
   message("Test Passed: The dataset was successfully loaded.")
 } else {
   stop("Test Failed: The dataset could not be loaded.")
 }
 
-
 #### Test data ####
-
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if all column headers in 'sim_data' are unique
+if (length(unique(names(sim_data))) == length(names(sim_data))) {
+  message("Test Passed: All column headers in 'sim_data' are unique.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: There are duplicate column headers in 'sim_data'.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
+# Check if the 'state' column contains only valid American state names
+valid_states <- c(
+  "Alabama", "Alaska", "Arizona", "Arkansas", "California", 
+  "Colorado", "Connecticut", "Delaware", "Florida", "Georgia", 
+  "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", 
+  "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", 
[…]

### 2024-11-03 scripts/03-1-clean_data.R

@@ -1,0 +1,65 @@
+#### Preamble ####
+# Purpose: Cleans the raw presidential general election polls data into usable links
+# an analysis dataset
+# Author: Tina Kim
+# Date: 7 October 2024
+# Contact: tinak.kim@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean Raw data ####
+raw_data <- read_csv("data/01-raw_data/raw_data.csv")
+
+# Find the top 3 most frequent pollster in the raw data
+top_pollsters <- raw_data %>%
+  count(pollster, sort = TRUE) %>%
+  top_n(3, n) %>%
+  pull(pollster)
+
+cleaned_data <- raw_data %>%
+  # Filter the raw data to only top 3 most frequent pollster names
+  filter(pollster %in% top_pollsters) %>%
+  # Remove columns with completely empty rows
+  select(where(~ any(!is.na(.)))) 
+
+# Calculate missing values for each of the three pollsters
+missing_values_summary <- cleaned_data %>%
+  group_by(pollster) %>%
+  summarize(
+    # Missing number of variables across all columns for each pollster
+    missing_count = sum(is.na(across(everything()))),
+    # Total number of variables for each pollster
+    total_count = n() * ncol(cleaned_data),
+    .groups = 'drop'
+  ) %>%
+  # (Missing # of variables / total # of variables) ratio for each pollster
+  mutate(missing_ratio = missing_count / total_count) %>%
+  # Order from smallest to biggest ratio value
+  arrange(missing_ratio)
+
+# Find the pollster with the lowest missing values ratio because it has
+# the most complete data which could lead to a more reliable analysis
+smallest_ratio_pollster <- missing_values_summary %>%
+  slice(1) %>%
+  pull(pollster)
+
+# Filter cleaned data to include only the smallest ratio pollster.
+# This pollster is a good balance of high number of observations and a low
+# number of missing values.
+cleaned_data <- cleaned_data %>%
+  filter(pollster == smallest_ratio_pollster)
+
+# Convert 'start_date' to Date format to match 'end_date' format and filter out rows before July 21st, 2024
+cleaned_data <- cleaned_data %>%
+  mutate(
+    start_date = as.Date(start_date, format = "%m/%d/%y"),
[…]

### 2024-11-03 scripts/03-2-clean-links-state-data.R

@@ -1,0 +1,48 @@
+#### Preamble ####
+# Purpose: Organizes states poll links
+# an analysis dataset
+# Author: Gadiel David Flores Jimenez
+# Date: 7 October 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to cleaned poll data
+# Any other information needed? Intall googledrive and googleAuthR
+library(tidyverse)
+
+# Load the poll links data
+poll_links_data <- read_csv("data/02-analysis_data/pollster_data/emerson_list_data.csv")
+
+#### Pivot for URL crosstabs ####
+url_pivot <- poll_links_data %>%
+  group_by(state) %>%
+  mutate(url_rank = row_number()) %>%
+  select(state, url_crosstab, url_rank) %>%
+  pivot_wider(
+    names_from = url_rank,
+    values_from = url_crosstab,
+    names_prefix = "url_crosstab_"
+  ) %>%
+  ungroup()
+
+#### Shift non-NA values to the left in relevant columns ####
+# Define columns to shift
+columns_to_shift <- names(url_pivot)[!(names(url_pivot) %in% c("state"))]
+
+# Function to shift values to the left
+shift_left <- function(row) {
+  non_na_values <- row[!is.na(row)]  # Keep only non-NA values
+  c(non_na_values, rep(NA, length(row) - length(non_na_values)))  # Fill the remaining with NA
+}
+
+# Apply shift_left function
+new_poll_data_clean <- url_pivot %>%
+  mutate(across(all_of(columns_to_shift), ~ shift_left(.)))
+
+#### Remove columns that are NA ####
+new_poll_data_clean <- new_poll_data_clean %>%
+  select(where(~ any(!is.na(.)))) 
+
+#### Save the cleaned dataset ####
+write_csv(new_poll_data_clean, "data/02-analysis_data/pollster_data/pollster_link_data.csv")
+
+

### 2024-11-03 scripts/03-3-api_extract_clean_data.R

@@ -1,0 +1,455 @@
+#### Preamble ####
+# Purpose: Extracts data from links into a summarize dataset
+# an analysis dataset
+# Author: Gadiel David Flores
+# Date: 7 October 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to cleaned poll data
+# Any other information needed? Intall googledrive and googleAuthR
+
+#### Load Libraries and files ####
+library(googledrive)
+library(readxl)
+library(tidyverse)
+library(dplyr)
+library(dplyr)
+library(lubridate)
+library(readr)
+library(janitor)
+library(arrow)
+state_data <- read_csv("data/02-analysis_data/pollster_data/pollster_link_data.csv")
+#### Helper Functions ####
+
+# Function to check if a string is a Google Sheets link
+is_google_sheet_link <- function(text) {
+  grepl("docs.google.com/spreadsheets", text)
+}
+# Function to check if a string is an Excel file link
+is_excel_link <- function(text) {
+  grepl("\\.xlsx$|\\.xls$|view.officeapps.live.com/op/view.aspx\\?src=.*\\.xlsx", text)
+}
+# Function to download and read a Google Sheet
+download_and_read_google_sheet <- function(url) {
+  # Download the file from the Google Sheets URL
+  file <- drive_download(as_id(url), path = "temp_file.xlsx", overwrite = TRUE)
+  # Read the file
+  data <- read_excel(file$local_path)
+  # Return the data
+  return(data)
+}
+
+# Function to replace NA values with the previous non-NA value
+replace_na_with_previous <- function(df, column_index) {
+  last_non_na <- NA  # Variable to store the last non-NA value
+  
+  for (i in 1:nrow(df)) {
+    if (is.na(df[[column_index]][i])) {
+      df[[column_index]][i] <- last_non_na  # Replace NA with last non-NA value
+    } else {
+      last_non_na <- df[[column_index]][i]  # Update last non-NA value
+    }
+  }
+  
+  return(df)
+}
+
+
+#### Main Function to Process Google and Excel Sheets####
+process_google_sheet <- function(sheet_url, output_path) {
+  # Step 1: Download and read the Google Sheet
+  sheet_data <- download_and_read_google_sheet(sheet_url)
+  
+  # Step 2: Define unwanted values to clean
[…]

### 2024-11-03 scripts/03-api-extract-clean_data.R

@@ -1,441 +1,0 @@
-#### Preamble ####
-# Purpose: Extracts data from links into a summarize dataset
-# an analysis dataset
-# Author: Gadiel David Flores
-# Date: 7 October 2024
-# Contact: davidgadiel.flores@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Need to cleaned poll data
-# Any other information needed? Intall googledrive and googleAuthR
-
-#### Load Libraries and files ####
-library(googledrive)
-library(readxl)
-library(tidyverse)
-library(dplyr)
-library(dplyr)
-library(lubridate)
-library(readr)
-library(janitor)
-library(arrow)
-state_data <- read_csv("data/02-analysis_data/pollster_data/pollster_link_data.csv")
-#### Helper Functions ####
-
-# Function to check if a string is a Google Sheets link
-is_google_sheet_link <- function(text) {
-  grepl("docs.google.com/spreadsheets", text)
-}
-# Function to check if a string is an Excel file link
-is_excel_link <- function(text) {
-  grepl("\\.xlsx$|\\.xls$|view.officeapps.live.com/op/view.aspx\\?src=.*\\.xlsx", text)
-}
-# Function to download and read a Google Sheet
-download_and_read_google_sheet <- function(url) {
-  # Download the file from the Google Sheets URL
-  file <- drive_download(as_id(url), path = "temp_file.xlsx", overwrite = TRUE)
-  # Read the file
-  data <- read_excel(file$local_path)
-  # Return the data
-  return(data)
-}
-
-# Function to replace NA values with the previous non-NA value
-replace_na_with_previous <- function(df, column_index) {
-  last_non_na <- NA  # Variable to store the last non-NA value
-  
-  for (i in 1:nrow(df)) {
-    if (is.na(df[[column_index]][i])) {
-      df[[column_index]][i] <- last_non_na  # Replace NA with last non-NA value
-    } else {
-      last_non_na <- df[[column_index]][i]  # Update last non-NA value
-    }
-  }
-  
-  return(df)
-}
-
-
-#### Main Function to Process Google and Excel Sheets####
-process_google_sheet <- function(sheet_url, output_path) {
-  # Step 1: Download and read the Google Sheet
-  sheet_data <- download_and_read_google_sheet(sheet_url)
-  
-  # Step 2: Define unwanted values to clean
[…]

### 2024-11-03 scripts/03-clean-links-state-data.R

@@ -1,48 +1,0 @@
-#### Preamble ####
-# Purpose: Organizes states poll links
-# an analysis dataset
-# Author: Gadiel David Flores Jimenez
-# Date: 7 October 2024
-# Contact: davidgadiel.flores@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Need to cleaned poll data
-# Any other information needed? Intall googledrive and googleAuthR
-library(tidyverse)
-
-# Load the poll links data
-poll_links_data <- read_csv("data/02-analysis_data/pollster_data/emerson_list_data.csv")
-
-#### Pivot for URL crosstabs ####
-url_pivot <- poll_links_data %>%
-  group_by(state) %>%
-  mutate(url_rank = row_number()) %>%
-  select(state, url_crosstab, url_rank) %>%
-  pivot_wider(
-    names_from = url_rank,
-    values_from = url_crosstab,
-    names_prefix = "url_crosstab_"
-  ) %>%
-  ungroup()
-
-#### Shift non-NA values to the left in relevant columns ####
-# Define columns to shift
-columns_to_shift <- names(url_pivot)[!(names(url_pivot) %in% c("state"))]
-
-# Function to shift values to the left
-shift_left <- function(row) {
-  non_na_values <- row[!is.na(row)]  # Keep only non-NA values
-  c(non_na_values, rep(NA, length(row) - length(non_na_values)))  # Fill the remaining with NA
-}
-
-# Apply shift_left function
-new_poll_data_clean <- url_pivot %>%
-  mutate(across(all_of(columns_to_shift), ~ shift_left(.)))
-
-#### Remove columns that are NA ####
-new_poll_data_clean <- new_poll_data_clean %>%
-  select(where(~ any(!is.na(.)))) 
-
-#### Save the cleaned dataset ####
-write_csv(new_poll_data_clean, "data/02-analysis_data/pollster_data/pollster_link_data.csv")
-
-

### 2024-11-03 scripts/03-clean_data.R

@@ -1,57 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw presidential general election polls data into usable links
-# an analysis dataset
-# Author: Tina Kim
-# Date: 7 October 2024
-# Contact: tinak.kim@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Need to have downloaded the data
-# Any other information needed? None
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Clean Raw data ####
-raw_data <- read_csv("data/01-raw_data/raw_data.csv")
-
-# Find the top 3 most frequent pollster in the raw data
-top_pollsters <- raw_data %>%
-  count(pollster, sort = TRUE) %>%
-  top_n(3, n) %>%
-  pull(pollster)
-
-cleaned_data <- raw_data %>%
-  # Filter the raw data to only top 3 most frequent pollster names
-  filter(pollster %in% top_pollsters) %>%
-  # Remove columns with completely empty rows
-  select(where(~ any(!is.na(.)))) 
-
-# Calculate missing values for each of the three pollsters
-missing_values_summary <- cleaned_data %>%
-  group_by(pollster) %>%
-  summarize(
-    # Missing number of variables across all columns for each pollster
-    missing_count = sum(is.na(across(everything()))),
-    # Total number of variables for each pollster
-    total_count = n() * ncol(cleaned_data),
-    .groups = 'drop'
-  ) %>%
-  # (Missing # of variables / total # of variables) ratio for each pollster
-  mutate(missing_ratio = missing_count / total_count) %>%
-  # Order from smallest to biggest ratio value
-  arrange(missing_ratio)
-
-# Find the pollster with the lowest missing values ratio because it has
-# the most complete data which could lead to a more reliable analysis
-smallest_ratio_pollster <- missing_values_summary %>%
-  slice(1) %>%
-  pull(pollster)
-
-# Filter cleaned data to include only the smallest ratio pollster.
-# This pollster is a good balance of high number of observations and a low
-# number of missing values.
-cleaned_data <- cleaned_data %>%
-  filter(pollster == smallest_ratio_pollster)
-
-#### Save data ####
-write_csv(cleaned_data, "data/02-analysis_data/pollster_data/emerson_list_data.csv")

### 2024-11-03 scripts/03-merge_datasets.R

@@ -1,159 +1,0 @@
-#### Preamble ####
-# Purpose: Extracts data from links into a summarize dataset
-# an analysis dataset
-# Author: Gadiel David Flores
-# Date: 7 October 2024
-# Contact: davidgadiel.flores@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Need to cleaned poll data
-# Any other information needed? Intall googledrive and googleAuthR
-
-# Load necessary libraries
-library(dplyr)
-library(lubridate)
-library(readr)
-library(janitor)
-library(arrow)
-
-#### Merge datasets ####
-# Read and clean the data
-data <- read_csv("data/02-analysis_data/pollster_data/emerson_list_data.csv") |>
-  clean_names()
-
-# Clean the data (treat NA state as National, convert dates to date format)
-cleaned_data <- data |>
-  mutate(
-    state = if_else(is.na(state), "National", state), # Treat NA as National
-    end_date = mdy(end_date), # Convert end_date to date format
-    start_date = mdy(start_date) # Convert start_date to date format if needed
-  ) |>
-  mutate(
-    num_respondents = round((pct / 100) * sample_size, 0) # Convert percentage to number (applies to any candidate)
-  )
-
-# Filter to include only Democrats (DEM) and Republicans (REP)
-filtered_data <- cleaned_data |>
-  filter(party %in% c("DEM", "REP"))
-
-# Extract relevant information for each unique state and party combination
-state_summary <- filtered_data |>
-  group_by(state, party) |>
-  filter(end_date == max(end_date, na.rm = TRUE)) |>
-  summarize(
-    start_date = min(start_date, na.rm = TRUE), # Earliest start date for each state and party
-    end_date = max(end_date, na.rm = TRUE), # Latest end date for each state and party
-    total_sample_size = sum(sample_size, na.rm = TRUE), # Total sample size for each state and party
-    pct = first(pct), # Use the pct value associated with the latest entry
-    num_respondents = sum(num_respondents, na.rm = TRUE) # Sum of number of respondents for each state and party
-  ) |>
-  ungroup()
-
-# Read existing cleaned dataset
-cleaned_dataset <- read_csv("data/02-analysis_data/pollster_data/summarized_questions_data.csv")
-
-# Create separate columns for Democrats and Republicans
-pivoted_data <- state_summary %>%
-  select(state, party, start_date, end_date, total_sample_size, pct, num_respondents) %>%
[…]

### 2024-11-03 scripts/06-model_data.R

@@ -5,18 +5,18 @@
 # Contact: davidgadiel.flores@mail.utoronto.ca
 # License: MIT
 
-#### Workspace setup ####
+#### Updated Workspace Setup ####
 library(here)
 library(arrow)
 library(dplyr)
-library(brms)     
+library(brms)
 
 # Data Import
 analysis_data <- read_parquet(here::here("data/02-analysis_data/timeline_data.parquet"))
 
 #### Model Setup for Democrat Voters ####
 
-# Define column references 
+# Define updated predictor columns
 predictor_columns <- c(
   "What is the highest level of education you have attained? College graduate",
   "What is the highest level of education you have attained? High school or less",
@@ -24,10 +24,10 @@ predictor_columns <- c(
   "Can you please tell me your gender? Male",
   "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
   "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
+  "Although you are undecided, which candidate do you lean toward? Donald Trump",
+  "Although you are undecided, which candidate do you lean toward? Kamala Harris",
   "Who did you vote for in the 2020 election? Donald Trump",
-  "Who did you vote for in the 2020 election? Joe Biden",
-  "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
-  "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove"
+  "Who did you vote for in the 2020 election? Joe Biden"
 )
 
 # Filter analysis_data to only include rows where party is DEM
@@ -56,10 +56,9 @@ model_data_dem <- analysis_data_dem %>%
 predictor_formula <- paste(predictor_columns, collapse = " + ")
 formula <- as.formula(paste("pct_scaled ~", predictor_formula))
 
-# Set priors with specific priors for `Who did you vote for...` columns
-priors <- c(
+# Set priors with a specific prior for `Who did you vote for in the 2020 election? Joe Biden`
+priors_dem <- c(
   prior(normal(0, 5), class = "b"),
-  prior(normal(0, 1), class = "b", coef = "Who.did.you.vote.for.in.the.2020.election..Donald.Trump"),
[…]

### 2024-11-04 paper/paper.qmd
cleaned data

@@
 
+poll_data <- poll_data %>%
+  mutate(across(12:ncol(.), ~ round(as.numeric(.), 1)))
+
+time_data <- time_data %>%
+  mutate(across(5:ncol(.), ~ round(as.numeric(.), 1)))
+
 ```
@@
     past_vote_trump = mean(`Who did you vote for in the 2020 election? Donald Trump`, na.rm = TRUE)
-  ) %>%
-  mutate(across(everything(), ~ replace_na(., 0))) %>%  # Replace NA values with 0
-  mutate(across(where(is.numeric), ~ round(., 1)))  # Round to the nearest tenth
+  )
 
@@
 #| label: fig-party-extremes
-#| fig-cap: Demographic Differences Between Maryland and South Dakota
+#| fig-cap: Demographic Differences Between California and South Dakota
 #| echo: false
@@
   select(-contains('Mode')) %>%             # Remove columns containing 'Mode' in their names
-  select(-contains('registration'))
+  select(-contains('registration')) %>%
+  select(-democrat_pct, -republican_pct)    # Exclude 'democrat_pct' and 'republican_pct'
 
@@
     ) %>%
-    mutate(Group = recode(Group, "Blue_Mean" = "Maryland", "Red_Mean" = "South Dakota"))
+    mutate(Group = recode(Group, "Blue_Mean" = "California", "Red_Mean" = "South Dakota"))
 }
@@
     # Update the colors and labels for the groups
-    scale_fill_manual(values = c("Maryland" = "blue", "South Dakota" = "red")) +
+    scale_fill_manual(values = c("California" = "blue", "South Dakota" = "red")) +
     labs(
-      title = "Demographic Differences Between Maryland and South Dakota",
+      title = "Demographic Differences Between California and South Dakota",
       x = "Demographic Variable",
@@
 # Main analysis function
-analyze_state_differences <- function(data, blue_state, red_state, top_n_vars = 4) {
+analyze_state_differences <- function(data, blue_state, red_state, top_n_vars = 5) {
   diff_df <- compute_mean_differences(data, blue_state, red_state)
@@
 
-# Run the analysis for Maryland and South Dakota
-analysis_results <- analyze_state_differences(poll_summary, blue_state = "Maryland", red_state = "South Dakota", top_n_vars = 5)
+# Run the analysis for California and South Dakota
+analysis_results <- analyze_state_differences(poll_summary, blue_state = "California", red_state = "South Dakota", top_n_vars = 6)
 
@@
 analysis_results$plot
-
 ```
@@
     male_pct = round(mean(.data[[column_refs$male]], na.rm = TRUE), 1),
-    lean_trump_pct = round(mean(.data[[column_refs$lean_trump]], na.rm = TRUE), 1),
-    lean_kamala_pct = round(mean(.data[[column_refs$lean_Kamala]], na.rm = TRUE), 1),
     caucasian_pct = round(mean(.data[[column_refs$caucasian]], na.rm = TRUE), 1),
@@
 
-# Select relevant numeric columns for correlation matrix, excluding PartyDifference
+# Select relevant numeric columns for correlation matrix, excluding lean_trump and lean_kamala
 corr_data <- poll_summary %>%
@@
     Male = male_pct,
-    LeanTrump = lean_trump_pct,
-    LeanKamala = lean_kamala_pct,
     Caucasian = caucasian_pct,
@@
   theme(
-    plot.title = element_text(size = 12, hjust = 0.5),  # Title size and centered
+    plot.title = element_text(size = 10, hjust = 0.5),  # Title size and centered
     axis.text = element_text(size = 1),  # Axis text size
@@
 # Model {#sec-model}
-## Model Development
-
-### Forecasting Vote Outcomes in Key Swing States
+
+## Forecasting Vote Outcomes in Key Swing States
 
@@
 
-### Model Design and Approach
+## Model Design and Approach
 
@@
 
-### Key Components of Our Models
+## Key Components of Our Models
 
@@
 
-### Assumptions Underlying Our Models
+## Assumptions Underlying Our Models
 
@@
 
-### Bayesian Modeling Implementation
+## Bayesian Modeling Implementation
 
@@
 
-### Results and Interpretation
+## Results and Interpretation
 
@@
 # Generate predictions for Democratic and Republican models
-#predictions_dem <- posterior_predict(bayesian_model_dem, newdata = swing_data)
-#predictions_rep <- posterior_predict(bayesian_model_rep, newdata = swing_data)
+predictions_dem <- posterior_predict(bayesian_model_dem, newdata = swing_data)
+predictions_rep <- posterior_predict(bayesian_model_rep, newdata = swing_data)
 
 # Function to summarize predictions by state
-#aggregate_summary <- function(predictions, state_labels) {
-  #summary_df <- data.frame(State = character(), mean = numeric(), lower = numeric(), upper = numeric())
-  #for (state in unique(state_labels)) {
-    #state_indices <- which(state_labels == state)
-    #state_predictions <- predictions[, state_indices]
-    #state_mean <- mean(state_predictions)
-    #state_lower <- quantile(state_predictions, 0.025)
-    #state_upper <- quantile(state_predictions, 0.975)
-    #summary_df <- rbind(summary_df, data.frame(State = state, mean = state_mean, lower = state_lower, upper = state_upper))
-  #}
-  #return(summary_df)
-#}
+aggregate_summary <- function(predictions, state_labels) {
+  summary_df <- data.frame(State = character(), mean = numeric(), lower = numeric(), upper = numeric())
+  for (state in unique(state_labels)) {
+    state_indices <- which(state_labels == state)
+    state_predictions <- predictions[, state_indices]
+    state_mean <- mean(state_predictions)
+    state_lower <- quantile(state_predictions, 0.025)
+    state_upper <- quantile(state_predictions, 0.975)
+    summary_df <- rbind(summary_df, data.frame(State = state, mean = state_mean, lower = state_lower, upper = state_upper))
+  }
+  return(summary_df)
+}
 
 # Summarize predictions for both Democrat and Republican models
-#dem_summary <- aggregate_summary(predictions_dem, swing_data$State)
-#rep_summary <- aggregate_summary(predictions_rep, swing_data$State)
+dem_summary <- aggregate_summary(predictions_dem, swing_data$State)
+rep_summary <- aggregate_summary(predictions_rep, swing_data$State)
 
 # Combine summaries into the final prediction summary
-#prediction_summary <- data.frame(
-  #State = dem_summary$State,
-  #DEM_Mean = dem_summary$mean,
-  #DEM_Lower = dem_summary$lower,
-  #DEM_Upper = dem_summary$upper,
-  #REP_Mean = rep_summary$mean,
-  #REP_Lower = rep_summary$lower,
-  #REP_Upper = rep_summary$upper
-#)
-
+prediction_summary <- data.frame(
+  State = dem_summary$State,
+  DEM_Mean = dem_summary$mean,
+  DEM_Lower = dem_summary$lower,
+  DEM_Upper = dem_summary$upper,
+  REP_Mean = rep_summary$mean,
+  REP_Lower = rep_summary$lower,
+  REP_Upper = rep_summary$upper
+)
 ```
@@
 # Modify the prediction summary
-#prediction_summary <- prediction_summary %>%
-  #select(State, DEM_Mean, REP_Mean) %>%  # Keep only the mean columns
-  #mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))
+prediction_summary <- prediction_summary %>%
+  select(State, DEM_Mean, REP_Mean) %>%  # Keep only the mean columns
+  mutate(Winner = ifelse(DEM_Mean > REP_Mean, "Democrat", "Republican"))
 
 # Create a formatted table
-#kable(
-  #prediction_summary,
-  #col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
-  #format = "latex", 
-  #digits = 2,
-  #escape = FALSE
-#) %>%
-  #kable_styling(
-    #latex_options = c("striped", "hold_position", "condensed"), 
-    #full_width = FALSE, 
-    #font_size = 10
-#)
+kable(
+  prediction_summary,
+  col.names = c("State", "Democrat Points", "Republican Points", "Winner"),
+  format = "latex", 
+  digits = 2,
+  escape = FALSE
+) %>%
+  kable_styling(
+    latex_options = c("striped", "hold_position", "condensed"), 
+    full_width = FALSE, 
+    font_size = 10
+  )
 
@@
 
-## Additional Data & Model Details
+# Additional data details
 Include any technical details on data cleaning, model diagnostics, and posterior checks.
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-```
-
-\newpage
-
 # References

### 2024-11-04 scripts/03-3-api_extract_clean_data.R

@@ -364,12 +364,6 @@ updated_dataset <- cleaned_dataset %>%
 updated_dataset <- updated_dataset %>%
   select(state, starts_with("democrat"), starts_with("republican"), everything())
 
-updated_dataset <- updated_dataset %>%
-  mutate(across(-c(1:11), ~ if (is.character(.)) as.numeric(.) else .))
-
-updated_dataset <- updated_dataset_1 %>%
-  mutate(across(where(is.numeric), ~ round(., 1)))
-
 # Save dataset
 write_parquet(updated_dataset, "data/02-analysis_data/summarized_state_poll_data.parquet")
 
@@ -448,8 +442,5 @@ for (i in seq_len(nrow(state_data))) {
   # Append new_row as a row to timeline_data
   timeline_data <- rbind(timeline_data, as.data.frame(new_row, stringsAsFactors = FALSE, check.names = FALSE))
 }
-
-timeline_data <- timeline_data %>%
-  mutate(across(where(is.numeric), ~ round(., 1)))
 
 write_parquet(timeline_data, "data/02-analysis_data/timeline_data.parquet")

### 2024-11-04 scripts/06-model_data.R

@@ -56,14 +56,14 @@ model_data_dem <- analysis_data_dem %>%
 predictor_formula <- paste(predictor_columns, collapse = " + ")
 formula <- as.formula(paste("pct_scaled ~", predictor_formula))
 
-# Set priors with a specific prior for `Who did you vote for in the 2020 election? Joe Biden`
+# Adjusted priors for the Democrat Model
 priors_dem <- c(
   prior(normal(0, 5), class = "b"),
   prior(normal(0, 1), class = "b", coef = "Who.did.you.vote.for.in.the.2020.election..Joe.Biden"),
-  prior(beta(1, 1), class = "phi")  # Prior for precision parameter
+  prior(gamma(2, 0.1), class = "phi")  # Use gamma prior for precision parameter
 )
 
-# Fit the Democrat Model
+# Fit the Democrat Model with updated control parameters
 dem_model <- brm(
   formula = formula,
   data = model_data_dem,
@@ -74,7 +74,7 @@ dem_model <- brm(
   warmup = 2000,
   cores = 4,
   seed = 123,
-  control = list(adapt_delta = 0.9, max_treedepth = 15)
+  control = list(adapt_delta = 0.9999, max_treedepth = 20)
 )
 
 ### Save the Democrat Model ###
@@ -105,14 +105,14 @@ model_data_rep <- analysis_data_rep %>%
 
 ### Specify and Fit the Republican Model ###
 
-# Set priors with a specific prior for `Who did you vote for in the 2020 election? Donald Trump`
+# Adjusted priors for the Republican Model
 priors_rep <- c(
   prior(normal(0, 5), class = "b"),
   prior(normal(0, 1), class = "b", coef = "Who.did.you.vote.for.in.the.2020.election..Donald.Trump"),
-  prior(beta(1, 1), class = "phi")  # Prior for precision parameter
+  prior(gamma(2, 0.1), class = "phi")  # Use gamma prior for precision parameter
 )
 
-# Fit the Republican Model
+# Fit the Republican Model with updated control parameters
 rep_model <- brm(
   formula = formula,
   data = model_data_rep,
@@ -123,7 +123,7 @@ rep_model <- brm(
   warmup = 2000,
   cores = 4,
   seed = 123,
-  control = list(adapt_delta = 0.9, max_treedepth = 15)
+  control = list(adapt_delta = 0.9999, max_treedepth = 20)
 )
 
 ### Save the Republican Model ###

### 2024-11-04 paper/paper.qmd
Fixes some prose error in Appendix A

@@
 ## Appendix A: Pollster Methodology Overview {-}
-Emerson College Polling was initially a college polling exercise, and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs).
+Emerson College Polling was initially a college polling exercise and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs).
 For the 2024 USA Presidential Election pollster, the Emerson College Polling defined the population to be "likely voters" of 2024 United States Presidential Election.
@@
 In the MMS-to-Online approach, the target population were sent text messages with graphics that invite them to take a "screening questionnaire", and those who pass the questionnaire can move on to take the survey.
-The selected respondents were based on "state voter files provided by Aristotle." The Online Opt-in Panel approach involves respondents from the targeted population being invited to finish the survey by means of an online opt-in panel "provided by CINT" (@AboutUs).
+The selected respondents were based on state voter files "provided by Aristotle." The Online Opt-in Panel approach involves respondents from the targeted population being invited to finish the survey by means of an online opt-in panel "provided by CINT" (@AboutUs).
 Specifically, in the Online Panel approach, selected voters were pre-matched to "L2 voter file data provided by Rep Data" (@Polling).
@@
 Therefore, after analyzing the advantages and disadvantages of this pollster about the 2024 United States Presidential Elections by Emerson College Polling, we can conclude that Emerson College Polling can improve their pollster by implementing the following approaches.
-Emerson can consider increasing the sample size in order to make their survey more representative of the targeted population, explicitly describing their definition of the targeted population "likely voters" and the model used to determine characteristics of the targeted population.
-Emerson can also consider changing from simple random to stratified sampling, based on different states or other demographical features based on their targeted population model, making their sample more representative of the population, decrease the length of the questionnaire, and also explicitly explain their treatment of non-response, which will increase the validity of the study.
+Emerson can consider increasing the sample size to make their survey more representative of the targeted population, explicitly describing their definition of the targeted population "likely voters" and the model used to determine characteristics of the targeted population.
+Emerson can also consider changing from simple random to stratified sampling, based on different states or other demographic features based on their targeted population model, making their sample more representative of the population, decrease the length of the questionnaire, and also explicitly explain their treatment of non-response, which will increase the validity of the study.
 Overall, this pollster is still highly valid, and generally reliable for predicting the 2024 United States Presidential Election.

### 2024-11-04 paper/paper.qmd
Updated the abstract

@@
 date-format: long
-abstract: Our analysis of swing state polls for the 2024 U.S.
-Presidential Election shows a close race.
-Republicans lead in most key states like Michigan, Georgia, Nevada, Wisconsin, and Pennsylvania, while Democrats have a slight edge in Arizona.
-The national average also leans Republican, suggesting a Republican win in the 2024 Election.
+abstract: This paper analyzes recent swing state polling data to forecast outcomes for the 2024 U.S.
+Presidential Election.
+Our findings indicate a close race, with Republicans leading in key states such as Michigan, Georgia, Nevada, Wisconsin, and Pennsylvania, while Democrats hold a slight edge in Arizona.
+The national polling average also leans towards Republican, suggesting a Republican win in the 2024 Election.
+These insights give us a clearer picture of voter preferences across states and highlight the importance of targeted campaign strategies.
 format: pdf

### 2024-11-04 paper/paper.qmd
Removed redundant libraries already part of tidyverse

@@
 library(tidyverse)
-library(dplyr)
 library(knitr)
-library(lubridate)
-library(ggplot2)
 library(ggcorrplot)

### 2024-11-04 paper/paper.qmd
Updated the introduction section

@@
 
-*Overview paragraph:* The 2024 United States of America Presidential Election is to elect the fourty-seventh president of USA, with the two candidates as Donald Trump, the fourty-fifth president of the United States and representative of the Republican Party, and Kamala Harris, current vice president of the United States and representative of the Democratic Party, after Joe Biden announced that he will not continue participate in this presidential election anymore (@BidenQuits).
-Democracy can be described as liberal and left-leaning, with the emphasis of economy intervention with implementation of policies such as minimum wage and progressive tax rates.
+The 2024 United States Presidential Election will determine the next president of the United States of America, with the outcome likely to affect both domestic and international policy and economic trends.
+Former President Donald Trump, the 45th president of the United States, is representing the Republican Party.
+In contrast, Vice President Kamala Harris is now the Democratic Party's candidate in the 2024 U.S.
+Presidential Election, following President Joe Biden's decision to withdraw from the race (@BidenQuits).
+Democracy can be described as liberal and left-leaning, with the emphasis on economy intervention with implementation of policies such as minimum wage and progressive tax rates.
 Members of the Republican party tend to follow the free economy, and can be described as right-leaning (@RepDem).
-The results of the 2024 Presidential Election will determine which political party, Republican or Democratic, will become the leader of the United States for the next four years, largely affecting the policies implemented in the United States during this time period.
-Due to the strong nation power of the United States, this may even affect the world's economic trends in the four years.
-Therefore, predicting results of the upcoming 2024 United States Presidential Election is a necessary action.
-
-*Estimand paragraph:* To predict results of the 2024 United States of America Presidential Election, our aim is clearly to conclude which candidate, either Harris or Trump, is more likely to win this election.
-Our response variable is in turn the probability of one candidate winning the election based on poll data, with our chosen poll as Emerson.
-When the supporting percentage of one candidate, or the party s/he represents, is only slightly less than 50%, we cannot clearly determine whether s/he would be predicted to have been elected as the fourty-seventh US President or not, since there is a group of people who did not choose to participate in the voting process.
-Therefore, to increase accuracy of predictions, we will build two models, one model with the estimand as the probability of Harris, or the Democratic party, winning the election, and another with the estimand as the probability of Trump, or the Republican party, winning the election.
-By comparing the two predictions, one can get a final prediction of who is more likely to be elected as the next President of the United States of America.
+
+The outcome of the 2024 Presidential Election will determine whether the Republican or Democratic Party will lead the United States for the next four years.
+Given the United States' significant global influence, accurately predicting the results of this election carries substantial implications for shaping policy direction, ensuring economic stability, and influencing international relations.
+Reliable predictions enable parties to refine their campaigns and mobilize voters, influencing turnout and perceptions of competitiveness.
+Additionally, these predictions provide stakeholders with valuable insights that help adjust strategies and facilitate discussions on pressing societal issues.
+Ultimately, effective election forecasting is essential for understanding governance and its impact on society.
+
+This paper addresses a gap in election forecasting by using recent polling data from key swing states to provide a prediction of the 2024 election outcome.
+We aim to determine which candidate—Kamala Harris or Donald Trump—is more likely to win this election.
+Our response variable is in turn the probability of one candidate winning the election based on our chosen Emerson poll data.
+When one candidate’s support hovers just below 50%, it becomes challenging to predict their election as the 47th President of the United States, especially considering the potential influence of non-participating voters.
+To increase the accuracy of our predictions, we will develop two models: one estimating the probability of Harris, representing the Democratic Party, winning the election, and the other estimating the probability of Trump, representing the Republican Party, winning.
+By comparing these predictions, we aim to arrive at a clearer conclusion about who is more likely to be elected as the next President of the United States.
 
@@
 
-*Why it matters paragraph:* Explain the broader significance of accurately forecasting election results for politics, society, and policymaking.
-
-*Telegraphing paragraph:* 
-The remainder of this paper is structured as follows: Section 2 details the data and measurement process; Section 3 covers model development and results; Section 4 discusses implications and future steps.
-The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured as follows: @sec-data details the data and measurement process; @sec-results covers model development and results; @sec-discussion discusses implications and future steps.
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2024-11-04 paper/paper.qmd
Updated introduction section and pdf file

@@
 
-The remainder of this paper is structured as follows: @sec-data details the data and measurement process; @sec-results covers model development and results; @sec-discussion discusses implications and future steps.
+The remainder of this paper is structured as follows: @sec-data details the data and measurement process; @sec-model covers the model; @sec-results provides results; and @sec-discussion discusses implications and future steps.
 

### 2024-11-04 paper/paper.qmd
Updated models

@@
 The table highlights key trends and differences across states in terms of party lean and demographic distribution.
-@fig-predictors-pct-summary-1 and @fig-predictors-pct-summary-2 provides a snapshot of these findings.
-
-```{r}
-#| label: fig-predictors-pct-summary
-#| tbl-cap: "Latest Poll Points by Party"
+@tbl-predictors-pct-summary-1 and @tbl-predictors-pct-summary-2 provides a snapshot of these findings.
+
+```{r}
+#| label: tbl-predictors-pct-summary-1
+#| tbl-cap: "Voter Party Preferences by State"
 #| echo: false
@@
 
+# Create summary table
+poll_summary_party <- poll_summary %>%
+  select(State, `Democrat Pct`, `Republican Pct`, `Voted Biden 2020`, `Voted Trump 2020`)
+
 # Display table for party percentages
@@
 
+```
+
+```{r}
+#| label: tbl-predictors-pct-summary-2
+#| tbl-cap: "Demographic Breakdowns by State"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create summary table
+poll_summary_demographics <- poll_summary %>%
+  select(State, `High Educated`, `Low Educated`, `Female`, 
+         `Male`, `Nonbinary`, `Caucasian`, `Minority Ethnicity`)
+
 # Display table for demographic variables
@@
 ## Measurement  
-We chose Emerson as our primary pollster, and the reasoning for this is detailed in @sec-appenA.
+We chose Emerson as our primary pollster, and the reasoning for this is detailed in @sec-appendix.
 Emerson's polls offered crucial information, such as polling dates, party affiliations, sample sizes, and the percentage of support for each political party.
@@
 #| label: tbl-final-summary
+#| tbl-cap: "Final Summary Table of 95% Confidence Intervals for Common States"
 #| echo: false
@@
   final_summary_table,
-  caption = "Final Summary Table of 95% Confidence Intervals for Common States",
   col.names = c("State", "DEM Mean (%)", "DEM Lower 95%", "DEM Upper 95%", "REP Mean (%)", "REP Lower 95%", "REP Upper 95%"),
@@
 #| label: fig-points-over-time
-#| tbl-cap: "Parties Approval and Leaning Preferences"
+#| fig-cap: "Parties Approval and Leaning Preferences"
 #| echo: false
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 

### 2024-11-04 paper/paper.qmd
Minor fixes in appendix A

@@
 For the 2024 USA Presidential Election pollster, the Emerson College Polling defined the population to be "likely voters" of 2024 United States Presidential Election.
-Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modelling and the specific model is not indicated, so the specific validity of the targeted population cannot be determine (@Polling).
+Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modelling and the specific model is not indicated, so the specific validity of the targeted population cannot be determined (@Polling).
 
@@
 From the methodology of recruiting people indicated in the previous paragraph, we can see that while the targeted population of the national poll are likely voters, the sampling frame are likely voters that use phones, either mobile or landline.
-Emerson College Polling chooses 1000 samples from their targeted population, in this case is all likely voters of the 2024 United States Presidential Election, and it is then "weighted by gender, education, race, age, party affiliation, and region based on 2024 likely voter modeling" (@Polling) for each national poll.
+Emerson College Polling chooses 1000 samples from their targeted population of likely voters of the 2024 United States Presidential Election, and it is then "weighted by gender, education, race, age, party affiliation, and region based on 2024 likely voter modeling" (@Polling) for each national poll.
 As most people in the United States use phones by some means, the target population and sampling frame are highly similar, increasing the overall validity of the survey.
@@
 Random sampling tends to reduce bias, simplify analysis, and is also easy to implement, but at the same time also requires much time and money.
-In consideration with the sample size of 1000 people compared to the targeted population of likely voters, which contains most of the population of the United States of America, there is a high possibility of existing selection bias, occurring when the participants are not representative of the population.
+In consideration with the sample size of 1000 people compared to the targeted population of likely voters, which contains most of the population of the United States of America, there is a high possibility of existing selection bias occurring when the participants are not representative of the population.
 
 There is also no specific indication of how non-response is treated, but the results section of national polls, such as the one from September 29 to October 1, 2024, implicates that non-response are eliminated from final recorded results of the survey.
-Therefore, we cannot ignore the possibility of non-response bias, occurring when non-respondents differ significantly from respondents.
+Therefore, we cannot ignore the possibility of non-response bias occurring when non-respondents differ significantly from respondents.
 For example, such non-response bias occurs when non-respondents are from California, while respondents are from other states except California.
@@
 The questionnaire is generally written in an objective voice, and all questions only allow participants to make only one choice.
-All questions in the questionnaire are relatively-common, including ones about ethnicity, age range, region, and education which collects demographical data, and ones that are related to actual presidential-election predictions, including opinions about Joe Biden's performance, the two major parties (Democratic and Republican), whether s/he would vote, and voting inclinations.
+All questions in the questionnaire are relatively-common, including ones about ethnicity, age range, region, and education which collects demographic data, and ones that are related to actual presidential-election predictions, including opinions about Joe Biden's performance, the two major parties (Democratic and Republican), whether s/he would vote, and voting inclinations.
 The only issue observed in the questionnaire is that it is too lengthy, containing more than 20 questions, which may decrease participants' motivations to complete the survey, even after starting it (@Polling).

### 2024-11-04 paper/paper.qmd
Makes some further edits, related to prose and wording, as well as adding more details, to Appendix A.

@@
 For the 2024 USA Presidential Election pollster, the Emerson College Polling defined the population to be "likely voters" of 2024 United States Presidential Election.
-Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modelling and the specific model is not indicated, so the specific validity of the targeted population cannot be determined (@Polling).
+Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modelling and the specific model is not indicated, so we can't determine the scope of the targeted population (@Polling).
 
@@
 In the "MMS-to-Online" and "IVR(Interactive Voice Response)" approaches, random sampling is used, and in the "Online Opt-in Panel" approach, there is no indication of a specific sampling approach (@AboutUs).
-We can conclude that the general approach used in this specific pollster is random sampling.
+We can conclude that the general approach used in this specific pollster is random sampling, since Emerson Polling first chooses a population frame, and then each individual in this population frame has an equal chance to be chosen.
 Random sampling tends to reduce bias, simplify analysis, and is also easy to implement, but at the same time also requires much time and money.
@@
 Emerson can consider increasing the sample size to make their survey more representative of the targeted population, explicitly describing their definition of the targeted population "likely voters" and the model used to determine characteristics of the targeted population.
-Emerson can also consider changing from simple random to stratified sampling, based on different states or other demographic features based on their targeted population model, making their sample more representative of the population, decrease the length of the questionnaire, and also explicitly explain their treatment of non-response, which will increase the validity of the study.
+Emerson can also consider changing from simple random to stratified sampling (dividing the population into independent 'subsets', then performing random sampling in each group), based on different states or other demographic features based on their targeted population model, making their sample more representative of the population, decrease the length of the questionnaire, and also explicitly explain their treatment of non-response, which will increase the validity of the study.
 Overall, this pollster is still highly valid, and generally reliable for predicting the 2024 United States Presidential Election.

### 2024-11-04 paper/paper.qmd
Completed the introduction section

--- 2024-11-04 paper/paper.qmd
@@
 
-*Results paragraph:* Summarize the key findings of the model, highlighting its predictive accuracy and implications.
+In our models, we identified eight pivotal swing states—Georgia, Michigan, Nevada, New Hampshire, North Carolina, Pennsylvania, and Wisconsin—crucial for predicting the 2024 election outcome.
+We constructed Bayesian generalized linear models to estimate the percentage likelihood of candidates Trump and Harris winning in these states, incorporating education level, gender distribution, ethnic demographics, historical voting patterns, approval ratings of Joe Biden, and current leanings toward Trump or Harris.
+Our election forecast resulted in 6 out of 8 pivotal states, as well as the national polling average, leaning towards the Republican party.
 

### 2024-11-04 paper/paper.qmd
Updated introduction and appendix formatting

@@
 In our models, we identified eight pivotal swing states—Georgia, Michigan, Nevada, New Hampshire, North Carolina, Pennsylvania, and Wisconsin—crucial for predicting the 2024 election outcome.
-We constructed Bayesian generalized linear models to estimate the percentage likelihood of candidates Trump and Harris winning in these states, incorporating education level, gender distribution, ethnic demographics, historical voting patterns, approval ratings of Joe Biden, and current leanings toward Trump or Harris.
+We constructed Bayesian generalized linear models to estimate the percentage likelihood of candidates Trump and Harris winning in these states, incorporating education level, gender distribution, ethnic demographics, historical voting patterns, and current leanings toward Trump or Harris.
 Our election forecast resulted in 6 out of 8 pivotal states, as well as the national polling average, leaning towards the Republican party.
@@
 ## Measurement  
-We chose Emerson as our primary pollster, and the reasoning for this is detailed in @sec-appendix.
+We chose Emerson as our primary pollster, and the reasoning for this is detailed in @sec-appendix-A.
 Emerson's polls offered crucial information, such as polling dates, party affiliations, sample sizes, and the percentage of support for each political party.
@@
 
-# Appendix {#sec-appendix}
-
-## Appendix A: Pollster Methodology Overview {-}
+# Appendix {#sec-appendix-A}
+
+## Pollster Methodology Overview {-}
 Emerson College Polling was initially a college polling exercise and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs).
@@
 
-## Appendix B: Idealized Survey Design for $100K Budget {-}
+# Appendix: Idealized Survey Design for $100K Budget {#sec-appendix-B}
 

### 2024-11-04 paper/paper.qmd
Minor fix in appendix C
Updated the discussion section

@@
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-Discuss what the model reveals about the election forecast and its potential impact on understanding voting behavior.
-
-*Person C*: Discuss limitations of the model and areas for further improvement.
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+## Predictive Insights and Electoral Dynamics
+
+Our model provides insightful predictions on the 2024 U.S.
+presidential election landscape, highlighting both the strengths and potential uncertainties within the swing states.
+By examining the demographic and historical predictors, we observed significant patterns in voter leanings and demographic shifts, particularly in critical swing states such as Georgia, Michigan, and Pennsylvania.
+The model suggests that minor shifts in demographics or campaign focus within these states could lead to substantial impacts on the election outcome.
+For example, Georgia, which exhibits an almost even split in party leanings, underscores the importance of targeted outreach to specific demographics, like college-educated voters and minority groups, which show slight inclinations towards the Democratic party.
+
+## The Role of Demographic Factors in Predicting Party Support
+
+The analysis confirms that demographic variables such as education level, ethnicity, and gender hold substantial predictive value in determining party affiliation and voter preference.
+Our correlation matrix shows that higher education levels, for instance, are more closely associated with Democratic leanings, whereas lower education levels trend towards Republican support.
+Racial demographics also show a clear pattern, with minority groups leaning Democratic and White/Caucasian groups displaying stronger support for Republican candidates.
+Additionally, our results on gender differences show a modest trend, with women generally favoring Democratic candidates and men leaning Republican.
+This dynamic reinforces the importance of gender-specific appeals and messaging in campaign strategies.
+The model's findings imply that campaigns aiming to shift voter sentiment in swing states may benefit from tailored approaches that address the priorities and concerns specific to these demographic groups.
+
+## Weaknesses and Next Steps
+
+While the model provides useful forecasts, there are inherent limitations and biases that may affect its predictions.
+One limitation is the assumption of linearity between predictor variables and vote share, which may not fully capture the complex dynamics of voter preferences in certain regions.
+Furthermore, the model’s independence assumption may overlook interconnected factors among states or demographic groups, particularly as social and economic trends impact neighboring states in similar ways.
+
+The use of historical voting data can also introduce potential biases, especially as the political landscape can shift unexpectedly due to national events or changing candidate appeal.
+The reliance on only Emerson polling data may also introduce biases from potential underrepresentation or overrepresentation of certain demographic groups.
+Future models should try to incorporate multiple polling sources and adjust for potential sampling biases across states.
+
+Moving forward, enhancements could include non-linear models to capture more complex relationships between demographic variables and voter behavior.
+Additionally, incorporating data on emerging voter concerns, such as economic shifts and social issues, would allow the model to better adapt to the evolving political landscape.
+Expanding the dataset to include more recent and varied polling sources may help improve prediction accuracy.
+
+Moreover, applying regularization techniques, such as LASSO or Ridge regression, might reduce overfitting and improve the generalizability of predictions across states.
+Incorporating voter sentiment analysis from social media or news coverage could provide real-time insights into voter sentiment, enabling the model to adapt to the ever-changing political dynamics.
+Lastly, building models that consider interactions among predictors, such as gender and education, could offer a more nuanced view of the factors influencing voter behavior, making predictions more responsive to shifts within pivotal swing states.
 
@@
 
-# Additional data details
+# Appendix {#sec-appendix-C}
 Include any technical details on data cleaning, model diagnostics, and posterior checks.
+
 # References

### 2024-11-04 paper/paper.qmd
Minor fix in introduction section
Updated the README and made consistency changes in the paper

@@
 Presidential Election.
-Our findings indicate a close race, with Republicans leading in key states such as Michigan, Georgia, Nevada, Wisconsin, and Pennsylvania, while Democrats hold a slight edge in Arizona.
+Our findings indicate a close race, with Republicans leading in key states such as Georgia, Nevada, North Carolina, New Hampshire, Wisconsin, and Pennsylvania, while Democrats hold a slight edge in Michigan and Arizona.
 The national polling average also leans towards Republican, suggesting a Republican win in the 2024 Election.
@@
 
-This paper addresses a gap in election forecasting by using recent polling data from key swing states to provide a prediction of the 2024 election outcome.
+This paper uses recent polling data from key swing states to provide a prediction of the 2024 election outcome.
 We aim to determine which candidate—Kamala Harris or Donald Trump—is more likely to win this election.
@@
 
-In our models, we identified eight pivotal swing states—Georgia, Michigan, Nevada, New Hampshire, North Carolina, Pennsylvania, and Wisconsin—crucial for predicting the 2024 election outcome.
+In our models, we identified eight pivotal swing states—Michigan, Georgia, Nevada, New Hampshire, North Carolina, Pennsylvania, Wisconsin, and Arizona—crucial for predicting the 2024 election outcome.
 We constructed Bayesian generalized linear models to estimate the percentage likelihood of candidates Trump and Harris winning in these states, incorporating education level, gender distribution, ethnic demographics, historical voting patterns, and current leanings toward Trump or Harris.
@@
 
-From this analysis, we identified eight critical swing states: Georgia, Michigan, National, Nevada, New Hampshire, North Carolina, Pennsylvania, and Wisconsin.
+From this analysis, we identified eight critical swing states: Michigan, Georgia, Nevada, New Hampshire, North Carolina, Pennsylvania, Wisconsin, and Arizona.
 These states will be central to our election prediction.

### 2024-11-04 paper/paper.qmd
Makes further edits of the paper's qmd to fix errors (wording, being more specific, prose, etc.), mostly in discussions and Appendix A.

@@
 Our model provides insightful predictions on the 2024 U.S.
-presidential election landscape, highlighting both the strengths and potential uncertainties within the swing states.
+presidential election landscape, highlighting both the strengths and potential uncertainties within the swing states, and ultimately providing predictions of results of the 2024 United States of America Presidential Elections.
 By examining the demographic and historical predictors, we observed significant patterns in voter leanings and demographic shifts, particularly in critical swing states such as Georgia, Michigan, and Pennsylvania.
@@
 
-# Appendix {#sec-appendix-A}
-
-## Pollster Methodology Overview {-}
-Emerson College Polling was initially a college polling exercise and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs).
+# Appendix: Pollster Methdology Overview - Weaknesses and Potential Improvements{#sec-appendix-A}
+
+Emerson College Polling initially developed from a classroom polling exercise and is now an "innovative, nationally-ranked" non-partisan polling center (@AboutUs).
 For the 2024 USA Presidential Election pollster, the Emerson College Polling defined the population to be "likely voters" of 2024 United States Presidential Election.
-Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modelling and the specific model is not indicated, so we can't determine the scope of the targeted population (@Polling).
+Emerson College Polling does indicate that the targeted population of "likely voters" is based on "2024 Likely Voter Modeling." However, the methodology for this modeling and the specific model is not indicated, so we can not determine the scope of the targeted population (@Polling).
 
@@
 Finally, the IVR approach involves making automatic telephone calls to selected "likely voters." Participants then answer the survey using their "touch-tone telephone".
-The email approach indicated in Emerson College Polling's official methodology (@AboutUs) is not mentioned in the National Polling, so it can be assumed that this approach is not used in our chosen pollster data.
+The email approach indicated in Emerson College Polling's official methodology (@AboutUs) is not mentioned in the National Polling (@Polling), so it can be assumed that this approach is not used in our chosen pollster data.
 
 From the methodology of recruiting people indicated in the previous paragraph, we can see that while the targeted population of the national poll are likely voters, the sampling frame are likely voters that use phones, either mobile or landline.
-Emerson College Polling chooses 1000 samples from their targeted population of likely voters of the 2024 United States Presidential Election, and it is then "weighted by gender, education, race, age, party affiliation, and region based on 2024 likely voter modeling" (@Polling) for each national poll.
+Here, note that the definition of likely voters is set by Emerson College Polling, and the datasets where each sample is chosen from is described in the three approaches by Emerson's methodologies.
+Emerson College Polling chooses 1000 samples from their targeted population of likely voters of the 2024 United States Presidential Election, and it is then "weighted by gender, education, race, age, party affiliation, and region based on 2024 likely voter modeling" for each national poll.
 As most people in the United States use phones by some means, the target population and sampling frame are highly similar, increasing the overall validity of the survey.

### 2024-11-04 paper/paper.qmd
Makes necessary citations in Appendix A
Polished and added references to packages

@@
 
-Our primary objective was to clean, organize, and analyze U.S.
-presidential election polling data to provide insights into voter preferences by state.
-This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries.
-Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis.
-Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns.
-The table highlights key trends and differences across states in terms of party lean and demographic distribution.
-@tbl-predictors-pct-summary-1 and @tbl-predictors-pct-summary-2 provides a snapshot of these findings.
+Our goal was to clean, organize, and analyze U.S.
+presidential election polling data to highlight voter preferences by state.
+We used **`knitr`** (@knitr) for report generation, ensuring our analysis was clear and well-documented, and **`kableExtra`** (@kableExtra) to enhance table outputs for easy interpretation.
+**`ggcorrplot`** (@ggcorrplot) helped us visualize correlations, revealing relationships between variables, while **`zoo`** (@zoo) efficiently handled time series data, such as shifts in voter preferences.
+The **`here`** (@here) package ensured consistent file paths, making our workflow reproducible.
+This streamlined approach provided a comprehensive, polished analysis of key trends and demographic distributions.
+@tbl-predictors-pct-summary-1 and @tbl-predictors-pct-summary-2 provide a snapshot of these findings, showcasing the results of our organized and thorough data analysis.
 
@@
 ```{r}
-#| label: fig-party-extremes
-#| fig-cap: Demographic Differences Between California and South Dakota
+#| label: fig-selected-predictors
+#| fig-cap: Differences in Selected Predictors Between California and South Dakota
 #| echo: false
@@
 
-# Prepare the poll_summary data frame
+# Update the predictor_columns to exclude gender
+predictor_columns <- c(
+  "What is the highest level of education you have attained? College graduate",
+  "What is the highest level of education you have attained? High school or less",
+  "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
+  "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
+  "Who did you vote for in the 2020 election? Donald Trump",
+  "Who did you vote for in the 2020 election? Joe Biden",
+  "Although you are undecided, which candidate do you lean toward? Donald Trump",
+  "Although you are undecided, which candidate do you lean toward? Kamala Harris"
+)
+
+# Update the short_labels to exclude gender
+short_labels <- c(
+  "College Graduate",
+  "High School or Less",
+  "White or Caucasian",
+  "Hispanic or Latino",
+  "Voted Trump 2020",
+  "Voted Biden 2020",
+  "Leaning Trump",
+  "Leaning Harris"
+)
+
+# Prepare the poll_summary data frame with specific predictors
 poll_summary <- poll_data %>%
-  mutate(party_diff = democrat_pct - republican_pct) %>%
-  select(-c(2, 3, 4, 6, 7, 8, 9, 11)) %>%  # Remove specific columns by index
-  select(-contains('Mode')) %>%             # Remove columns containing 'Mode' in their names
-  select(-contains('registration')) %>%
-  select(-democrat_pct, -republican_pct)    # Exclude 'democrat_pct' and 'republican_pct'
-
-# Function to compute mean differences for all numeric variables
+  select(state, all_of(predictor_columns))
+
+# Function to compute mean differences for selected predictors
 compute_mean_differences <- function(data, blue_state, red_state) {
-  numeric_vars <- data %>%
-    select(-state, -party_diff) %>%
-    select(where(is.numeric)) %>%
-    names()
+  selected_vars <- predictor_columns
 
@@
     filter(state == blue_state) %>%
-    summarise(across(all_of(numeric_vars), mean, na.rm = TRUE))
+    summarise(across(all_of(selected_vars), mean, na.rm = TRUE))
 
@@
     filter(state == red_state) %>%
-    summarise(across(all_of(numeric_vars), mean, na.rm = TRUE))
+    summarise(across(all_of(selected_vars), mean, na.rm = TRUE))
 
   diff_df <- tibble(
-    Variable = numeric_vars,
+    Variable = short_labels,  # Use shortened labels
     Blue_Mean = as.numeric(blue_means[1, ]),
@@
 # Function to prepare data for plotting
-prepare_plot_data <- function(diff_df, top_n = 6) {
-  top_diff_vars <- diff_df %>%
-    arrange(desc(Abs_Difference)) %>%
-    slice(1:top_n)
-
-  top_diff_vars %>%
+prepare_plot_data <- function(diff_df) {
+  plot_data <- diff_df %>%
     pivot_longer(
@@
     mutate(Group = recode(Group, "Blue_Mean" = "California", "Red_Mean" = "South Dakota"))
+
+  plot_data
 }
@@
     ) +
-    # Update the colors and labels for the groups
+    # Customize the colors and labels for the groups
     scale_fill_manual(values = c("California" = "blue", "South Dakota" = "red")) +
     labs(
-      title = "Demographic Differences Between California and South Dakota",
-      x = "Demographic Variable",
+      title = "Differences in Selected Predictors Between California and South Dakota",
+      x = "Predictor Variable",
       y = "Mean Percentage",
@@
     theme(
-      axis.text.x = element_text(angle = 45, hjust = 1, size = 4)
+      axis.text.x = element_text(angle = 45, hjust = 1, size = 10)
     )
@@
 # Main analysis function
-analyze_state_differences <- function(data, blue_state, red_state, top_n_vars = 5) {
+analyze_selected_predictors <- function(data, blue_state, red_state) {
   diff_df <- compute_mean_differences(data, blue_state, red_state)
 
-  plot_data <- prepare_plot_data(diff_df, top_n = top_n_vars)
+  plot_data <- prepare_plot_data(diff_df)
 
@@
     plot = plot,
-    top_variables = diff_df %>% arrange(desc(Abs_Difference)) %>% slice(1:top_n_vars)
+    differences = diff_df
   )
@@
 
-# Ensure 'state' column exists in poll_summary
-if (!"state" %in% colnames(poll_summary)) {
-  stop("The 'state' column does not exist in poll_summary.")
-}
-
 # Run the analysis for California and South Dakota
-analysis_results <- analyze_state_differences(poll_summary, blue_state = "California", red_state = "South Dakota", top_n_vars = 6)
+analysis_results <- analyze_selected_predictors(poll_summary, blue_state = "California", red_state = "South Dakota")
 
@@
 analysis_results$plot
+
 ```
@@
 ```
-
-```{r}
-#| label: fig-predictors-party-pct-2
-#| fig-cap: Example Voting and Leaning Predictor Variable Visualization
-#| echo: false
-#| warning: false
-#| message: false
-
-# Calculate poll_summary_2 with rounding to the nearest tenth
-poll_summary_2 <- poll_data %>%
-  group_by(state) %>%
-  summarise(
-    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1),
-    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1),
-    vote_trump_pct = round(mean(.data[[column_refs$vote_trump]], na.rm = TRUE), 1),
-    vote_biden_pct = round(mean(.data[[column_refs$vote_biden]], na.rm = TRUE), 1),
-    lean_trump_pct = round(mean(.data[[column_refs$lean_trump]], na.rm = TRUE), 1),
-    lean_kamala_pct = round(mean(.data[[column_refs$lean_Kamala]], na.rm = TRUE), 1)
-  ) %>%
-  mutate(party_diff = round(dem_total_pct - rep_total_pct, 1)) %>%
-  drop_na()
-
-# Select relevant numeric columns for correlation matrix, excluding PartyDifference
-corr_data_2 <- poll_summary_2 %>%
-  select(
-    Democrat = dem_total_pct,
-    Republican = rep_total_pct,
-    VotedTrump2020 = vote_trump_pct,
-    VotedBiden2020 = vote_biden_pct,
-    LeanTrump = lean_trump_pct,
-    LeanKamala = lean_kamala_pct
-  )
-
-# Calculate the correlation matrix
-cor_matrix_2 <- cor(corr_data_2, use = "complete.obs")
-
-# Set correlations less than -0.9 to NA
-cor_matrix_2[cor_matrix_2 < -0.9] <- NA
-
-# Visualize the correlation matrix without hierarchical clustering
-ggcorrplot(cor_matrix_2, 
-           hc.order = FALSE,  
-           type = "lower", 
-           lab = TRUE, 
-           lab_size = 2,  # Size for correlation labels
-           title = "Correlation Matrix of Voting Preferences and State Lean",
-           colors = c("blue", "white", "red"),  
-           legend.title = "Correlation") +
-  theme(
-    plot.title = element_text(size = 10, hjust = 0.5),  # Title size and centered
-    axis.text = element_text(size = 1),  # Axis text size
-    legend.text = element_text(size = 10),  # Legend text size
-    legend.title = element_text(size = 10),  # Legend title size
-    axis.title = element_text(size = 1)  # Axis labels size
-  )
-
-```
-
 The correlation matrix in @fig-predictors-party-pct shows how different predictors relate to party preferences.
@@
 Blue shading indicates stronger Democratic support, while red shows stronger Republican support.
-Key observations include higher education being linked to Democratic support and lower education favoring Trump in 2020.
+Key observations include higher education being linked to Democratic support and lower education favoring Republican.
 Minority groups generally lean Democratic, while Caucasian/White voters tend to support Trump.
-Past voter behavior shows Biden supporters remain Democratic, and Trump supporters disapprove of Biden.
 Additionally, women lean slightly more Democratic compared to men, who often lean Republican.
@@
 
-Lastly, we examine Democratic and Republican support alongside Biden's approval over time, marking July 21st when Joe Biden concluded his campaign, as reported by @globalnews2024.
-During this period, Democratic support was initially low but increased following this change.
+Lastly, we examine Democratic and Republican support, along with the population's leanings toward Kamala Harris or Donald Trump over time, starting from July 21st, when Joe Biden concluded his campaign, as reported by @globalnews2024.
+During this period, Democratic support has been decreasing steadily since September.
+Trump lost some support in September but gained it back, surpassing Kamala by early October.
+As the election approaches, both Democrats and Republicans have similar levels of overall support; however, the differences between support for Trump and Harris remain significant.
 By analyzing specific demographic factors and previous voting behavior, we aim to predict the 2024 election outcome.
@@
 - **Priors (Past Voting Behavior)**: This looks at who people voted for in the past, helping us understand ongoing political support or disapproval.
-- **Biden's Approval Rating**: This assesses how approval or disapproval of Joe Biden impacts Democratic support, providing insight into shifting voter sentiments.
+- **Population Leaning Towards Harris**: This assesses how the population is feeling at the moment regarding who they may vote for, providing insight into shifting voter sentiments.
+- **Population Leaning Towards Trump**: This assesses how the population is feeling at the moment regarding who they may vote for, providing insight into shifting voter sentiments.
 
@@
 
-We constructed two separate Bayesian generalized linear models to estimate the percentage likelihood of Trump and Harris winning in each swing state.
-Our models look at demographic and historical data that let us to generate probabilistic predictions for both candidates.
+We constructed two separate Bayesian linear models to estimate the percentage likelihood of Trump and Harris winning in each swing state.
+Using **`brms`** (@brms) for Bayesian regression modeling, our models included demographic and historical data, allowing us to generate probabilistic predictions for both candidates.
 These models provide insights into which candidate is more likely to secure a win, even in closely contested states.
@@
 Sensible default priors were adjusted to reflect our prior knowledge about voting behavior, and we conducted posterior predictive checks to assess the models' performance.
-
-## Results and Interpretation
-
-The results, summarized in Table @tbl-modelresults, highlight the predicted vote shares for each candidate in the swing states.
-For each state, we report the posterior mean, as well as 95% credible intervals, to capture the range of plausible outcomes.
-Our findings underscore the impact of demographic and historical factors on voting behavior, providing a nuanced understanding of which states are likely to be battlegrounds in the upcoming election.
-
-By using a Bayesian approach, we not only make point estimates but also quantify the uncertainty surrounding these predictions.
-This is crucial for election forecasting, where even small shifts in voter behavior can have significant implications.
-Our analysis will inform campaign strategies and identify states where mobilizing specific demographics could be pivotal.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+The results, summarized in Table @tbl-modelresults, highlight the predicted vote shares for each candidate in the swing states.
+For each state, we report the posterior mean, as well as 95% credible intervals, to capture the range of plausible outcomes.
+Our findings underscore the impact of demographic and historical factors on voting behavior, providing a nuanced understanding of which states are likely to be battlegrounds in the upcoming election.
+
+By using a Bayesian approach, we not only make point estimates but also quantify the uncertainty surrounding these predictions.
+This is crucial for election forecasting, where even small shifts in voter behavior can have significant implications.
+Our analysis will inform campaign strategies and identify states where mobilizing specific demographics could be pivotal.
 
@@
 Emerson can also consider changing from simple random to stratified sampling (dividing the population into independent 'subsets', then performing random sampling in each group), based on different states or other demographic features based on their targeted population model, making their sample more representative of the population, decrease the length of the questionnaire, and also explicitly explain their treatment of non-response, which will increase the validity of the study.
-Overall, this pollster is still highly valid, and generally reliable for predicting the 2024 United States Presidential Election.
+Overall, this pollster is still highly valid, and generally reliable for predicting the 2024 United States Presidential Election (@tellingstories).
 
@@
 
-# Appendix {#sec-appendix-C}
-Include any technical details on data cleaning, model diagnostics, and posterior checks.
+# Appendix: Data Cleaning {#sec-appendix-C}
+The focus was cleaning up and organizing poll data from different states to create a clear and detailed timeline using data from Google Sheets and Excel files.
+To make this process easier and more efficient, we used several packages that helped us download, clean, and organize the information so it could be easily used later.
+
+First, we used the **`googledrive`** package (@googledrive), which allowed our code to connect to Google Sheets and download the poll data.
+This made it possible for us to get the most up-to-date information.
+Once we had the data, we used another set of packages called **`tidyverse`** (@tidyverse), which includes **`dplyr`** (@dplyr) and **`readr`** (@readr), to clean up the information and organize it.
+For data coming from Excel files, we used the **`readxl`** package (@readxl), which made it easy to read and work with the information from those files.
+
+We also used **`lubridate`** (@lubridate) to make sure dates were handled correctly, so the timeline made sense and everything was in order from oldest to newest.
+
+The **`janitor`** package (@janitor) was helpful for cleaning up column names.
+We also used **`testthat`** (@testthat) to test our data, making sure everything was correct and worked as expected.
+Finally, we used a package called **`arrow`** (@arrow) to save the data in a file format called Parquet, which makes the data smaller and faster to work with.
 

### 2024-11-04 scripts/03-3-api_extract_clean_data.R

@@ -12,12 +12,10 @@
 library(googledrive)
 library(readxl)
 library(tidyverse)
-library(dplyr)
-library(dplyr)
 library(lubridate)
-library(readr)
 library(janitor)
 library(arrow)
+
 state_data <- read_csv("data/02-analysis_data/pollster_data/pollster_link_data.csv")
 #### Helper Functions ####
 

### 2024-11-04 scripts/04-test_analysis_data.R

@@ -1,69 +1,132 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Testing cleaned pollster data for the 2024 US presidential election,
+#          ensuring data quality and consistency with expected formats
+# Author: Gadiel David Flores
+# Date: October 29, 2024
+# Contact: davidgadiel.flores@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# 
+# Pre-requisites:
+# - The `tidyverse`, `lubridate`, and `testthat` packages must be installed and loaded.
+# - Ensure that the following scripts have been run in the specified order:
+#   1. 02-download_data.R
+#   2. 03-1-clean_data.R
+#   3. 03-2-clean-links-state-data.R
+#   4. 03-3-api_extract_clean_data.R
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(lubridate)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+# Load the analysis data
+analysis_data <- read_csv(here::here("data/02-analysis_data/timeline_data.csv"))
 
+# Test Data Structure
 
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+test_that("Data has correct structure and identifiers", {
+  # Test for required columns
+  expected_cols <- c("state", "party", "start_date", "pct", "end_date")
+  expect_true(all(expected_cols %in% colnames(analysis_data)), 
+              info = "Missing one or more expected columns.")
+  
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
+# Test Data Types and Formats
+
+test_that("Variables have correct data types", {
+  # Test numeric variables
+  numeric_vars <- c("pct")
+  for (var in numeric_vars) {
+    expect_true(is.numeric(analysis_data[[var]]), 
+                info = paste(var, "should be numeric"))
+  }
+  
+  # Test character variables
[…]

### 2024-11-04 paper/paper.qmd
Fixed some wording

@@
 - **Model Structure**: Each model incorporates an intercept to represent baseline support and predictor coefficients to quantify the effect of each variable on the vote share.
-We use a Beta distribution to account for the bounded nature of percentage outcomes, ensuring our predictions remain between 0 and 100%.
 
@@
 2. **Independence**: We assume that vote share predictions for each state are independent, conditional on the included predictors.
-3.
-**Non-Normality**: Given the non-normal distribution of our response variables, using a generalized linear model is more appropriate.
-Our analysis revealed that the distribution of Democratic support resembles a Poisson-like distribution, while Republican support shows multimodal characteristics.
+3. **Non-Normality**: Given the non-normal distribution of our response variables, using a generalized linear model is more appropriate.
 4. **Priors**: We applied weakly informative priors to regularize our predictions, with specific priors for those who voted for Trump and Biden in 2020, aligning with historical voting trends.
@@
 
-We implemented our models using the `brms` package in R, which allows for flexible Bayesian regression modeling.
 The models were fit with four chains, running 6,000 iterations each, to ensure convergence and reliable posterior estimates.
@@
 
-The results, summarized in Table @tbl-modelresults, highlight the predicted vote shares for each candidate in the swing states.
-For each state, we report the posterior mean, as well as 95% credible intervals, to capture the range of plausible outcomes.
-Our findings underscore the impact of demographic and historical factors on voting behavior, providing a nuanced understanding of which states are likely to be battlegrounds in the upcoming election.
+The results, summarized in Table @tbl-modelresults, show the predicted vote shares for each candidate in the swing states.
+Our findings show the impact of demographic and historical factors on voting behavior.
 
@@
 This is crucial for election forecasting, where even small shifts in voter behavior can have significant implications.
-Our analysis will inform campaign strategies and identify states where mobilizing specific demographics could be pivotal.
 
